### 2015-04-28 article-00-headers.tex
Initial ShareLaTeX Import

\documentclass[envcountsame]{llncs}

\usepackage{makeidx}    % allows indexgeneration
\usepackage{graphicx}   % allows \includegraphics
\usepackage{cite}       % enables autonumbering in cites

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Additional libraries
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{color}

%
% LISTINGS
%

\usepackage{listings} % see manual: ftp://ftp.funet.fi/pub/TeX/CTAN/macros/latex/contrib/listings/listings.pdf

\lstset{language=TeX,
backgroundcolor=\color[gray]{0.9},
numberstyle=\scriptsize,
basicstyle=\ttfamily\scriptsize,
numbers=none,
stepnumber=1,
frame=none,
breaklines=true,
breakautoindent=true}

\AtBeginDocument{%
  \renewcommand{\thelstlisting}{\arabic{lstlisting}}
}

\spnewtheorem{ontologyRule}{Ontology Rule}{\bfseries}{\rmfamily}

\spnewtheorem{datasetsRule}{Datasets Rule}{\bfseries}{\rmfamily}

%
% TABLES
%

\usepackage{tabularx} % see manual: http://en.wikibooks.org/wiki/LaTeX/Tables#The_tabularx_package

%\usepackage{multirow}

%\usepackage{svg}        % allows adding scalable vector graphic files

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% User-defined commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \newcommand{\definition}[1]{\textbf{#1}}
%\newcommand{\name}[1]{\emph{#1}}
\newcommand{\name}[1]{\textsf{\footnotesize{}#1}}
\newcommand{\TODO}[1]{\textbf{TODO: }\textit{#1}}
\newcommand{\REVIEW}[1]{\textit{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DOCUMENT
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
%
\frontmatter          % for the preliminaries
%
\pagestyle{headings}  % switches on printing of running heads
\addtocmark{IFC-EXPRESS-to-OWL-2} % additional mark in the TOC
%

%\vspace{1cm}
%
\title{Publishing Datasets about Building Designs \\Based on a Multilayer Building Ontology}
%
\titlerunning{Building ontology}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Nam Vu Hoang \and Seppo T\"orm\"a}
%
%\authorrunning{} % abbreviated author list (for running head)
%
\institute{Department of Computer Science, \\
Aalto University School of Science, 
Espoo, Finland\\
\email{\{nam.vuhoang,seppo.torma\}@aalto.fi}}

\maketitle              % typeset the title of the contribution

\begin{abstract}

The volume of data produced in the construction industry is growing all the time. However, building data is currently much less accessible than other smart city data related to geography, traffic, or services. We present an ontology derived from the conceptualization of IFC, the existing standard for representing building designs, and a tool to convert building designs from IFC files to RDF datasets that conform to the ontology. Replacing the traditional point-to-point exchange of IFC files with publication of RDF datasets  by using Linked Data principles and tools enables information integration across different IFC-based and other smart city datasets, and the granular access to individual components. Moreover, it paves the way to a completely distributed and loosely coupled architecture for building information management. Finally, opening selected parts of construction data to a wider set of stakeholders creates opportunities for new and innovative applications in the smart cities of the future.

\keywords{IFC, BIM, Ontology, Linked Data, OWL, RDF}

\end{abstract}

\input{article-01-introduction.tex}

\input{article-02-main.tex}

\input{article-03-bibliography.tex}

\clearpage
\addtocmark[2]{Author Index} % additional numbered TOC entry
\renewcommand{\indexname}{Author Index}
\printindex
\clearpage
\addtocmark[2]{Subject Index} % additional numbered TOC entry
\markboth{Subject Index}{Subject Index}
\renewcommand{\indexname}{Subject Index}
% \input{subjidx.ind}
\end{document}

### 2015-04-28 article-01-introduction.tex
Initial ShareLaTeX Import

\section{Introduction}
\label{sec:Introduction}

Construction industry is becoming increasingly data intensive with the growing use of
information and communication technology in all phases of a building life-cycle. The design phase is
carried out with Building Information Modeling (BIM) tools and systems that produce elaborate
object-oriented 3D models, and the construction phase with a wide range of construction planning,
scheduling, and coordination systems. In the operational stage the building is monitored,
controlled, and maintained based on sensor data and human observations.

In the future, buildings will act as hubs of a wide variety of real-time and historic data about
spaces, indoor locations, indoor routes, assets, people, sensors, materials, equipment, connections
to surrounding networks (traffic, infrastructure, communication), and so on.

A huge amount of design information is already created in the building design phase 
using BIM tools and there is a widely adopted standard, IFC (Industry Foundation Classes) to
represent building designs. IFC has been specified to improve interoperability between different
disciplines in the design phase of a project, including the architecture, structural engineering,
MEP engineering models, and so on. Nowadays, the IFC standard is supported by all the major BIM
vendors; their BIM tools can export building models as IFC files that can be exchanged with other
parties.

The design models can provide the framework with which the information gathered from other sources
can be organized and visualized.  However, despite the IFC standard, the design information is currently
difficult to access and utilize. The sharing of building models based on file exchange does not work
sufficiently well in practice. Utilization of related information from different models requires
repeated manual work from the designers. In the longer run, file exchange cannot provide a fruitful
basis for open utilization of building information and emergence of new, innovative applications.

We have developed a WebIFC converter to derive an OWL ontology from the IFC schema (called ifcOWL),
and RDF datasets from IFC files (called ifcRDF).

The ifcOWL ontology is used to classify concepts, characterize possible relationships, and define possible constraints of data in ifcRDF datasets which replicate the content of original IFC-STEP data files. Thus, all terms defined in the ifcOWL ontology must be based on the corresponding constructs of the original IFC-EXPRESS schema. Ideally, there should be only one ifcOWL ontology for one IFC schema (see Fig. \ref{fig:ifcOWL-single}). However, taking account that ifcRDF datasets are automatically generated, data validation is almost never required. So in most of cases it is wisely to use a lightweight ontology instead of a detailed one in order to facilitate more effective reasoning.

\begin{figure}[h]
\centering
\includegraphics[width=0.75\textwidth]{images/ifcOWL-single.png}
\caption{A single ifcOWL ontology}
\label{fig:ifcOWL-single}
\end{figure}

Various levels of details of ifcOWL ontologies are defined in the Table \ref{tab:ifcOWL-levels}. An ifcOWL ontology version can be considered as extension of all ontology versions with smaller level numbers.

\begin{table}[h]
\centering
\footnotesize
\begin{tabularx}{\textwidth}{|X|X|X|}
\hline
\textbf{Level (name)} & \textbf{Detail} & \textbf{Note}   \\

\hline
0 (None) & Schema meta data &   \\

\hline
1 & All type names &   \\

\hline
2 & Entity type hierarchy &   \\

\hline
3 & Enumeration \& Select data types & \\

\hline
4 (Lite) & Property names and value types &   \\

\hline
5 & Keys &   \\

\hline
6 (Standard) & Inverse properties &   \\

\hline
7 & Property cardinality restrictions  &   \\

\hline
8 (Advanced) & List type cardinality restrictions  &   \\

\hline
9 (Full) & Property value types with exact value ranges  & Not supported by the converter  \\

\hline
\end{tabularx}
\caption{Levels of details of ifcOWL ontology}
\label{tab:ifcOWL-levels}
\end{table}

Having multiple versions of the same ontology and knowing exactly that every ifcRDF dataset will be compatible with any related ifcOWL ontology (see Fig. \ref{fig:ifcOWL-multilayered}) user can choose the best one based on the requirements of the concrete use case and tools used. For instance, when it is required to query entity properties, extract the type hierarchy, or convert ifcRDF datasets back to object-oriented data models it is enough to use \textbf{ifcOWL-Lite} ontology. It case of reasoning about data including inverse properties, it is recommended to use \textbf{ifcOWL-Standard} ontology. When data validation, for example, checking type and cardinality constraints is needed \textbf{ifcOWL-Advanced} ontology is the most reasonable choice.

\begin{figure}[h]
\centering
\includegraphics[width=0.75\textwidth]{images/ifcOWL-multilayer.png}
\caption{A multilayer ifcOWL ontology}
\label{fig:ifcOWL-multilayered}
\end{figure}

### 2015-04-28 article-02-main.tex
Initial ShareLaTeX Import

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Converting EXPRESS Schemas to OWL
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Generating ifcOWL Ontologies From IFC Specifications}

\subsection{Common Conversion Rules}

\TODO{Explain: why an OWL 2 profile has to be being chosen, how that choice will impact to the reasoning and querying performance, whether we can use other OWL 2 EL or QL for ifcOWL-Lite ontology}

\begin{ontologyRule}[OWL profile, ifcOWL layers: all]
All ifc\-OWL ontologies must be based on OWL 2 RL profile\footnote{Our IFC-EXPRESS-to-OWL converter tool which is available on http://drumbeat.cs.hut.fi also allows choosing other OWL profiles: OWL Lite, OWL DL, OWL Full, OWL 2 EL, OWL 2 QL, OWL 2 Full}.
\end{ontologyRule}

This profile has been chosen because in contrast to OWL 2 EL and OWL 2 QL, it supports such needed things as: enumerations of individuals (see \ref{ifcOWL:enum-types}), named individuals (see \ref{ifcRDF:naming-individuals}), etc.\cite{w3c:owl2-profiles}.

\begin{ontologyRule}[Domain and range information, ifcOWL layers: all]
Property domain and range constraints specified by \name{rdfs:domain} and \name{rdfs:range} should not be included in ifcOWL ontologies.
\end{ontologyRule}

Properties \name{rdfs:\-domain} and \name{rdfs:\-range} are mainly used to infer types, but not to check type constraints like in programming languages \cite{w3c:owl-guide}. Moreover, the basic facilities given by them do not provide any direct way to indicate property restrictions that are local to a class \cite{w3c:rdf-schema}. Although it is possible to combine use \name{rdfs:domain} and \name{rdfs:range} with sub-property hierarchies, direct support for such declarations is provided by OWL constructs, e.g. \name{owl:\-onProperty}.

\begin{ontologyRule}[Namespaces, ifcOWL layers: all]
All constructs related purely to EXPRESS, STEP or IFC specifications must be belong to namespaces with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly. At this moment the namespaces are defined as in Listing \ref{ifcOWL-namespaces}. Instead of \name{IFCXXX} there should be IFC versions, e.g. \name{IFC2x3} or \name{IFC4}.
\end{ontologyRule}

\begin{lstlisting}[caption={Namespace definitions},label=ifcOWL-namespaces]
@prefix expr: <http://drumbeat.cs.hut.fi/owl/EXPRESS#>
@prefix step: <http://drumbeat.cs.hut.fi/owl/STEP#>
@prefix ifc: <http://drumbeat.cs.hut.fi/owl/IFCXXX#>
\end{lstlisting}

\subsection{Components of an EXPRESS schema}
The IFC specification is written as an EXPRESS schema. An EXPRESS schema defines series of data types, functions and rules by using EXPRESS data definition language specified in ISO10303-11 \cite{wiki:express,noauthor:ifc-guide}.

There are six groups of data types: simple data types, entity data types, enumeration data types, defined (declared) data types, select data types and aggregation data types. All types must be declared obviously by using constructs TYPE and ENTITY except the built-in simple data types. Definitions and characteristics of data types are considered below. EXPRESS schemata also contain definitions of functions, rules and type constraints most of which cannot be reflected in OWL 2 at all.

\subsection{Simple Data Type Declarations}
\label{ifcOWL:simple-type}

Simple data types built in EXPRESS are: \name{STRING}, \name{BI\-NA\-RY}, \name{IN\-TE\-GER}, \name{REAL}, \name{NUM\-BER}, \name{BOO\-LEAN} and \name{LO\-GI\-CAL}. However, type \name{LOGICAL} which has three possible values \name{TRUE}, \name{FALSE} and \name{UNKNOWN} is considered by us as an enumeration data type (see \ref{ifcOWL:enum-types}). The same to type \name{BOOLEAN} which is like a primitive boolean data type but semantically close and met with the same frequency in IFC4 as \name{LOGICAL}.

On the other hand, there are two defined data types in IFC (not in EXPRESS) that also can be referred to the simple-data-type group: \name{Ifc\-Time\-Stamp} which is a so-called TimeSpan data type and can be easily converted to \name{xsd:date\-Time}, and \name{Ifc\-Globally\-Unique\-Id} which represents set of compressed Globally Unique Identifiers (GUIDs), i.e. strings with fixed length (see Listing \ref{ifcOWL-additional-simple-types}).

\begin{lstlisting}[caption={Two defined data types considered as simple data types},label=ifcOWL-additional-simple-types]
(* number of seconds elapsed since the beginning of the year 1970 *)
TYPE IfcTimeStamp = INTEGER; END_TYPE;

(* compressed GUID *)
TYPE IfcGloballyUniqueId = STRING(22) FIXED; END_TYPE;
\end{lstlisting}

\begin{ontologyRule}[Simple data types, ifcOWL layers: all]
All simple data types except \emph{BOOLEAN} and \emph{LOGICAL} are converted to OWL classes which have only one property \name{rdf:value}. The range of that property should be the most similar primitive XSD type supported by OWL 2 \cite{w3c:owl2-profiles} and recommended to use in Linked Data Platform Best Practices and Guideline \cite{w3c:ldp-best-practices}. For \name{STRING} and \name{Ifc\-Globally\-UniqueId} it is \name{xsd:\-string}, for \name{BINARY} -- \name{xsd:\-hexBinary}, for \name{INTEGER} -- \name{xsd:\-integer}, for \name{REAL} and \name{NUMBER} -- \name{xsd:\-double}, for \name{Ifc\-Time\-Stamp} -- \name{xsd:\-dateTime} (see example on Listing \ref{lst:ifcOWL-simple-types}).
\end{ontologyRule}

\begin{lstlisting}[caption={Simple data types as instances of \name{owl:Class}}, label=lst:ifcOWL-simple-types]
expr:STRING a owl:Class;
    rdfs:subClassOf [ a owl:Restriction;
                      owl:onProperty rdf:value;
                      owl:allValuesFrom xsd:double; ] .
ifc:IfcTimeStamp a owl:Class;
    rdfs:subClassOf [ a owl:Restriction;
                      owl:onProperty rdf:value;
                      owl:allValuesFrom xsd:dateTime; ]
\end{lstlisting}

The selected XSD types conform to the value spaces of the original EXPRESS and IFC data types better than other ones. For instance, type \name{xsd:double} is better than \name{xsd:decimal} as it is an IEEE 64-bit floating-point data type and supports scientific notation \cite{w3c:xsd,datapic:xsd}.

The reason of using class boxing technique is to make simple data types be instances of \name{owl:Class}. Instead of that they could be either simply replaced with primitive XSD data types, or defined as shown in the Listing , but in these cases they will be instances of \name{rdfs:Datatype}.  and cannot be used for union classes

\begin{lstlisting}[caption={Simple data types as instances of \name{rdfs:Datattype}},label=lst:ifcOWL-simple-types-2]
# if LEVEL_OF_DETAIL is higher than STANDARD
expr:STRING owl:equivalentClass [
    rdf:type        rdfs:Datatype ;
    owl:onDatatype  xsd:string ;
    owl:withRestrictions ()
]
\end{lstlisting}

Other considered conversion options were:
\begin{itemize}
\item To replace simple data types with the primitive XSD data types. But in this case all select data types (see \ref{subsec:ifcOWL:select-types}) cannot be defined by means of property \name{owl:unionOf} the range of which

all properties in any OWL profile except OWL Full must be either \name{owl:}
\end{itemize}

Some of simple data types may be specified together with optional size parameter, e.g. \name{STRING(lengthInBytes)} or \name{BINARY(sizeInBits)}. If \name{\$LEVEL\-\_OF\-\_DETAILS} is less than \name{STANDARD} (see \ref{subsec:multilayer-ontology-approach}), these parameters are ignored, i.e. \name{STRING(N)} is considered as equivalent to \name{STRING}. Otherwise additional constraints are added to the primitive data type by using property \name{owl:onDatatype} and data type facets \name{xsd:maxLength} or \name{xsd:length} \cite{w3c:owl2-syntax,w3c:xmlschema11-2} (see Listing \ref{ifcOWL:length-constraint}.

\begin{lstlisting}[caption={Length constraint},label=ifcOWL:length-constraint]
# if LEVEL_OF_DETAIL is higher than STANDARD
expr:STRING22 owl:equivalentClass [
    rdf:type        rdfs:Datatype ;
    owl:onDatatype  xsd:string ;
    owl:withRestrictions (
        [ xsd:length "22"^^xsd:integer ]
    )
]
\end{lstlisting}

\begin{lstlisting}[caption={},label=1]
expr:STRING rdf:type owl:Class;
    rdfs:subClassOf [
        owl:Restriction;
        owl:onProperty expr:hasValue;
        owl:allValuesFrom xsd:string;
    ].

ifc:ifcLabel rdf:type owl:Class;
    rdfs:subClassOf expr:STRING.
\end{lstlisting}

\begin{lstlisting}[caption={},label=1]

% http://answers.semanticweb.com/questions/229/how-to-define-a-own-notation-datatype

http://semantic-web-grundlagen.de/wiki/Guide_to_OWL_2_for_OWL_1_users#Datatypes

http://www2.informatik.hu-berlin.de/~wandelt/SW201213/9OWL2.pdf

Using that in OWL 2 DL is no problem. If you want to stay within one of the profiles though note that defining datatypes through such restrictions isn't allowed in any of them. So for increased interoperability (also with OWL 1 and RDFS) you could just use xsd:string instead. That should be perfectly fine with SKOS.

expr:STRING rdf:type rdfs:Datatype;
 owl:onDatatype xsd:string;
 owl:withRestrictions ().

ifc:ifcLabel rdf:type owl:Class;
    rdfs:subClassOf expr:STRING.
\end{lstlisting}

\subsection{Entity Data Type Declarations}

\subsection{Enumeration Type Declarations}
\label{ifcOWL:enum-types}

\section{Generating ifcRDF Datasets From IFC Data}

\subsection{Naming individuals}
\label{ifcRDF:naming-individuals}

### 2015-04-28 article-02-main.tex
Changed rules for simple datatypes

@@
 \section{Generating ifcOWL Ontologies From IFC Specifications}
+\label{sec:ifcOWL}
 
@@
 All constructs related purely to EXPRESS, STEP or IFC specifications must be belong to namespaces with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly.
-At this moment the namespaces are defined as in Listing \ref{ifcOWL-namespaces}.
+At this moment the namespaces are defined as in Listing \ref{lst:ifcOWL-namespaces}.
 Instead of \name{IFCXXX} there should be IFC versions, e.g.
@@
 
-\begin{lstlisting}[caption={Namespace definitions},label=ifcOWL-namespaces]
+\begin{lstlisting}[caption={Namespace definitions},label=lst:ifcOWL-namespaces]
 @prefix expr: <http://drumbeat.cs.hut.fi/owl/EXPRESS#>
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Components of an EXPRESS schema
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Components of an EXPRESS schema}
 The IFC specification is written as an EXPRESS schema.
-An EXPRESS schema defines series of data types, functions and rules by using EXPRESS data definition language specified in ISO10303-11 \cite{wiki:express,noauthor:ifc-guide}.
+An EXPRESS schema defines series of data\-types, functions and rules by using EXPRESS data definition language specified in ISO10303-11 \cite{wiki:express,noauthor:ifc-guide}.
 
-There are six groups of data types: simple data types, entity data types, enumeration data types, defined (declared) data types, select data types and aggregation data types.
-All types must be declared obviously by using constructs TYPE and ENTITY except the built-in simple data types.
-Definitions and characteristics of data types are considered below.
+There are six groups of data\-types: simple data\-types, entity data\-types, enumeration data\-types, defined (declared) data\-types, select data\-types and aggregation data\-types.
+All types must be declared obviously by using constructs TYPE and ENTITY except the built-in simple data\-types.
+Definitions and characteristics of data\-types are considered below.
 EXPRESS schemata also contain definitions of functions, rules and type constraints most of which cannot be reflected in OWL 2 at all.
 
-\subsection{Simple Data Type Declarations}
-\label{ifcOWL:simple-type}
+\subsection{Simple Datatype Declarations}
+\label{subsec:ifcOWL-simple-type}
 
-Simple data types built in EXPRESS are: \name{STRING}, \name{BI\-NA\-RY}, \name{IN\-TE\-GER}, \name{REAL}, \name{NUM\-BER}, \name{BOO\-LEAN} and \name{LO\-GI\-CAL}.
-However, type \name{LOGICAL} which has three possible values \name{TRUE}, \name{FALSE} and \name{UNKNOWN} is considered by us as an enumeration data type (see \ref{ifcOWL:enum-types}).
-The same to type \name{BOOLEAN} which is like a primitive boolean data type but semantically close and met with the same frequency in IFC4 as \name{LOGICAL}.
+Simple data\-types built in EXPRESS are: \name{STRING}, \name{BI\-NA\-RY}, \name{IN\-TE\-GER}, \name{REAL}, \name{NUM\-BER}, \name{BOO\-LEAN} and \name{LO\-GI\-CAL}.
+However, type \name{LOGICAL} which has three possible values \name{TRUE}, \name{FALSE} and \name{UNKNOWN} is considered by us as an enumeration data\-type (see \ref{ifcOWL:enum-types}).
+The same to type \name{BOOLEAN} which is like a primitive boolean data\-type but semantically close and met with the same frequency in IFC4 as \name{LOGICAL}.
 
-On the other hand, there are two defined data types in IFC (not in EXPRESS) that also can be referred to the simple-data-type group: \name{Ifc\-Time\-Stamp} which is a so-called TimeSpan data type and can be easily converted to \name{xsd:date\-Time}, and \name{Ifc\-Globally\-Unique\-Id} which represents set of compressed Globally Unique Identifiers (GUIDs), i.e.
-strings with fixed length (see Listing \ref{ifcOWL-additional-simple-types}).
+On the other hand, there are two defined data\-types in IFC (not in EXPRESS) that also can be referred to the simple-data-type group: \name{Ifc\-Time\-Stamp} which is a so-called TimeSpan data\-type and can be easily converted to \name{xsd:date\-Time}, and \name{Ifc\-Globally\-Unique\-Id} which represents set of compressed Globally Unique Identifiers (GUIDs), i.e.
+strings with fixed length (see Listing \ref{lst:ifcOWL-additional-simple-types}).
 
-\begin{lstlisting}[caption={Two defined data types considered as simple data types},label=ifcOWL-additional-simple-types]
+\begin{lstlisting}[caption={Two defined data\-types considered as simple data\-types}, label=lst:ifcOWL-additional-simple-types]
 (* number of seconds elapsed since the beginning of the year 1970 *)
@@
 
-\begin{ontologyRule}[Simple data types, ifcOWL layers: all]
-All simple data types except \emph{BOOLEAN} and \emph{LOGICAL} are converted to OWL classes which have only one property \name{rdf:value}.
-The range of that property should be the most similar primitive XSD type supported by OWL 2 \cite{w3c:owl2-profiles} and recommended to use in Linked Data Platform Best Practices and Guideline \cite{w3c:ldp-best-practices}.
-For \name{STRING} and \name{Ifc\-Globally\-UniqueId} it is \name{xsd:\-string}, for \name{BINARY} -- \name{xsd:\-hexBinary}, for \name{INTEGER} -- \name{xsd:\-integer}, for \name{REAL} and \name{NUMBER} -- \name{xsd:\-double}, for \name{Ifc\-Time\-Stamp} -- \name{xsd:\-dateTime} (see example on Listing \ref{lst:ifcOWL-simple-types}).
+\begin{ontologyRule}[Simple data\-types, ifcOWL layers: all]
+All simple data\-types except \emph{BOOLEAN} and \emph{LOGICAL} are the same as their most similar primitive XSD data\-types which are supported by OWL 2 \cite{w3c:owl2-profiles} and are among preferred standard data\-types in Linked Data Platform \cite{w3c:ldp-best-practices} (see Listing \ref{lst:ifcOWL-simple-types}).
+However, it is recommended to use directly XSD s like (e.g.
+\name{xsd:string}) instead of the EXPRESS data\-type constructs (e.g.
+\name{expr:STRING}).
 \end{ontologyRule}
 
-\begin{lstlisting}[caption={Simple data types as instances of \name{owl:Class}}, label=lst:ifcOWL-simple-types]
-expr:STRING a owl:Class;
-    rdfs:subClassOf [ a owl:Restriction;
-                      owl:onProperty rdf:value;
-                      owl:allValuesFrom xsd:double; ] .
-ifc:IfcTimeStamp a owl:Class;
-    rdfs:subClassOf [ a owl:Restriction;
-                      owl:onProperty rdf:value;
-                      owl:allValuesFrom xsd:dateTime; ]
+% For \name{STRING} and \name{Ifc\-Globally\-UniqueId} it is \name{xsd:\-string}, for \name{BINARY} -- \name{xsd:\-hexBinary}, for \name{INTEGER} -- \name{xsd:\-integer}, for \name{REAL} and \name{NUMBER} -- \name{xsd:\-double}, for \name{Ifc\-Time\-Stamp} -- \name{xsd:\-dateTime}.
+
+\begin{lstlisting}[caption={Simple data types}, label=lst:ifcOWL-simple-types]
+expr:STRING, ifc:IfcGlobalUniqueId owl:sameAs xsd:string .
+expr:BINARY owl:sameAs xsd:hexBinary .
+expr:INTEGER owl:sameAs xsd:integer .
+expr:REAL, expr:NUMBER owl:sameAs xsd:double .
+ifc:IfcTimeStamp owl:sameAs xsd:dateTime .
 \end{lstlisting}
 
-The selected XSD types conform to the value spaces of the original EXPRESS and IFC data types better than other ones.
-For instance, type \name{xsd:double} is better than \name{xsd:decimal} as it is an IEEE 64-bit floating-point data type and supports scientific notation \cite{w3c:xsd,datapic:xsd}.
+`Most similar XSD data\-types' mean the types from XSD schema that conform to the value spaces of the original EXPRESS and IFC data\-types better than other ones.
+For instance, type \name{xsd:double} is more similar to \name{REAL} than \name{xsd:decimal} as it is an IEEE 64-bit floating-point data\-type and supports scientific notation \cite{w3c:xsd,datapic:xsd}.
 
-The reason of using class boxing technique is to make simple data types be instances of \name{owl:Class}.
-Instead of that they could be either simply replaced with primitive XSD data types, or defined as shown in the Listing , but in these cases they will be instances of \name{rdfs:Datatype}.
+\begin{ontologyRule}[Simple data\-types with sizes, ifcOWL layers: all]If \name{STRING} and \name{BINARY} data\-types are specified with a size parameter, i.e.
+\name{STRING (max\-Length\-In\-Bytes)} or \name{BINARY(size\-In\-Bits)}, then they will be named in format \name{\textless type\_name \textgreater\textless size\textgreater} (e.g.
+\name{STRING22}).
+In layers \lite{} and \standard{} they are equivalent to data\-types without size.
+But in layer \advanced{}, in order to specify the size parameter, additional constraints must be added by using property \name{owl:onDatatype} and data\-type facets \name{xsd:maxLength} or \name{xsd:length} \cite{w3c:owl2-syntax,w3c:xmlschema11-2} (see example in Listing \ref{lst:ifcOWL-simple-types-with-size}).
+\end{ontologyRule}
+
+\begin{lstlisting}[caption={Size constraint of simple s},
+label=lst:ifcOWL-simple-types-with-size]
+# only for Lite, Standard layers
+expr:STRING22 owl:sameAs expr:STRING .
+
+# only for Advanced layer
+expr:STRING22 owl:equivalentClass
+    [   rdf:type        rdfs:Datatype ;
+        owl:onDatatype  xsd:string ;
+        owl:withRestrictions ( [ xsd:length "22"^^xsd:integer ] ) ] .
+\end{lstlisting}
+
+% \begin{lstlisting}[caption={},label=1]
+% expr:STRING rdf:type owl:Class;
+%     rdfs:subClassOf [
+%         owl:Restriction;
+%         owl:onProperty expr:hasValue;
+%         owl:allValuesFrom xsd:string;
+%     ].
+
+% \begin{lstlisting}[caption={Simple data\-types as instances of \name{owl:Class}}, label=lst:ifcOWL-simple-types]
+% expr:STRING a owl:Class;
+%     rdfs:subClassOf [ a owl:Restriction;
+%                       owl:onProperty rdf:value;
+%                       owl:allValuesFrom xsd:double; ] .
+% ifc:IfcTimeStamp a owl:Class;
+%     rdfs:subClassOf [ a owl:Restriction;
+%                       owl:onProperty rdf:value;
+%                       owl:allValuesFrom xsd:dateTime; ]
+% \end{lstlisting}
+
+% The reason of using class boxing technique is to make simple data\-types be instances of \name{owl:Class}.
+Instead of that they could be either simply replaced with primitive XSD data\-types, or defined as shown in the Listing , but in these cases they will be instances of \name{rdfs:Datatype}.
 and cannot be used for union classes
 
-\begin{lstlisting}[caption={Simple data types as instances of \name{rdfs:Datattype}},label=lst:ifcOWL-simple-types-2]
-# if LEVEL_OF_DETAIL is higher than STANDARD
-expr:STRING owl:equivalentClass [
-    rdf:type        rdfs:Datatype ;
-    owl:onDatatype  xsd:string ;
-    owl:withRestrictions ()
-]
-\end{lstlisting}
+% \begin{lstlisting}[caption={Simple data\-types as instances of \name{rdfs:Datattype}},label=lst:ifcOWL-simple-types-2]
+% # if LEVEL_OF_DETAIL is higher than STANDARD
+% expr:STRING owl:equivalentClass [
+%     rdf:type        rdfs:Datatype ;
+%     owl:onDatatype  xsd:string ;
+%     owl:withRestrictions ()
+% ]
+% \end{lstlisting}
 
-Other considered conversion options were:
-\begin{itemize}
-\item To replace simple data types with the primitive XSD data types.
-But in this case all select data types (see \ref{subsec:ifcOWL:select-types}) cannot be defined by means of property \name{owl:unionOf} the range of which
+% Other considered conversion options were:
+% \begin{itemize}
+% \item To replace simple data\-types with the primitive XSD data\-types.
+But in this case all select data\-types (see \ref{subsec:ifcOWL:select-types}) cannot be defined by means of property \name{owl:unionOf} the range of which
 
-all properties in any OWL profile except OWL Full must be either \name{owl:}
-\end{itemize}
+% all properties in any OWL profile except OWL Full must be either \name{owl:}
+% \end{itemize}
 
-Some of simple data types may be specified together with optional size parameter, e.g.
-\name{STRING(lengthInBytes)} or \name{BINARY(sizeInBits)}.
-If \name{\$LEVEL\-\_OF\-\_DETAILS} is less than \name{STANDARD} (see \ref{subsec:multilayer-ontology-approach}), these parameters are ignored, i.e.
-\name{STRING(N)} is considered as equivalent to \name{STRING}.
-Otherwise additional constraints are added to the primitive data type by using property \name{owl:onDatatype} and data type facets \name{xsd:maxLength} or \name{xsd:length} \cite{w3c:owl2-syntax,w3c:xmlschema11-2} (see Listing \ref{ifcOWL:length-constraint}.
+% ifc:ifcLabel rdf:type owl:Class;
+%     rdfs:subClassOf expr:STRING.
+% \end{lstlisting}
 
-\begin{lstlisting}[caption={Length constraint},label=ifcOWL:length-constraint]
-# if LEVEL_OF_DETAIL is higher than STANDARD
-expr:STRING22 owl:equivalentClass [
-    rdf:type        rdfs:Datatype ;
-    owl:onDatatype  xsd:string ;
-    owl:withRestrictions (
-        [ xsd:length "22"^^xsd:integer ]
-    )
-]
-\end{lstlisting}
+% \begin{lstlisting}[caption={},label=1]
 
-\begin{lstlisting}[caption={},label=1]
-expr:STRING rdf:type owl:Class;
-    rdfs:subClassOf [
-        owl:Restriction;
-        owl:onProperty expr:hasValue;
-        owl:allValuesFrom xsd:string;
-    ].
+% % http://answers.semanticweb.com/questions/229/how-to-define-a-own-notation-data\-type
 
-ifc:ifcLabel rdf:type owl:Class;
-    rdfs:subClassOf expr:STRING.
-\end{lstlisting}
+% http://semantic-web-grundlagen.de/wiki/Guide_to_OWL_2_for_OWL_1_users#Datatypes
 
-\begin{lstlisting}[caption={},label=1]
+% http://www2.informatik.hu-berlin.de/~wandelt/SW201213/9OWL2.pdf
 
-% http://answers.semanticweb.com/questions/229/how-to-define-a-own-notation-datatype
-
-http://semantic-web-grundlagen.de/wiki/Guide_to_OWL_2_for_OWL_1_users#Datatypes
-
-http://www2.informatik.hu-berlin.de/~wandelt/SW201213/9OWL2.pdf
-
-Using that in OWL 2 DL is no problem.
-If you want to stay within one of the profiles though note that defining datatypes through such restrictions isn't allowed in any of them.
+% Using that in OWL 2 DL is no problem.
+If you want to stay within one of the profiles though note that defining data\-types through such restrictions isn't allowed in any of them.
 So for increased interoperability (also with OWL 1 and RDFS) you could just use xsd:string instead.
@@
 
-expr:STRING rdf:type rdfs:Datatype;
- owl:onDatatype xsd:string;
- owl:withRestrictions ().
+% expr:STRING rdf:type rdfs:Datatype;
+%  owl:onDatatype xsd:string;
+%  owl:withRestrictions ().
 
-ifc:ifcLabel rdf:type owl:Class;
-    rdfs:subClassOf expr:STRING.
-\end{lstlisting}
+% ifc:ifcLabel rdf:type owl:Class;
+%     rdfs:subClassOf expr:STRING.
+% \end{lstlisting}
 
-\subsection{Entity Data Type Declarations}
+\subsection{Defined Datatype Declarations}
+\label{subsec:ifcOWL-defined-types}
+
+% \item \textbf{Defined or declared data type}: This further specializes other data\-types — e.g., define a datatype \emph{PositiveInteger} that is of type integer with a positive value.
+
+\subsection{Entity Datatype Declarations}
 

### 2015-04-28 article-02-main.tex
Save draft for select types and defines types

@@
 
-% \begin{lstlisting}[caption={Simple data\-types as instances of \name{owl:Class}}, label=lst:ifcOWL-simple-types]
-% expr:STRING a owl:Class;
-%     rdfs:subClassOf [ a owl:Restriction;
-%                       owl:onProperty rdf:value;
-%                       owl:allValuesFrom xsd:double; ] .
-% ifc:IfcTimeStamp a owl:Class;
-%     rdfs:subClassOf [ a owl:Restriction;
-%                       owl:onProperty rdf:value;
-%                       owl:allValuesFrom xsd:dateTime; ]
-% \end{lstlisting}
-
 % The reason of using class boxing technique is to make simple data\-types be instances of \name{owl:Class}.
@@
 
+\subsection{Select Datatype Declarations}
+\label{subsec:ifcOWL-select-types}
+
+A select data\-type represents an union of some defined, entity data\-types.
+For instance, data\-type \name{Ifc\-Trim\-ming\-Se\-lect} is an union of entity data\-type \name{Ifc\-Car\-te\-sian\-Point} and defined data\-type \name{Ifc\-Pa\-ra\-me\-ter\-Va\-lue}.
+
+\begin{ontologyRule}[caption={Defined data\-types}]
+A select data\-type is converted to an \name{owl:Class} that is equivalent to a union of \name{owl:Class} statements (see example on Listing \ref{lst:ifcOWL-select-types}).
+\end{ontologyRule}
+
+\begin{lstlisting}[caption={Select data\-types, label=lst:ifcOWL-select-types]
+ifc:IfcTrimmingSelect owl:equivalentClass [
+   rdf:type     owl:Class ;
+   owl:unionOf  ( ifc:IfcCartesianPoint ifc:IfcparameterValue ) ] .
+\end{lstlisting}
+
+\TODO{Explain what if another OWL profile is used, and it doesn't support owl:unionOf --> use rdfs:subClassOf}
+
 \subsection{Defined Datatype Declarations}
 \label{subsec:ifcOWL-defined-types}
+
+A defined (declared, named) data\-type is usually specified as equivalent to a simple data\-type or another defined data\-type, with or without a constraint.
+For instance, type \name{Ifc\-Length\-Mea\-sure} is equal to \name{REAL}, and type \name{Ifc\-Po\-si\-ti\-ve\-Length\-Mea\-su\-re} is a set of positive \name{Ifc\-Length\-Mea\-sure} values.
+
+In contrast to simple data\-types, defined data\-types can be used for declaration of select data\-types (see above), and most of defined data\-types have meaningful names.
+\TODO{When an entity property has select type which is a defined type...}
+%Moreover, only a few of entity type properties in IFC are specified with simple data type.
+
+\begin{ontologyRule}[caption={Defined data\-types}]
+A defined data\-type is converted to an \name{owl:Class} with a single property \name{rdf:value} that has the underlying data\-type of the defined data\-type (see example on Listing \ref{lst:ifcOWL-defined-types}).
+\end{ontologyRule}
+
+\begin{lstlisting}[caption={Defined data\-types}, label=lst:ifcOWL-defined-types]
+ifc:IfcAreaMeasure a owl:Class;
+    rdfs:subClassOf [ a owl:Restriction;
+                      owl:onProperty rdf:value;
+                      owl:allValuesFrom xsd:double; ] .
+\end{lstlisting}
+
+Defined data\-types must be declared as wrapper classes because they must be instances of \name{owl:Class} can be used as a class-option of an union class (see \ref{subsec:ifcOWL-select-types}).
 
@@
 \subsection{Entity Datatype Declarations}
+\label{subsec:ifcOWL-entity-types}
+
+Each entity data\-type represents a category of entity instances (entities).
+The main relationships between entities are inheritance and association (aggregation is considered as a special case of association).
+
+similar to object model, single inheritance with properties.
 
 \subsection{Enumeration Type Declarations}
-\label{ifcOWL:enum-types}
+\label{subsec:ifcOWL-enum-types}
 

### 2015-04-29 article-00-headers.tex
Changed rules for simple datatypes
Wrote about defined, entity types. Removed listing captions to save space.

@@
 \usepackage{color}
+\usepackage{linegoal}
 
@@
 
-\lstset{language=TeX,
+% language Ruby is used to enable comment lines started with hash character
+
+\lstset{language=Ruby,
+aboveskip=6pt,
+belowskip=6pt,
 backgroundcolor=\color[gray]{0.9},
@@
 
+%\usepackage{calc}
+
+\graphicspath{{images/}}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %\newcommand{\name}[1]{\emph{#1}}
+
+% programming language names
 \newcommand{\name}[1]{\textsf{\footnotesize{}#1}}
-\newcommand{\TODO}[1]{\textbf{TODO: }\textit{#1}}
-\newcommand{\REVIEW}[1]{\textit{#1}}
+
+% highlight texts
+\newcommand{\highlight}[1]{\colorbox{yellow}{\parbox[t]{\linegoal}{#1}}}
+
+% TODO text
+\newcommand{\TODO}[1]{\highlight{\textbf{TODO: }{#1}}}
+
+%\newcommand{\REVIEW}[1]{\textit{#1}}
+
+% ifcOWL layer names
+\newcommand{\lite}{\textbf{Lite}}
+\newcommand{\standard}{\textbf{Standard}}
+\newcommand{\advanced}{\textbf{Advanced}}
+
+\newcommand{\ifclite}{\textbf{ifcOWL-Lite}}
+\newcommand{\ifcstandard}{\textbf{ifcOWL-Standard}}
+\newcommand{\ifcadvanced}{\textbf{ifcOWL-Advanced}}
 
@@
 \begin{document}
+
 %
@@
 The volume of data produced in the construction industry is growing all the time.
-However, building data is currently much less accessible than other smart city data related to geography, traffic, or services.
+However, building data is currently difficult to access in an open manner.
 We present an ontology derived from the conceptualization of IFC, the existing standard for representing building designs, and a tool to convert building designs from IFC files to RDF datasets that conform to the ontology.

### 2015-04-29 article-01-introduction.tex
Wrote about defined, entity types. Removed listing captions to save space.

@@
 Construction industry is becoming increasingly data intensive with the growing use of
-information and communication technology in all phases of a building life-cycle. The design phase is
-carried out with Building Information Modeling (BIM) tools and systems that produce elaborate
-object-oriented 3D models, and the construction phase with a wide range of construction planning,
-scheduling, and coordination systems. In the operational stage the building is monitored,
-controlled, and maintained based on sensor data and human observations.
+information and communication technology in all phases of a building life-cycle. 
+In the future, buildings will act as hubs of a wide variety of real-time and accumulated data about
+spaces, indoor locations, indoor routes, assets, people, sensors, materials, equipment, connections
+to surrounding networks (traffic, infrastructure, communication), detected issues, and activities in the building.
 
-In the future, buildings will act as hubs of a wide variety of real-time and historic data about
-spaces, indoor locations, indoor routes, assets, people, sensors, materials, equipment, connections
-to surrounding networks (traffic, infrastructure, communication), and so on.
-
-A huge amount of design information is already created in the building design phase 
-using BIM tools and there is a widely adopted standard, IFC (Industry Foundation Classes) to
-represent building designs. IFC has been specified to improve interoperability between different
-disciplines in the design phase of a project, including the architecture, structural engineering,
-MEP engineering models, and so on. Nowadays, the IFC standard is supported by all the major BIM
+Currently, a huge amount of basic design information about buildings are produced  
+using modern Building Information Modeling (BIM) tools. There is a widely adopted standard, 
+IFC (Industry Foundation Classes) that provides a common representation of building designs developed in different design disciplines: architecture, structural engineering,
+MEP engineering, and so on. Nowadays, the IFC standard is supported by all major BIM
 vendors; their BIM tools can export building models as IFC files that can be exchanged with other
@@
 
-The design models can provide the framework with which the information gathered from other sources
-can be organized and visualized.  However, despite the IFC standard, the design information is currently
-difficult to access and utilize. The sharing of building models based on file exchange does not work
-sufficiently well in practice. Utilization of related information from different models requires
-repeated manual work from the designers. In the longer run, file exchange cannot provide a fruitful
-basis for open utilization of building information and emergence of new, innovative applications.
+The BIM models could provide a natural framework to organize and manage information gathered from other sources. However, a big obstacle to that is the difficulty to access and utilize BIM models. 
+In the realm of IFC the sharing of building models is based on point-to-point file exchange which 
+requires repeated manual work from designers.
+Information cannot be used in an online, granular, and interlinked manner, which provides a poor basis for open utilization of building information and the emergence of new, innovative applications.
 
-We have developed a WebIFC converter to derive an OWL ontology from the IFC schema (called ifcOWL),
-and RDF datasets from IFC files (called ifcRDF).
+To make building information more readily accessible, we have developed a \emph{WebIFC converter} to derive an OWL ontology from the IFC schema (\emph{ifcOWL}), and RDF datasets from IFC files (\emph{ifcRDF}).
+The ifcOWL ontology is used to classify concepts, characterize possible relationships, and define constraints in ifcRDF datasets which replicate the content of original IFC data files.
+All terms defined in the ifcOWL ontology must therefore be based on the corresponding constructs in the original IFC schema.
+The conversion from IFC data files to RDF is straightforward and can be done without information loss.
 
-The ifcOWL ontology is used to classify concepts, characterize possible relationships, and define possible constraints of data in ifcRDF datasets which replicate the content of original IFC-STEP data files.
-Thus, all terms defined in the ifcOWL ontology must be based on the corresponding constructs of the original IFC-EXPRESS schema.
-Ideally, there should be only one ifcOWL ontology for one IFC schema (see Fig.
-\ref{fig:ifcOWL-single}).
-However, taking account that ifcRDF datasets are automatically generated, data validation is almost never required.
-So in most of cases it is wisely to use a lightweight ontology instead of a detailed one in order to facilitate more effective reasoning.
+Ideally, there should also be just one ifcOWL ontology for each version of the IFC schema (Fig.
+\ref{fig:ifcOWL-layers}, a).
+However, the conversion from the EXPRESS data definition language (that is used to specify the IFC schema) to OWL can be done in several different ways for the following reasons.
+Firstly, there is a \emph{mismatch} between the modeling constructs between EXPRESS and OWL.
+Secondly, for most practical purposes, the IFC data can be considered \emph{valid and consistent} with respect to the schema, since the IFC models are exported from BIM tools with certified converters.
+Thirdly, as the master data of the design models will remain in the native formats of the BIM tools and evolves over time, the icfRDF can in practice only be used in a \emph{read-only} fashion.
+Fourthly, the conversion of all possible IFC schema information will lead into a \emph{extensive} OWL ontology that mostly contains information not needed in practical applications or use cases.
+
+Consequently, we have converted the IFC schema into a \emph{flexible, multilayer ifcOWL ontology}.
+Different representational constructs are introduced at different levels.
+The ifcRDF is compatible with any of the layers; different layers just provide different amounts of type information that can be used with RDF.
 
@@
 \centering
-\includegraphics[width=0.75\textwidth]{images/ifcOWL-single.png}
-\caption{A single ifcOWL ontology}
-\label{fig:ifcOWL-single}
+\includegraphics{images/ifcOWL-approaches.jpg}
+\caption{Two approaches of creating and using ifcOWL ontology}
+\label{fig:ifcOWL-layers}
 \end{figure}
 
-Various levels of details of ifcOWL ontologies are defined in the Table \ref{tab:ifcOWL-levels}.
-An ifcOWL ontology version can be considered as extension of all ontology versions with smaller level numbers.
+Different levels of details of the ifcOWL ontology are the following:
 
-\begin{table}[h]
-\centering
-\footnotesize
-\begin{tabularx}{\textwidth}{|X|X|X|}
-\hline
-\textbf{Level (name)} & \textbf{Detail} & \textbf{Note}   \\
+\begin{enumerate}
+\item
+    \textbf{ifcOWL-Lite} (compatible with OWL 2 EL, OWL 2 QL, OWL 2 RL)
+    \begin{itemize}
+        \item Schema meta data
+        \item All type definitions and hierarchies
+        \item All entity type properties (names and range types)
+    \end{itemize}
+\item
+    \textbf{ifcOWL-Standard} (compatible with OWL 2 RL)
+    \begin{itemize}
+        \item Schema meta data
+        \item All entity type keys
+        \item All entity type inverse properties
+    \end{itemize}
+\item
+    \textbf{ifcOWL-Advanced} (compatible with OWL 2 RL)
+    \begin{itemize}
+        \item All properties' cardinality constraints
+        \item All list types' cardinality constraints
+    \end{itemize}
+\end{enumerate}
 
-\hline
-0 (None) & Schema meta data &   \\
+Since all ifcRDF datasets are compatible with all layers of the ifcOWL ontology (Fig.
+\ref{fig:ifcOWL-layers}, b), a user can choose the best one based on the requirements of the concrete use case and tools.
+For instance, many Linked Data applications that need to query entity properties, extract the type hierarchy, or convert ifcRDF datasets back to object-oriented data models it is sufficient to use the \textbf{ifcOWL-Lite} ontology.
+In case of reasoning about data that includes inverse properties, it is recommended to use \textbf{ifcOWL-Standard} ontology.
+If an RDF dataset needs to be modified and data validated using type and cardinality constraints, \textbf{ifcOWL-Advanced} ontology is the right choice.
 
-\hline
-1 & All type names &   \\
-
-\hline
-2 & Entity type hierarchy &   \\
-
-\hline
-3 & Enumeration \& Select data types & \\
-
-\hline
-4 (Lite) & Property names and value types &   \\
-
-\hline
-5 & Keys &   \\
-
-\hline
-6 (Standard) & Inverse properties &   \\
-
-\hline
-7 & Property cardinality restrictions  &   \\
-
-\hline
-8 (Advanced) & List type cardinality restrictions  &   \\
-
-\hline
-9 (Full) & Property value types with exact value ranges  & Not supported by the converter  \\
-
-\hline
-\end{tabularx}
-\caption{Levels of details of ifcOWL ontology}
-\label{tab:ifcOWL-levels}
-\end{table}
-
-Having multiple versions of the same ontology and knowing exactly that every ifcRDF dataset will be compatible with any related ifcOWL ontology (see Fig.
-\ref{fig:ifcOWL-multilayered}) user can choose the best one based on the requirements of the concrete use case and tools used.
-For instance, when it is required to query entity properties, extract the type hierarchy, or convert ifcRDF datasets back to object-oriented data models it is enough to use \textbf{ifcOWL-Lite} ontology.
-It case of reasoning about data including inverse properties, it is recommended to use \textbf{ifcOWL-Standard} ontology.
-When data validation, for example, checking type and cardinality constraints is needed \textbf{ifcOWL-Advanced} ontology is the most reasonable choice.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.75\textwidth]{images/ifcOWL-multilayer.png}
-\caption{A multilayer ifcOWL ontology}
-\label{fig:ifcOWL-multilayered}
-\end{figure}
+% \begin{figure}
+% \centering
+% %\def\svgwidth{\textwidth}
+% \def\svgscale{1}
+% \input{images/image.pdf_tex}
+% %{\includesvg{images/test-2}}
+% %\includegraphics[width=0.75\textwidth]{images/ifcOWL-multilayer-2.png}
+% \caption{A multilayer ifcOWL ontology}
+% \label{fig:ifcOWL-multilayered}
+% \end{figure}

### 2015-04-29 article-02-main.tex
Wrote about defined, entity types. Removed listing captions to save space.

--- 2015-04-28 article-02-main.tex
@@
 
+\begin{ontologyRule}[OWL profile]
+\ifclite{} layer must be compatible with any OWL 2 DL profile: OWL 2 EL, OWL 2 QL or OWL 2 RL.
+\ifcstandard{}, \ifcadvanced{} must be compatible with OWL 2 RL\footnote{Our WebIFC converter tool allows forced choice of other OWL profiles for any layer: OWL Lite, OWL DL, OWL Full, and OWL 2 Full}.
+\end{ontologyRule}
+
 \TODO{Explain: why an OWL 2 profile has to be being chosen, how that choice will impact to the reasoning and querying performance, whether we can use other OWL 2 EL or QL for ifcOWL-Lite ontology}
 
-\begin{ontologyRule}[OWL profile, ifcOWL layers: all]
-All ifc\-OWL ontologies must be based on OWL 2 RL profile\footnote{Our IFC-EXPRESS-to-OWL converter tool which is available on http://drumbeat.cs.hut.fi also allows choosing other OWL profiles: OWL Lite, OWL DL, OWL Full, OWL 2 EL, OWL 2 QL, OWL 2 Full}.
-\end{ontologyRule}
+OWL 2 RL has been chosen for \ifcstandard{}, \ifcadvanced{} because in contrast to OWL 2 EL and OWL 2 QL, it supports such needed things as: union data\-types (see \ref{subsec:ifcOWL-select-types}), enumerations of individuals (see \ref{subsec:ifcOWL-enum-types}), anonymous individuals (see \ref{subsec:ifcRDF-naming-individuals}), etc.\cite{w3c:owl2-profiles}.
 
-This profile has been chosen because in contrast to OWL 2 EL and OWL 2 QL, it supports such needed things as: enumerations of individuals (see \ref{ifcOWL:enum-types}), named individuals (see \ref{ifcRDF:naming-individuals}), etc.\cite{w3c:owl2-profiles}.
-
-\begin{ontologyRule}[Domain and range information, ifcOWL layers: all]
+\begin{ontologyRule}[Domain and range information]
 Property domain and range constraints specified by \name{rdfs:domain} and \name{rdfs:range} should not be included in ifcOWL ontologies.
@@
 
-\begin{ontologyRule}[Namespaces, ifcOWL layers: all]
-All constructs related purely to EXPRESS, STEP or IFC specifications must be belong to namespaces with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly.
-At this moment the namespaces are defined as in Listing \ref{lst:ifcOWL-namespaces}.
-Instead of \name{IFCXXX} there should be IFC versions, e.g.
-\name{IFC2x3} or \name{IFC4}.
-\end{ontologyRule}
+\begin{ontologyRule}[Namespaces]
+All constructs related to EXPRESS, STEP or IFC specifications must be belong to namespaces with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly.
+At this moment the namespaces are defined as below (istead of \name{IFCXXX} there should be IFC versions, e.g.
+\name{IFC2x3} or \name{IFC4}):
 
-\begin{lstlisting}[caption={Namespace definitions},label=lst:ifcOWL-namespaces]
+%\begin{lstlisting}[caption={Namespace definitions},label=lst:ifcOWL-namespaces]
+\begin{lstlisting}
 @prefix expr: <http://drumbeat.cs.hut.fi/owl/EXPRESS#>
@@
 \end{lstlisting}
+\end{ontologyRule}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Simple Datatype Declarations
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \subsection{Simple Datatype Declarations}
@@
 Simple data\-types built in EXPRESS are: \name{STRING}, \name{BI\-NA\-RY}, \name{IN\-TE\-GER}, \name{REAL}, \name{NUM\-BER}, \name{BOO\-LEAN} and \name{LO\-GI\-CAL}.
-However, type \name{LOGICAL} which has three possible values \name{TRUE}, \name{FALSE} and \name{UNKNOWN} is considered by us as an enumeration data\-type (see \ref{ifcOWL:enum-types}).
+However, type \name{LOGICAL} which has three possible values \name{TRUE}, \name{FALSE} and \name{UNKNOWN} is considered by us as an enumeration data\-type (see \ref{subsec:ifcOWL-enum-types}).
 The same to type \name{BOOLEAN} which is like a primitive boolean data\-type but semantically close and met with the same frequency in IFC4 as \name{LOGICAL}.
 
-On the other hand, there are two defined data\-types in IFC (not in EXPRESS) that also can be referred to the simple-data-type group: \name{Ifc\-Time\-Stamp} which is a so-called TimeSpan data\-type and can be easily converted to \name{xsd:date\-Time}, and \name{Ifc\-Globally\-Unique\-Id} which represents set of compressed Globally Unique Identifiers (GUIDs), i.e.
-strings with fixed length (see Listing \ref{lst:ifcOWL-additional-simple-types}).
+\begin{ontologyRule}[Simple data\-types]
+All simple data\-types except \emph{BOOLEAN} and \emph{LOGICAL} are declared as the same as their most similar primitive XSD data\-types which are supported by OWL 2 \cite{w3c:owl2-profiles} and are among preferred standard data\-types in Linked Data Platform \cite{w3c:ldp-best-practices}:
 
-\begin{lstlisting}[caption={Two defined data\-types considered as simple data\-types}, label=lst:ifcOWL-additional-simple-types]
-(* number of seconds elapsed since the beginning of the year 1970 *)
-TYPE IfcTimeStamp = INTEGER; END_TYPE;
-
-(* compressed GUID *)
-TYPE IfcGloballyUniqueId = STRING(22) FIXED; END_TYPE;
-\end{lstlisting}
-
-\begin{ontologyRule}[Simple data\-types, ifcOWL layers: all]
-All simple data\-types except \emph{BOOLEAN} and \emph{LOGICAL} are the same as their most similar primitive XSD data\-types which are supported by OWL 2 \cite{w3c:owl2-profiles} and are among preferred standard data\-types in Linked Data Platform \cite{w3c:ldp-best-practices} (see Listing \ref{lst:ifcOWL-simple-types}).
-However, it is recommended to use directly XSD s like (e.g.
-\name{xsd:string}) instead of the EXPRESS data\-type constructs (e.g.
-\name{expr:STRING}).
-\end{ontologyRule}
-
-% For \name{STRING} and \name{Ifc\-Globally\-UniqueId} it is \name{xsd:\-string}, for \name{BINARY} -- \name{xsd:\-hexBinary}, for \name{INTEGER} -- \name{xsd:\-integer}, for \name{REAL} and \name{NUMBER} -- \name{xsd:\-double}, for \name{Ifc\-Time\-Stamp} -- \name{xsd:\-dateTime}.
-
-\begin{lstlisting}[caption={Simple data types}, label=lst:ifcOWL-simple-types]
+% \begin{lstlisting}[caption={Simple data types}, label=lst:ifcOWL-simple-types]
+\begin{lstlisting}
 expr:STRING, ifc:IfcGlobalUniqueId owl:sameAs xsd:string .
@@
 expr:REAL, expr:NUMBER owl:sameAs xsd:double .
-ifc:IfcTimeStamp owl:sameAs xsd:dateTime .
 \end{lstlisting}
+\end{ontologyRule}
 
@@
 For instance, type \name{xsd:double} is more similar to \name{REAL} than \name{xsd:decimal} as it is an IEEE 64-bit floating-point data\-type and supports scientific notation \cite{w3c:xsd,datapic:xsd}.
+It is recommended to use XSD data\-type names directly when they are needed.
 
 \begin{ontologyRule}[Simple data\-types with sizes, ifcOWL layers: all]If \name{STRING} and \name{BINARY} data\-types are specified with a size parameter, i.e.
-\name{STRING (max\-Length\-In\-Bytes)} or \name{BINARY(size\-In\-Bits)}, then they will be named in format \name{\textless type\_name \textgreater\textless size\textgreater} (e.g.
-\name{STRING22}).
-In layers \lite{} and \standard{} they are equivalent to data\-types without size.
-But in layer \advanced{}, in order to specify the size parameter, additional constraints must be added by using property \name{owl:onDatatype} and data\-type facets \name{xsd:maxLength} or \name{xsd:length} \cite{w3c:owl2-syntax,w3c:xmlschema11-2} (see example in Listing \ref{lst:ifcOWL-simple-types-with-size}).
-\end{ontologyRule}
+\name{STRING (max\-Length\-In\-Bytes)} or \name{BINARY(size\-In\-Bits)}, then in layers \lite{} and \standard{} they are defined as equivalent to data\-types without size.
+But in layer \advanced{}, the size parameter must be declared by using property \name{owl:onDatatype} and data\-type facets \name{xsd:maxLength} or \name{xsd:length} \cite{w3c:owl2-syntax,w3c:xmlschema11-2}, for instance:
 
-\begin{lstlisting}[caption={Size constraint of simple s},
-label=lst:ifcOWL-simple-types-with-size]
+% \begin{lstlisting}[caption={Size constraint of simple datatypes},label=lst:ifcOWL-simple-types-with-size]
+\begin{lstlisting}
 # only for Lite, Standard layers
@@
 expr:STRING22 owl:equivalentClass
-    [   rdf:type        rdfs:Datatype ;
+    [   a rdfs:Datatype ;
         owl:onDatatype  xsd:string ;
@@
 \end{lstlisting}
+\end{ontologyRule}
 
-% \begin{lstlisting}[caption={},label=1]
-% expr:STRING rdf:type owl:Class;
-%     rdfs:subClassOf [
-%         owl:Restriction;
-%         owl:onProperty expr:hasValue;
-%         owl:allValuesFrom xsd:string;
-%     ].
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Enumeration Datatype Declarations
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% \begin{lstlisting}[caption={Simple data\-types as instances of \name{owl:Class}}, label=lst:ifcOWL-simple-types]
-% expr:STRING a owl:Class;
-%     rdfs:subClassOf [ a owl:Restriction;
-%                       owl:onProperty rdf:value;
-%                       owl:allValuesFrom xsd:double; ] .
-% ifc:IfcTimeStamp a owl:Class;
-%     rdfs:subClassOf [ a owl:Restriction;
-%                       owl:onProperty rdf:value;
-%                       owl:allValuesFrom xsd:dateTime; ]
-% \end{lstlisting}
+\subsection{Enumeration Datatype Declarations}
+\label{subsec:ifcOWL-enum-types}
 
-% The reason of using class boxing technique is to make simple data\-types be instances of \name{owl:Class}.
-Instead of that they could be either simply replaced with primitive XSD data\-types, or defined as shown in the Listing , but in these cases they will be instances of \name{rdfs:Datatype}.
-and cannot be used for union classes
+An enumeration data\-type represents a set of possible fixed values.
 
-% \begin{lstlisting}[caption={Simple data\-types as instances of \name{rdfs:Datattype}},label=lst:ifcOWL-simple-types-2]
-% # if LEVEL_OF_DETAIL is higher than STANDARD
-% expr:STRING owl:equivalentClass [
-%     rdf:type        rdfs:Datatype ;
-%     owl:onDatatype  xsd:string ;
-%     owl:withRestrictions ()
-% ]
-% \end{lstlisting}
+\begin{ontologyRule}[Enumeration data\-types]
+Each enumeration data\-type in \textbf{ifc\-OWL} is defined as an \name{owl:Class} which is a subclass of \name{expr:\-Enum\-Class}.
+All possible values of enumeration data\-types are defined as named individuals.
+Enumeration data\-type memberships are defined by means of \name{rdf:type} in the \lite{} layer, and by means of \name{owl:oneOf} in \standard{} and \advanced{} layers as shown below:
+% \begin{lstlisting}[caption={Enumeration data\-types}, label=lst:ifcOWL-enum-types]
+\begin{lstlisting}
+# only for Lite layer
+ifc:IfcAssemblyPlaceEnum a owl:Class ; rdfs:subclassOf expr:EnumClass .
+ifc:SITE, ifc:FACTORY, ifc:NOTDEFINED rdf:type ifc:IfcAddressTypeEnum .
 
-% Other considered conversion options were:
-% \begin{itemize}
-% \item To replace simple data\-types with the primitive XSD data\-types.
-But in this case all select data\-types (see \ref{subsec:ifcOWL:select-types}) cannot be defined by means of property \name{owl:unionOf} the range of which
+# only for Standard, Advanced layers
+ifc:IfcAssemblyPlaceEnum a owl:Class ;  rdfs:subclassOf expr:EnumClass ;
+    owl:oneOf ( ifc:SITE ifc:FACTORY ifc:NOTDEFINED ) .
+\end{lstlisting}
+\end{ontologyRule}
 
-% all properties in any OWL profile except OWL Full must be either \name{owl:}
-% \end{itemize}
+The reason of two different approaches is that enumeration with more than one individual or literal are not supported in OWL 2 EL or QL.
+In both cases enumeration values are defined as individuals of one class so in terms of reasnoning logic they are equivalent.
+The enumeration values are defined as individuals because OWL 2 RL supports \name{Object\-One\-Of} to define enumeration of individuals, but not \name{Data\-One\-Of} \cite{w3c:owl2-profiles}.
+The fact that some values such as \name{ifc:NOT\-DE\-FI\-NED} belong to different enumeration data\-types should not cause any problems since all enumeration values in ifc\-RDF datasets are defined together with a concrete enumeration data\-type (see \ref{subsec:ifcRDF-enum-values}).
+It is also very easy to get all possible values of an enumeration data\-type by using simple SPARQL queries.
 
-% ifc:ifcLabel rdf:type owl:Class;
-%     rdfs:subClassOf expr:STRING.
-% \end{lstlisting}
+As mentioned above, simple data\-types \name{BOOLEAN} and \name{LOGICAL} are considered by us as enumeration data\-types.
+But these types and their values must be defined in the EXPRESS namespace, e.g.
+\name{expr:LOGICAL}, \name{expr:TRUE}, \name{expr:FALSE} or \name{expr:UNKNOWN}.
 
-% \begin{lstlisting}[caption={},label=1]
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Select Datatype Declarations
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% % http://answers.semanticweb.com/questions/229/how-to-define-a-own-notation-data\-type
+\subsection{Select Datatype Declarations}
+\label{subsec:ifcOWL-select-types}
 
-% http://semantic-web-grundlagen.de/wiki/Guide_to_OWL_2_for_OWL_1_users#Datatypes
+A select data\-type represents an union of other data\-types.
+Union members of a select data\-type usually are either entity data\-types, or defined data\-types all together.
+But there are a few exceptions, for instance, type \name{Ifc\-Trim\-ming\-Se\-lect} is an union of entity data\-type \name{Ifc\-Car\-te\-sian\-Point} and defined data\-type \name{Ifc\-Pa\-ra\-me\-ter\-Va\-lue}.
 
-% http://www2.informatik.hu-berlin.de/~wandelt/SW201213/9OWL2.pdf
+\begin{ontologyRule}[Select data\-types]
+A select data\-type must be converted to an \name{owl:Class} which is a subclass of \name{expr:\-Select\-Class}.
+Union-member data\-types are defined by means of \name{rdfs:\-subclassOf} in the \lite{} layer, and by means of \name{owl:unionOf} in \standard{} and \advanced{} layers as shown in the listing below.
+\end{ontologyRule}
 
-% Using that in OWL 2 DL is no problem.
-If you want to stay within one of the profiles though note that defining data\-types through such restrictions isn't allowed in any of them.
-So for increased interoperability (also with OWL 1 and RDFS) you could just use xsd:string instead.
-That should be perfectly fine with SKOS.
+%\begin{lstlisting}[caption={Select data\-types}, label=lst:ifcOWL-select-types]
+\begin{lstlisting}
+# only for Lite layer
+ifc:IfcTrimmingSelect a owl:Class ; rdfs:subclassOf expr:SelectClass .
+ifc:IfcCartesianPoint, ifc:IfcParameterValue
+    rdfs:subclassOf ifc:IfcTrimmingSelect .
 
-% expr:STRING rdf:type rdfs:Datatype;
-%  owl:onDatatype xsd:string;
-%  owl:withRestrictions ().
+# only for Standard, Advanced layers
+ifc:IfcTrimmingSelect a owl:Class ; rdfs:subclassOf expr:SelectClass ;
+    owl:unionOf ( ifc:IfcCartesianPoint ifc:IfcParameterValue ) ] .
+\end{lstlisting}
 
-% ifc:ifcLabel rdf:type owl:Class;
-%     rdfs:subClassOf expr:STRING.
-% \end{lstlisting}
+Similarly to the above case, two different approaches are used because property \name{owl:unionOf} supported only in OWL 2 RL, but not in OWL 2 EL and OWL 2 QL \cite{w3c:owl2-profiles}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Defined Datatype Declarations
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-% \item \textbf{Defined or declared data type}: This further specializes other data\-types — e.g., define a datatype \emph{PositiveInteger} that is of type integer with a positive value.
+A defined (declared, named) data\-type is usually specified as equivalent to a simple data\-type or another defined data\-type, with or without a additional constraint.
+For instance, type \name{Ifc\-Length\-Mea\-sure} is defined as equal to \name{REAL}, and type \name{Ifc\-Po\-si\-ti\-ve\-Length\-Mea\-su\-re} is a set of positive \name{Ifc\-Length\-Mea\-sure} values.
+
+\begin{ontologyRule}[Defined data\-types]
+If the defined data\-type is based on a simple data\-type then in ifcOWL it is declared as a \name{owl:Class} which is a subclass of \name{expr:\-Defined\-Class} has a single property \name{rdf:value} with that simple data\-type.
+If the defined data\-type is based on another defined data\-type then it is declared as a subclass of the \name{owl:Class} of that defined data\-type, for instance:
+
+%\begin{lstlisting}[caption={Defined data\-types}, label=lst:ifcOWL-defined-types]
+\begin{lstlisting}[aboveskip=3pt]
+ifc:IfcLengthMeasure a owl:Class ; rdfs:subClassOf expr:DefinedClass ;
+    rdfs:subClassOf [ a owl:Restriction; owl:onProperty rdf:value;
+                      owl:allValuesFrom xsd:double ] .
+
+ifc:IfcPositiveLengthMeasure a owl:Class ;
+    rdfs:subClassOf ifc:IfcLengthMeasure .
+\end{lstlisting}
+\end{ontologyRule}
+
+Defined data\-types are not declared as a \name{rdfs:Datatype}, but \name{owl:Class}es which wrap \name{rdfs:Datatype} values because in contract to simple data types they can be used for declaration of select data\-types (see \ref{subsec:ifcOWL-select-types}).
+
+Defined data\-type \name{Ifc\-Time\-Stamp} is declared as equivalent to \name{INTEGER} but represents so-called Unix time value that can be easily converted to \name{xsd:date\-Time}.
+So it wraps a \name{xsd:date\-Time} value instead of \name{xsd:integer}.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Entity Datatype Declarations
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \subsection{Entity Datatype Declarations}
+\label{subsec:ifcOWL-entity-types}
 
-\subsection{Enumeration Type Declarations}
-\label{ifcOWL:enum-types}
+Entity data\-types are identical to classes in object-oriented programming languages.
+Each type may have not more than one super\-type that it inherits attributes from.
+An attribute is defined with a data\-type, a minimum and a maximum cardinalities, can be marked as optional.
+
+\begin{ontologyRule}[Entity data\-types]An entity data\-type is converted to an \name{owl:\-Class} which is subclass of \name{expr:\-Entity\-Class} or its superclass.
+Entity attributes are defined by means of property \name{owl:on\-Pro\-per\-ty}.
+If an attribute is optional then its minimum qualified cardinality is 0, otherwise -- 1.
+Maximum qualified of all attributes are always 1.
+In case if an attribute is a list or a set of some data\-type then its type is considered as an external aggregated data\-type, for example:
+%\begin{lstlisting}[caption={Entity data\-types}, label=lst:ifcOWL-defined-types]
+\begin{lstlisting}
+ifc:IfcProperty a owl:Class ; rdfs:subClassOf expr:EntityClass .
+
+ifc:IfcComplexProperty a owl:Class ; rdfs:subClassOf ifc:IfcProperty ;
+    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:coordinates;
+                      owl:allValuesFrom LIST_1_3_IfcLengthMeasure ] ;
+    # only for Advanced layer
+    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:coordinates;
+                      owl:cardinality "1"^^xsd:integer ] .
+
+# only for Standard & Advanced layers
+[] a owl:AllDisjointClasses ;
+    owl:members ( ifc:IfcComplexProperty ifc:IfcSimpleProperty ) .
+\end{lstlisting}
+\end{ontologyRule}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Aggregated Datatype Declarations
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Aggregated Type Declarations}
+\label{subsec:ifcOWL-aggregated-types}
 

### 2015-04-29 article-02-main.tex
Updates from ShareLaTeX

@@
 
-OWL 2 RL has been chosen for \ifcstandard{}, \ifcadvanced{} because in contrast to OWL 2 EL and OWL 2 QL, it supports such needed things as: union data\-types (see \ref{subsec:ifcOWL-select-types}), enumerations of individuals (see \ref{subsec:ifcOWL-enum-types}), anonymous individuals (see \ref{subsec:ifcRDF-naming-individuals}), etc.\cite{w3c:owl2-profiles}.
+OWL 2 RL has been chosen for \ifcstandard{}, \ifcadvanced{} because in unlike OWL 2 EL and OWL 2 QL, it supports such needed things as: union data\-types (see \ref{subsec:ifcOWL-select-types}), enumerations of individuals (see \ref{subsec:ifcOWL-enum-types}), anonymous individuals (see \ref{subsec:ifcRDF-naming-individuals}), etc.\cite{w3c:owl2-profiles}.
 
@@
     rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:coordinates;
-                      owl:allValuesFrom LIST_1_3_IfcLengthMeasure ] ;
+                      owl:allValuesFrom ifc:LIST_1_3_IfcLengthMeasure ] ;
     # only for Advanced layer
@@
 
+Aggregation data\-types may be unordered (\name{SET} and \name{BAG}) or  ordered (\name{LIST} and \name{ARRAY}).
+\name{BAG} may contain a particular value more than once, unlike \name{SET}.
+\name{ARRAY} is specified with starting and ending indices while other types are specified with min and max cardinalities.
+
+In order to declare these types, new classes were created in a similar way as Ordered List Ontoloty \cite{olo:ontology}:
+\begin{itemize}
+\item Class \name{expr:\-Collec\-tion\-Class} has four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index} and a nonfunctional property \name{expr:slot} with type \name{expr:Slot}.
+\item Class \name{expr:Slot} has two functional properties \name{expr:item} (the actual collection item value), \name{expr:index} and also optional functional properties \name{expr:previous}, \name{expr:next}.
+\item Classes \name{expr:SetClass}, \name{expr:ListClass}, \name{expr:ArrayClass} and \name{expr:BagClass} are subclasses of \name{expr:\-Collec\-tion\-Class}.
+\end{itemize}
+
+Below is a fragment of definition of type \name{SET [2:?] OF IfcProfileDef}:
+
+\begin{lstlisting}
+# all layers:
+ ifc:SET_2_UNDEFINED_OF_IfcProfileDef a owl:Class ;
+    rdfs:subClassOf ifc:SET_2_UNDEFINED , ifc:SET_OF_IfcProfileDef .
+
+ ifc:SET_2_UNDEFINED, SET_OF_IfcProfileDef a owl:Class ;
+    rdfs:subClassOf expr:SetClass .
+
+# only for layer Advanced:
+ ifc:SET_2_UNDEFINED rdfs:subClassOf rdfs:subClassOf [ a owl:Restriction ;
+    owl:onProperty expr:slot ; owl:minCardinality "2"^^xsd:integer ] .
+
+ ifc:SET_OF_IfcProfileDef rdfs:subClassOf [ a owl:Restriction ;
+    owl:onProperty expr:slot ; owl:allValuesFrom ifc:SLOT_OF_IfcProfileDef ].
+
+ ifc:SLOT_OF_IfcProfileDef rdfs:subClassOf [ a owl:Restriction ; owl:onProperty expr:item ; owl:allValuesFrom ifc:IfcProfileDef ] .
+\end{lstlisting}
+
 \section{Generating ifcRDF Datasets From IFC Data}
 
-\subsection{Naming individuals}
-\label{ifcRDF:naming-individuals}
+\begin{lstlisting}
+:GUID_1916794F-5605-4499-9AB0-F155DE8D3B6C
+    a               ifc:IfcPropertySet ; # entity type
+    ifc:globalId    [ a ifc:IfcGloballyUniqueId ; # defined/select type
+                     rdf:value "0P5dbFLWL4cPgmyLNUZJji"^^xsd:string ] ;
+    ifc:hasProperties  _:b11 .
+
+_:b11 a ifc:SET_1_UNLIMITED_OF_IfcProperty ; # entity type
+    expr:size "3"^^xsd:integer ;
+    expr:itemType ifc:IfcProperty ;
+    expr:slot _:b12, _:b13, _:b14 .
+
+_:b12 a ifc:SLOT_OF_IfcProperty ;
+    expr:item [ a ifc:IfcPropertySingleValue ; # entity type
+                ifc:name "initial_GUID"^^xsd:string ;
+                ifc:nominalValue  _:b15 ] .
+\end{lstlisting}
+
+\subsection{Metadata}
+
+\TODO {Recommendations about model namespaces, naming individuals, defined types of select types }

### 2015-04-29 article-00-headers.tex
Wrote sth about ifcRDF

@@
 
-\spnewtheorem{ontologyRule}{Ontology Rule}{\bfseries}{\rmfamily}
+\spnewtheorem{principle}{Principle}{\bfseries}{\rmfamily}
 
-\spnewtheorem{datasetsRule}{Datasets Rule}{\bfseries}{\rmfamily}
+%\spnewtheorem{datasetsRule}{Datasets Rule}{\bfseries}{\rmfamily}
 
@@
 % ifcOWL layer names
-\newcommand{\lite}{\textbf{Lite}}
+\newcommand{\simple}{\textbf{Simple}}
 \newcommand{\standard}{\textbf{Standard}}
-\newcommand{\advanced}{\textbf{Advanced}}
+\newcommand{\extended}{\textbf{Extended}}
 
-\newcommand{\ifclite}{\textbf{ifcOWL-Lite}}
-\newcommand{\ifcstandard}{\textbf{ifcOWL-Standard}}
-\newcommand{\ifcadvanced}{\textbf{ifcOWL-Advanced}}
+\newcommand{\ifcowl}{\textbf{ifcOWL}}
+\newcommand{\ifcsimple}{\ifcowl{}-\simple{}}
+\newcommand{\ifcstandard}{\ifcowl{}-\standard{}}
+\newcommand{\ifcextended}{\ifcowl{}-\extended{}}
 
@@
 %
-\title{Publishing Datasets about Building Designs \\Based on a Multilayer Building Ontology}
+\title{Generating Datasets about Building Designs \\Based on a Multilayer Building Ontology}
 %
@@
 
-The volume of data produced in the construction industry is growing all the time.
-However, building data is currently difficult to access in an open manner.
+The volume of data produced in the construction industry is rapidly growing but, as of now, it is not available for open and flexible access and utilization.
 We present an ontology derived from the conceptualization of IFC, the existing standard for representing building designs, and a tool to convert building designs from IFC files to RDF datasets that conform to the ontology.
 Replacing the traditional point-to-point exchange of IFC files with publication of RDF datasets  by using Linked Data principles and tools enables information integration across different IFC-based and other smart city datasets, and the granular access to individual components.
-Moreover, it paves the way to a completely distributed and loosely coupled architecture for building information management.
-Finally, opening selected parts of construction data to a wider set of stakeholders creates opportunities for new and innovative applications in the smart cities of the future.
+It paves the way to a distributed and loosely coupled architecture for building information management.
+Opening selected parts of construction data to a wider set of stakeholders creates opportunities for innovative applications in the smart cities of the future.
 
-\keywords{IFC, BIM, Ontology, Linked Data, OWL, RDF}
+\keywords{Ontology, Linked Data, OWL, RDF, IFC, BIM}
 

### 2015-04-29 article-01-introduction.tex
Wrote sth about ifcRDF

@@
 spaces, indoor locations, indoor routes, assets, people, sensors, materials, equipment, connections
-to surrounding networks (traffic, infrastructure, communication), detected issues, and activities in the building.
+to surrounding networks (traffic, infrastructure, communication), detected issues, and 
+activities in the building.
 
-Currently, a huge amount of basic design information about buildings are produced  
-using modern Building Information Modeling (BIM) tools. There is a widely adopted standard, 
-IFC (Industry Foundation Classes) that provides a common representation of building designs developed in different design disciplines: architecture, structural engineering,
-MEP engineering, and so on. Nowadays, the IFC standard is supported by all major BIM
-vendors; their BIM tools can export building models as IFC files that can be exchanged with other
-parties.
+Currently, huge amounts of information about building designs is produced  
+using Building Information Modeling (BIM) tools. A widely adopted standard, 
+IFC (Industry Foundation Classes) provides a common representation for BIM models developed in 
+different design disciplines. All major BIM tools support IFC and can export BIM models to IFC files for exchange with other parties.
 
-The BIM models could provide a natural framework to organize and manage information gathered from other sources. However, a big obstacle to that is the difficulty to access and utilize BIM models. 
+BIM models could provide a natural framework to organize and manage information gathered from other 
+sources. However, a big obstacle is the difficulty to access and utilize BIM models. 
 In the realm of IFC the sharing of building models is based on point-to-point file exchange which 
-requires repeated manual work from designers.
-Information cannot be used in an online, granular, and interlinked manner, which provides a poor basis for open utilization of building information and the emergence of new, innovative applications.
+requires repeated manual work from designers. Information cannot be used in an online, granular, and 
+interlinked manner which provides a poor basis for open utilization of building information and the 
+emergence of new, innovative applications.
 
-To make building information more readily accessible, we have developed a \emph{WebIFC converter} to derive an OWL ontology from the IFC schema (\emph{ifcOWL}), and RDF datasets from IFC files (\emph{ifcRDF}).
-The ifcOWL ontology is used to classify concepts, characterize possible relationships, and define constraints in ifcRDF datasets which replicate the content of original IFC data files.
-All terms defined in the ifcOWL ontology must therefore be based on the corresponding constructs in the original IFC schema.
-The conversion from IFC data files to RDF is straightforward and can be done without information loss.
+To make building designs more openly accessible over the Web, we have developed the 
+\emph{WebIFC converter} to 
+derive OWL ontologies from the IFC schema (\ifcowl{}), and RDF datasets from IFC files 
+(\ifcrdf{}). The ifcOWL ontology is used to classify concepts, characterize possible relationships, 
+and define constraints in \ifcrdf{} datasets which replicate the content of original IFC data files. 
+All terms defined in the \ifcowl{} ontology must therefore be based on the corresponding constructs in 
+the original IFC schema. The conversion from IFC data files to RDF is straightforward and can be done 
+without information loss.
 
-Ideally, there should also be just one ifcOWL ontology for each version of the IFC schema (Fig.
-\ref{fig:ifcOWL-layers}, a).
-However, the conversion from the EXPRESS data definition language (that is used to specify the IFC schema) to OWL can be done in several different ways for the following reasons.
-Firstly, there is a \emph{mismatch} between the modeling constructs between EXPRESS and OWL.
-Secondly, for most practical purposes, the IFC data can be considered \emph{valid and consistent} with respect to the schema, since the IFC models are exported from BIM tools with certified converters.
-Thirdly, as the master data of the design models will remain in the native formats of the BIM tools and evolves over time, the icfRDF can in practice only be used in a \emph{read-only} fashion.
-Fourthly, the conversion of all possible IFC schema information will lead into a \emph{extensive} OWL ontology that mostly contains information not needed in practical applications or use cases.
-
-Consequently, we have converted the IFC schema into a \emph{flexible, multilayer ifcOWL ontology}.
-Different representational constructs are introduced at different levels.
-The ifcRDF is compatible with any of the layers; different layers just provide different amounts of type information that can be used with RDF.
+Ideally, there should also be just one \ifcowl{} ontology for each version of the IFC schema. However,
+there are possibilities and reasons to convert the IFC schema to OWL in several different ways for
+the following reasons. Firstly, there is a \emph{mismatch} between the modeling constructs of OWL and 
+EXPRESS data definition language used to specify the IFC schema, which creates
+different conversion options. Secondly, a simple version of \ifcowl{} \emph{compatible with
+restricted OWL profiles} (OWL 2 EL or QL) is useful to support ontology
+reasoning or inexact reasoning. Thirdly, for most practical purposes only a subset of ontology
+information is ever required, since no validation or checking of \ifcrdf{} data is needed; it can
+be considered \emph{practically valid} with respect to the schema, since IFC files are exported
+from BIM tools with certified converters, and icfRDF can sensibly only be used in a
+\emph{read-only} fashion, since BIM models will be maintained in the native formats of BIM
+tools. Fourthly, a full conversion of the IFC schema will make \ifcowl{}
+\emph{unnecessarily large}, and difficult to understand or manipulate.
 
@@
 \centering
-\includegraphics{images/ifcOWL-approaches.jpg}
-\caption{Two approaches of creating and using ifcOWL ontology}
+\includegraphics[angle=0,width=1.0\textwidth]{images/ifcOWL-multilayers.png}
+%\includegraphics{images/ifcOWL-multilayers.jpg}
+\caption{Multiple layers of \ifcowl{} ontology}
 \label{fig:ifcOWL-layers}
@@
 
-Different levels of details of the ifcOWL ontology are the following:
+Consequently, we have converted the IFC schema into a \emph{flexible, multilayer \ifcowl{} ontology}
+in a way that \emph{\ifcrdf{} is compatible with any of the layers} (Fig. \ref{fig:ifcOWL-layers}). 
+New representational constructs are introduced at more comprehensive levels which therefore 
+provide increasing detailed type information about \ifcrdf{} data. Different levels of details of the \ifcowl{} ontology are the following:
 
-\begin{enumerate}
-\item
-    \textbf{ifcOWL-Lite} (compatible with OWL 2 EL, OWL 2 QL, OWL 2 RL)
-    \begin{itemize}
-        \item Schema meta data
-        \item All type definitions and hierarchies
-        \item All entity type properties (names and range types)
-    \end{itemize}
-\item
-    \textbf{ifcOWL-Standard} (compatible with OWL 2 RL)
-    \begin{itemize}
-        \item Schema meta data
-        \item All entity type keys
-        \item All entity type inverse properties
-    \end{itemize}
-\item
-    \textbf{ifcOWL-Advanced} (compatible with OWL 2 RL)
-    \begin{itemize}
-        \item All properties' cardinality constraints
-        \item All list types' cardinality constraints
-    \end{itemize}
-\end{enumerate}
+\noindent
+\begin{tabular}{|l|l|l|}
+\hline
+\textbf{Layer} & Profiles & Included schema information \\
+\hline
+\hline
+\name{ifcOWL-Simple} & OWL 2 EL & Schema meta data \\
+& OWL 2 QL &  All type definitions and hierarchies \\
+& OWL 2 RL & All entity type properties (names, range types) \\
+\hline 
+\name{ifcOWL-Standard} & OWL 2 RL & Schema meta data \\
+& & All entity type keys \\
+& & All entity type inverse properties \\
+\hline
+\name{ifcOWL-Extended} & OWL 2 RL & All cardinality constraints of properties \\
+& & All cardinality constraints of list types \\
+\hline
+\end{tabular}
 
-Since all ifcRDF datasets are compatible with all layers of the ifcOWL ontology (Fig.
-\ref{fig:ifcOWL-layers}, b), a user can choose the best one based on the requirements of the concrete use case and tools.
-For instance, many Linked Data applications that need to query entity properties, extract the type hierarchy, or convert ifcRDF datasets back to object-oriented data models it is sufficient to use the \textbf{ifcOWL-Lite} ontology.
-In case of reasoning about data that includes inverse properties, it is recommended to use \textbf{ifcOWL-Standard} ontology.
-If an RDF dataset needs to be modified and data validated using type and cardinality constraints, \textbf{ifcOWL-Advanced} ontology is the right choice.
+Since all \ifcrdf{} datasets are compatible with all layers of the \ifcowl{} ontology 
+(Fig. \ref{fig:ifcOWL-layers}), a user can choose the best one based on the requirements of 
+the concrete use case and tools. For instance, many Linked Data applications that need to 
+query entity properties, extract the type hierarchy, or convert \ifcrdf{} datasets back to 
+object-oriented data models it is sufficient to use the \name{ifcOWL-Simple} ontology. 
+In case of reasoning about data that includes inverse properties, it is recommended to use 
+\name{ifcOWL-Standard} ontology. If an RDF dataset needs to be modified and data validated 
+using type and cardinality constraints, \name{ifcOWL-Extended} ontology is the correct choice.
 

### 2015-04-29 article-02-main.tex
Wrote sth about ifcRDF

@@
 \label{sec:ifcOWL}
+This section describes the derivation of the ifcOWL ontologies from the IFC specification.
+Since there are different options for the conversion, the choices are presented as conversion principles with justiciations.
 
-\subsection{Common Conversion Rules}
+\subsection{Common Conversion Principles}
 
-\begin{ontologyRule}[OWL profile]
-\ifclite{} layer must be compatible with any OWL 2 DL profile: OWL 2 EL, OWL 2 QL or OWL 2 RL.
-\ifcstandard{}, \ifcadvanced{} must be compatible with OWL 2 RL\footnote{Our WebIFC converter tool allows forced choice of other OWL profiles for any layer: OWL Lite, OWL DL, OWL Full, and OWL 2 Full}.
-\end{ontologyRule}
+\begin{principle}[Simple OWL profiles]
+\ifcowl{} should be compatible with as simple OWL 2 profiles as possible (to support a maximum variety of reasoners and other tools). 
+\end{principle}
 
-\TODO{Explain: why an OWL 2 profile has to be being chosen, how that choice will impact to the reasoning and querying performance, whether we can use other OWL 2 EL or QL for ifcOWL-Lite ontology}
+The first layer \ifcsimple{} can be specified to be compatible with any OWL 2 DL profile: OWL 2 EL, OWL 2 QL or OWL 2 RL.
+The other layers \ifcstandard{} and \ifcextended{} are compatible with OWL 2 RL\footnote{Our WebIFC converter tool allows forced choice of other OWL profiles for any layer: OWL Lite, OWL DL, OWL Full, and OWL 2 Full}.
+OWL 2 RL needed because it -- unlike OWL 2 EL or QL -- supports
+union data\-types (see \ref{subsec:ifcOWL-select-types}), enumerations of individuals (see \ref{subsec:ifcOWL-enum-types}), and anonymous individuals (see \ref{subsec:ifcRDF-naming-individuals}) required by the IFC schema.
+\cite{w3c:owl2-profiles}
 
-OWL 2 RL has been chosen for \ifcstandard{}, \ifcadvanced{} because in unlike OWL 2 EL and OWL 2 QL, it supports such needed things as: union data\-types (see \ref{subsec:ifcOWL-select-types}), enumerations of individuals (see \ref{subsec:ifcOWL-enum-types}), anonymous individuals (see \ref{subsec:ifcRDF-naming-individuals}), etc.\cite{w3c:owl2-profiles}.
+\begin{principle}[No domain and range restrictions]
+Property names should be kept simple and therefore domain and range constraints (specified by \name{rdfs:domain} and \name{rdfs:range}) should not be included in \ifcowl{} ontologies.
+\end{principle}
 
-\begin{ontologyRule}[Domain and range information]
-Property domain and range constraints specified by \name{rdfs:domain} and \name{rdfs:range} should not be included in ifcOWL ontologies.
-\end{ontologyRule}
+In the IFC schema the property names are \emph{local} to entity classes while in RDF they are \emph{global}. 
+In IFC the same property name can appear in multiple different classes. If the domain and range restrictions
+were included in  \ifcowl{}, the property names would need to be qualified with entity class names. This 
+would reduce the readability and usability of the \ifcowl{} significantly. In \ifcowl{} the domain and 
+range restrictions would mainly be used to infer types, since there is no need to check type constraints 
+like in programming languages \cite{w3c:owl-guide}. Moreover, the basic facilities given by them do not 
+provide any direct way to indicate property restrictions that are local to a class \cite{w3c:rdf-schema}. 
+Although it is possible to combine use \name{rdfs:domain} and \name{rdfs:range} with sub-property hierarchies, 
+direct support for such declarations is provided by OWL constructs, e.g. \name{owl:\-onProperty}.
 
-Properties \name{rdfs:\-domain} and \name{rdfs:\-range} are mainly used to infer types, but not to check type constraints like in programming languages \cite{w3c:owl-guide}.
-Moreover, the basic facilities given by them do not provide any direct way to indicate property restrictions that are local to a class \cite{w3c:rdf-schema}.
-Although it is possible to combine use \name{rdfs:domain} and \name{rdfs:range} with sub-property hierarchies, direct support for such declarations is provided by OWL constructs, e.g.
-\name{owl:\-onProperty}.
-
-\begin{ontologyRule}[Namespaces]
-All constructs related to EXPRESS, STEP or IFC specifications must be belong to namespaces with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly.
-At this moment the namespaces are defined as below (istead of \name{IFCXXX} there should be IFC versions, e.g.
-\name{IFC2x3} or \name{IFC4}):
+\begin{principle}[Namespaces]
+All constructs related to EXPRESS, STEP or IFC specifications must be belong to namespaces shown below with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (instead of \name{IFCXXX} there should be the name of IFC version, e.g.
+\name{IFC4_ADD1}):
+%\name{IFC2x3} or \name{IFC4}):
+%At this moment the namespaces are defined as below (instead of \name{IFCXXX} there should be IFC versions, e.g. 
+%\name{IFC2x3} or \name{IFC4}):
 
@@
 \end{lstlisting}
-\end{ontologyRule}
+\end{principle}
 
@@
 
-\begin{ontologyRule}[Simple data\-types]
+\begin{principle}[Simple data\-types]
 All simple data\-types except \emph{BOOLEAN} and \emph{LOGICAL} are declared as the same as their most similar primitive XSD data\-types which are supported by OWL 2 \cite{w3c:owl2-profiles} and are among preferred standard data\-types in Linked Data Platform \cite{w3c:ldp-best-practices}:
@@
 
-\begin{ontologyRule}[Simple data\-types with sizes, ifcOWL layers: all]If \name{STRING} and \name{BINARY} data\-types are specified with a size parameter, i.e.
-\name{STRING (max\-Length\-In\-Bytes)} or \name{BINARY(size\-In\-Bits)}, then in layers \lite{} and \standard{} they are defined as equivalent to data\-types without size.
-But in layer \advanced{}, the size parameter must be declared by using property \name{owl:onDatatype} and data\-type facets \name{xsd:maxLength} or \name{xsd:length} \cite{w3c:owl2-syntax,w3c:xmlschema11-2}, for instance:
+\begin{principle}[Simple data\-types with sizes, ifcOWL layers: all]If \name{STRING} and \name{BINARY} data\-types are specified with a size parameter, i.e.
+\name{STRING (max\-Length\-In\-Bytes)} or \name{BINARY(size\-In\-Bits)}, then in layers \simple{} and \standard{} they are defined as equivalent to data\-types without size.
+But in layer \extended{}, the size parameter must be declared by using property \name{owl:onDatatype} and data\-type facets \name{xsd:maxLength} or \name{xsd:length} \cite{w3c:owl2-syntax,w3c:xmlschema11-2}, for instance:
 
@@
 \begin{lstlisting}
-# only for Lite, Standard layers
+# only for Simple, Standard layers
 expr:STRING22 owl:sameAs expr:STRING .
 
-# only for Advanced layer
+# only for Extended layer
 expr:STRING22 owl:equivalentClass
@@
 
-\begin{ontologyRule}[Enumeration data\-types]
+\begin{principle}[Enumeration data\-types]
 Each enumeration data\-type in \textbf{ifc\-OWL} is defined as an \name{owl:Class} which is a subclass of \name{expr:\-Enum\-Class}.
 All possible values of enumeration data\-types are defined as named individuals.
-Enumeration data\-type memberships are defined by means of \name{rdf:type} in the \lite{} layer, and by means of \name{owl:oneOf} in \standard{} and \advanced{} layers as shown below:
+Enumeration data\-type memberships are defined by means of \name{rdf:type} in the \simple{} layer, and by means of \name{owl:oneOf} in \standard{} and \extended{} layers as shown below:
 % \begin{lstlisting}[caption={Enumeration data\-types}, label=lst:ifcOWL-enum-types]
 \begin{lstlisting}
-# only for Lite layer
+# only for Simple layer
 ifc:IfcAssemblyPlaceEnum a owl:Class ; rdfs:subclassOf expr:EnumClass .
@@
 
-# only for Standard, Advanced layers
+# only for Standard, Extended layers
 ifc:IfcAssemblyPlaceEnum a owl:Class ;  rdfs:subclassOf expr:EnumClass ;
@@
 
-\begin{ontologyRule}[Select data\-types]
+\begin{principle}[Select data\-types]
 A select data\-type must be converted to an \name{owl:Class} which is a subclass of \name{expr:\-Select\-Class}.
-Union-member data\-types are defined by means of \name{rdfs:\-subclassOf} in the \lite{} layer, and by means of \name{owl:unionOf} in \standard{} and \advanced{} layers as shown in the listing below.
-\end{ontologyRule}
+Union-member data\-types are defined by means of \name{rdfs:\-subclassOf} in the \simple{} layer, and by means of \name{owl:unionOf} in \standard{} and \extended{} layers as shown in the listing below.
+\end{principle}
 
@@
 \begin{lstlisting}
-# only for Lite layer
+# only for Simple layer
 ifc:IfcTrimmingSelect a owl:Class ; rdfs:subclassOf expr:SelectClass .
@@
 
-# only for Standard, Advanced layers
+# only for Standard, Extended layers
 ifc:IfcTrimmingSelect a owl:Class ; rdfs:subclassOf expr:SelectClass ;
@@
 
-\begin{ontologyRule}[Defined data\-types]
+\begin{principle}[Defined data\-types]
 If the defined data\-type is based on a simple data\-type then in ifcOWL it is declared as a \name{owl:Class} which is a subclass of \name{expr:\-Defined\-Class} has a single property \name{rdf:value} with that simple data\-type.
@@
 
-\begin{ontologyRule}[Entity data\-types]An entity data\-type is converted to an \name{owl:\-Class} which is subclass of \name{expr:\-Entity\-Class} or its superclass.
+\begin{principle}[Entity data\-types]An entity data\-type is converted to an \name{owl:\-Class} which is subclass of \name{expr:\-Entity\-Class} or its superclass.
 Entity attributes are defined by means of property \name{owl:on\-Pro\-per\-ty}.
@@
                       owl:allValuesFrom ifc:LIST_1_3_IfcLengthMeasure ] ;
-    # only for Advanced layer
+    # only for Extended layer
     rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:coordinates;
@@
 
-# only for Standard & Advanced layers
+# only for Standard & Extended layers
 [] a owl:AllDisjointClasses ;
@@
 
-# only for layer Advanced:
+# only for layer Extended:
  ifc:SET_2_UNDEFINED rdfs:subClassOf rdfs:subClassOf [ a owl:Restriction ;
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Generating ifcRDF Datasets From IFC Data
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Generating ifcRDF Datasets From IFC Data}
+\label{sec:ifcRDF}
+
+\subsection{Dataset Metadata}
+\label{subsec:ifcRDF-metadata}
+
+IFC data are exchanged by using ``STEP physical files'' (SPF).
+SPF format is defined in ISO 10303-21 \cite{wiki:step-file}.
+A SPF header section includes a list of STEP-specific metadata entities with types \name{FileDescription} \name{FileName} and \name{FileSchema} described in \cite{buildingSMART:ifc-header}.
+They are exported as objects of STEP entities, some attributes are duplicated as Dublin Core (DC) metadata annotations, for example:
+
+\begin{lstlisting}
+<http://example.org>  a void:DataSet ;
+    dcterms:created "2015-02-19T13:22:15Z"^^xsd:dateTime ;
+    dcterms:creator "John Smith"^^xsd:string .
+    step:fileName  [ a step:FileName ;
+        step:author "John Smith"^^xsd:string ;
+        step:originating_system "Tekla Structures 20.1"^^xsd:string ; ] ;
+        step:time_stamp "2015-02-19T13:22:15Z"^^xsd:dateTime  ] .
+\end{lstlisting}
+
+In a SPF data section each entity is declared with a line number, an entity data\-type and a list of values.
+All entities of \name{Ifc\-Root}-derived types are converted into URI resources with format \name{GUID\_\textless{}guid\textgreater{}} where \name{guid} must be an normal GUID, not a compressed IFC GUID which may include a sign \name{\$}.
+Regarding to other entities, user can choose whether to export them as anonymous individuals or name them in a format like \name{LINE\_\textless{}lineNumber\textgreater{}}.
+Note that line numbers are used only for linking entities inside a SPF, and they are unstable.
+Naming is required when the dataset is used together with OWL 2 EL or OWL 2 QL because these profiles do not allow anonymous individuals.
+The namespace of named entities is also defined by user.
+Below is an example of IFC data exported to an RDF dataset which is compatible with OWL 2 RL:
 
@@
     a               ifc:IfcPropertySet ; # entity type
-    ifc:globalId    [ a ifc:IfcGloballyUniqueId ; # defined/select type
+    ifc:globalId    [ a ifc:IfcGloballyUniqueId ; # defined type
                      rdf:value "0P5dbFLWL4cPgmyLNUZJji"^^xsd:string ] ;
@@
 
-\subsection{Metadata}
+%Data\-type of an attribute value is declared obviously only if the attribute has a select data\-type and the value has a defined data\-type.
 
-\TODO {Recommendations about model namespaces, naming individuals, defined types of select types }
+\section{Conclusion}
+\label{sec:conclusion}

### 2015-04-29 article-00-headers.tex
Updates from ShareLaTeX
Reduced number of headings

@@
 % TODO text
-\newcommand{\TODO}[1]{\highlight{\textbf{TODO: }{#1}}}
+% \newcommand{\TODO}[1]{\highlight{\textbf{TODO: }{#1}}}
+\newcommand{\TODO}[1]{{}}
 
@@
 \newcommand{\ifcextended}{\ifcowl{}-\extended{}}
+\newcommand{\ifcrdf}{\textbf{ifcRDF}}
 
@@
 We present an ontology derived from the conceptualization of IFC, the existing standard for representing building designs, and a tool to convert building designs from IFC files to RDF datasets that conform to the ontology.
-Replacing the traditional point-to-point exchange of IFC files with publication of RDF datasets  by using Linked Data principles and tools enables information integration across different IFC-based and other smart city datasets, and the granular access to individual components.
-It paves the way to a distributed and loosely coupled architecture for building information management.
-Opening selected parts of construction data to a wider set of stakeholders creates opportunities for innovative applications in the smart cities of the future.
+Replacing the traditional point-to-point exchange of IFC files with publication of RDF datasets using Linked Data principles and tools enables information integration across different IFC-based and other smart city datasets, and the granular access to individual components.
+It paves the way to a distributed and loosely coupled architecture for building information management, and provides opportunities for innovative applications.
 

### 2015-04-29 article-01-introduction.tex
Updates from ShareLaTeX
Reduced number of headings

@@
 Currently, huge amounts of information about building designs is produced  
-using Building Information Modeling (BIM) tools. A widely adopted standard, 
-IFC (Industry Foundation Classes) provides a common representation for BIM models developed in 
+using Building Information Modeling (BIM) tools \cite{eastman2011bim}. A widely adopted standard, 
+IFC (Industry Foundation Classes) \cite{ISO16739,liebich2010unveiling} provides a common representation for BIM models of 
 different design disciplines. All major BIM tools support IFC and can export BIM models to IFC files for exchange with other parties.
 
-BIM models could provide a natural framework to organize and manage information gathered from other 
-sources. However, a big obstacle is the difficulty to access and utilize BIM models. 
-In the realm of IFC the sharing of building models is based on point-to-point file exchange which 
-requires repeated manual work from designers. Information cannot be used in an online, granular, and 
-interlinked manner which provides a poor basis for open utilization of building information and the 
+BIM models could provide a natural framework to organize and manage information about buildings.
+However, a big obstacle is the difficulty to access and utilize BIM models which have traditionally been shared shared through point-to-point file exchange.
+They cannot be used in an online, granular, and
+interlinked manner and provides a poor basis for open utilization of building information and the 
 emergence of new, innovative applications.
-
-To make building designs more openly accessible over the Web, we have developed the 
-\emph{WebIFC converter} to 
+To make BIM models accessible on the Web \cite{torma2014wobd}, we have developed the 
+\name{WebIFC converter} to 
 derive OWL ontologies from the IFC schema (\ifcowl{}), and RDF datasets from IFC files 
-(\ifcrdf{}). The ifcOWL ontology is used to classify concepts, characterize possible relationships, 
-and define constraints in \ifcrdf{} datasets which replicate the content of original IFC data files. 
-All terms defined in the \ifcowl{} ontology must therefore be based on the corresponding constructs in 
-the original IFC schema. The conversion from IFC data files to RDF is straightforward and can be done 
-without information loss.
-
-Ideally, there should also be just one \ifcowl{} ontology for each version of the IFC schema. However,
-there are possibilities and reasons to convert the IFC schema to OWL in several different ways for
-the following reasons. Firstly, there is a \emph{mismatch} between the modeling constructs of OWL and 
-EXPRESS data definition language used to specify the IFC schema, which creates
-different conversion options. Secondly, a simple version of \ifcowl{} \emph{compatible with
-restricted OWL profiles} (OWL 2 EL or QL) is useful to support ontology
-reasoning or inexact reasoning. Thirdly, for most practical purposes only a subset of ontology
-information is ever required, since no validation or checking of \ifcrdf{} data is needed; it can
-be considered \emph{practically valid} with respect to the schema, since IFC files are exported
-from BIM tools with certified converters, and icfRDF can sensibly only be used in a
-\emph{read-only} fashion, since BIM models will be maintained in the native formats of BIM
-tools. Fourthly, a full conversion of the IFC schema will make \ifcowl{}
-\emph{unnecessarily large}, and difficult to understand or manipulate.
+(\ifcrdf{}). The \ifcowl{} specifies concepts, relations, and constraints for \ifcrdf{} datasets that replicate the content of IFC files. 
+All terms defined in \ifcowl{}  must thus be based on the corresponding constructs in 
+the IFC schema.
 
@@
 \centering
-\includegraphics[angle=0,width=1.0\textwidth]{images/ifcOWL-multilayers.png}
+\includegraphics[angle=0,width=0.9\textwidth]{images/ifcOWL-multilayers.png}
 %\includegraphics{images/ifcOWL-multilayers.jpg}
@@
 
-Consequently, we have converted the IFC schema into a \emph{flexible, multilayer \ifcowl{} ontology}
+The conversion of IFC files to \ifcrdf{} is straightforward and can be done 
+without information loss.
+Ideally, there would also be just one \ifcowl{} ontology for each version of the IFC schema. However,
+there are possibilities and reasons to convert the IFC schema to OWL in several different ways. Firstly, there is a \emph{mismatch} between the modeling constructs of OWL and 
+EXPRESS data definition language used to specify the IFC schema. Secondly, a simple version of \ifcowl{} \emph{compatible with
+restricted OWL profiles} (OWL 2 EL or QL) \cite{motik2009owl} is useful to support ontology
+reasoning or inexact reasoning. Thirdly, for most practical purposes only a subset of ontology
+information is required, as no validation or checking of \ifcrdf{} data is needed; it can
+be considered \emph{practically valid} w.r.t. the schema, since BIM tools use certified converters, and icfRDF can sensibly only be used in a
+\emph{read-only} fashion, since BIM models will be maintained in the native formats of BIM
+tools. Fourthly, a full conversion of the IFC schema will make \ifcowl{}
+\emph{unnecessarily large}, and difficult to understand or manipulate.
+
+Consequently, we convert the IFC schema into a \emph{flexible, multilayer \ifcowl{} ontology}
 in a way that \emph{\ifcrdf{} is compatible with any of the layers} (Fig. \ref{fig:ifcOWL-layers}). 
 New representational constructs are introduced at more comprehensive levels which therefore 
-provide increasing detailed type information about \ifcrdf{} data. Different levels of details of the \ifcowl{} ontology are the following:
+provide increasing detailed type information about \ifcrdf{} data:
 
-\noindent
+\vspace{2mm}
+% \noindent
+\begin{small}
 \begin{tabular}{|l|l|l|}
  \hline
-\textbf{Layer} & Profiles & Included schema information \\
+\textbf{Layer} & \textbf{Profiles} & \textbf{Included schema information} \\
 \hline
@@
 \end{tabular}
+\end{small}
 
-Since all \ifcrdf{} datasets are compatible with all layers of the \ifcowl{} ontology 
-(Fig. \ref{fig:ifcOWL-layers}), a user can choose the best one based on the requirements of 
-the concrete use case and tools. For instance, many Linked Data applications that need to 
-query entity properties, extract the type hierarchy, or convert \ifcrdf{} datasets back to 
-object-oriented data models it is sufficient to use the \name{ifcOWL-Simple} ontology. 
-In case of reasoning about data that includes inverse properties, it is recommended to use 
-\name{ifcOWL-Standard} ontology. If an RDF dataset needs to be modified and data validated 
-using type and cardinality constraints, \name{ifcOWL-Extended} ontology is the correct choice.
+%Since all \ifcrdf{} datasets are compatible with all layers of the \ifcowl{} ontology 
+%(Fig. \ref{fig:ifcOWL-layers}), the most appropriate one can be selected based on the requirements of 
+%a use case and tools. For instance, many Linked Data applications that need to 
+%query entity properties, use the type hierarchy, or convert \ifcrdf{} datasets back to 
+%object-oriented data models it is sufficient to use the \name{ifcOWL-Simple} ontology. 
+%In case of reasoning about data that includes inverse properties, it is recommended to use 
+%\name{ifcOWL-Standard} ontology. If an RDF dataset needs to be modified and data validated 
+%using type and cardinality constraints, \name{ifcOWL-Extended} is the correct choice.
+
+From the previous IFC-to-OWL/RDF conversions
+\cite{beetz2005ontology,beetz2009ifcowl,pauwels2011interoperability} our approach differs since it
+is based on OWL 2 profiles, produces a flexible multilayer ontology, and is specifically targeted to
+real-life Linked Data applications. It uses stable URIs for entities, in contrast to the previous approaches
+where the URIs were instable across models versions.
 

### 2015-04-29 article-02-main.tex
Reduced number of headings

@@
 
-\section{Generating ifcOWL Ontologies From IFC Specifications}
+\section{Principles of Generating ifcOWL Ontologies And ifcRDF Datasets}
 \label{sec:ifcOWL}
@@
 
-\begin{principle}[Simple OWL profiles]
-\ifcowl{} should be compatible with as simple OWL 2 profiles as possible (to support a maximum variety of reasoners and other tools). 
+\begin{principle}%[Simple OWL profiles]
+\ifcsimple{} should be compatible with as simple OWL 2 profiles as possible (to support a maximum variety of reasoners and other tools). 
 \end{principle}
@@
 The first layer \ifcsimple{} can be specified to be compatible with any OWL 2 DL profile: OWL 2 EL, OWL 2 QL or OWL 2 RL.
-The other layers \ifcstandard{} and \ifcextended{} are compatible with OWL 2 RL\footnote{Our WebIFC converter tool allows forced choice of other OWL profiles for any layer: OWL Lite, OWL DL, OWL Full, and OWL 2 Full}.
-OWL 2 RL needed because it -- unlike OWL 2 EL or QL -- supports
-union data\-types (see \ref{subsec:ifcOWL-select-types}), enumerations of individuals (see \ref{subsec:ifcOWL-enum-types}), and anonymous individuals (see \ref{subsec:ifcRDF-naming-individuals}) required by the IFC schema.
+The other layers \ifcstandard{} and \ifcextended{} are compatible with OWL 2 RL.
+OWL 2 RL needed because it -- unlike OWL 2 EL or QL -- supports such necessary things like enumerations of individuals, union classes, disjoint classes, and also anonymous individuals.
 \cite{w3c:owl2-profiles}
 
-\begin{principle}[No domain and range restrictions]
+\begin{principle}%[No domain and range restrictions]
 Property names should be kept simple and therefore domain and range constraints (specified by \name{rdfs:domain} and \name{rdfs:range}) should not be included in \ifcowl{} ontologies.
@@
 
-\begin{principle}[Namespaces]
+\begin{principle}%[Namespaces]
 All constructs related to EXPRESS, STEP or IFC specifications must be belong to namespaces shown below with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (instead of \name{IFCXXX} there should be the name of IFC version, e.g.
-\name{IFC4_ADD1}):
+\name{IFC4\_ADD1}):
 %\name{IFC2x3} or \name{IFC4}):
@@
 
+\begin{principle}All definitions of functions, rules, type constraints and derived attributes specified in EXPRESS after keywords \name{FUNCTION}, \name{RULE}, \name{WHERE} and \name{RULE} accordingly should be ignored.
+\end{principle}
+
+\TODO{Explain why functions, rules, etc. should be ignored}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\subsection{Components of an EXPRESS schema}
+\subsection{Datatype Declarations}
+\label{subsec:ifcOWL-types}
+
+\TODO{\textit{Write about EXPRESS in the Introduction section}:\\
 The IFC specification is written as an EXPRESS schema.
-An EXPRESS schema defines series of data\-types, functions and rules by using EXPRESS data definition language specified in ISO10303-11 \cite{wiki:express,noauthor:ifc-guide}.
+An EXPRESS schema defines series of data\-types, functions and rules by using EXPRESS data definition language specified in ISO10303-11 \cite{wiki:express,noauthor:ifc-guide}.}
 
-There are six groups of data\-types: simple data\-types, entity data\-types, enumeration data\-types, defined (declared) data\-types, select data\-types and aggregation data\-types.
-All types must be declared obviously by using constructs TYPE and ENTITY except the built-in simple data\-types.
-Definitions and characteristics of data\-types are considered below.
-EXPRESS schemata also contain definitions of functions, rules and type constraints most of which cannot be reflected in OWL 2 at all.
+An EXPRESS schema includes six groups of data\-types: built-in simple data\-types, entity data\-types, enumeration data\-types, defined data\-types, select data\-types and aggregation data\-types.
 
@@
 
-\subsection{Simple Datatype Declarations}
-\label{subsec:ifcOWL-simple-type}
+\subsubsection{Simple Datatypes} built in EXPRESS are: \name{STRING}, \name{BI\-NA\-RY}, \name{IN\-TE\-GER}, \name{REAL}, \name{NUM\-BER}, \name{BOO\-LEAN} and \name{LO\-GI\-CAL}.
+Types \name{STRING} and \name{BINARY} data\-types may be specified with a size parameter: \name{STRING (max\-Length\-In\-Bytes)} or \name{BINARY(size\-In\-Bits)}.
 
-Simple data\-types built in EXPRESS are: \name{STRING}, \name{BI\-NA\-RY}, \name{IN\-TE\-GER}, \name{REAL}, \name{NUM\-BER}, \name{BOO\-LEAN} and \name{LO\-GI\-CAL}.
-However, type \name{LOGICAL} which has three possible values \name{TRUE}, \name{FALSE} and \name{UNKNOWN} is considered by us as an enumeration data\-type (see \ref{subsec:ifcOWL-enum-types}).
-The same to type \name{BOOLEAN} which is like a primitive boolean data\-type but semantically close and met with the same frequency in IFC4 as \name{LOGICAL}.
-
-\begin{principle}[Simple data\-types]
-All simple data\-types except \emph{BOOLEAN} and \emph{LOGICAL} are declared as the same as their most similar primitive XSD data\-types which are supported by OWL 2 \cite{w3c:owl2-profiles} and are among preferred standard data\-types in Linked Data Platform \cite{w3c:ldp-best-practices}:
+\begin{principle}%[Simple data\-types]
+A simple data\-type (except \emph{BOOLEAN} and \emph{LOGICAL}) should be declared in \ifcowl{} as \name{owl:sameAs} of its most similar primitive XSD data\-type among all data\-types supported by any OWL 2 profile \cite{w3c:owl2-profiles} and preferred in Linked Data Platform \cite{w3c:ldp-best-practices}:
 
@@
 
-`Most similar XSD data\-types' mean the types from XSD schema that conform to the value spaces of the original EXPRESS and IFC data\-types better than other ones.
+Type \name{LOGICAL} -- with its three possible values \name{TRUE}, \name{FALSE} and \name{UNKNOWN} -- is considered as an enumeration data\-type.
+The same to \name{BOOLEAN} because OWL 2 EL and QL do not support type \name{xsd:boolean} \cite{w3c:owl2-profiles}.
+
+Statement `most similar XSD data\-type' means the type from XSD schema that conform to the value spaces of the EXPRESS simple data\-type better than any other type.
 For instance, type \name{xsd:double} is more similar to \name{REAL} than \name{xsd:decimal} as it is an IEEE 64-bit floating-point data\-type and supports scientific notation \cite{w3c:xsd,datapic:xsd}.
-It is recommended to use XSD data\-type names directly when they are needed.
+When it is needed to refer to an EXPRRESS simple data\-type, we recommend to use its XSD equivalent type directly.
 
-\begin{principle}[Simple data\-types with sizes, ifcOWL layers: all]If \name{STRING} and \name{BINARY} data\-types are specified with a size parameter, i.e.
-\name{STRING (max\-Length\-In\-Bytes)} or \name{BINARY(size\-In\-Bits)}, then in layers \simple{} and \standard{} they are defined as equivalent to data\-types without size.
+\begin{principle}%[Simple data\-types with sizes]
+Size parameters of types \name{STRING} and \name{BINARY} (if any) should be ignored in layers \simple{} and \standard{} and should be declared obviously -- by means of OWL 2 data\-type facets -- in layer \extended{}, for instance:
+
+% they are defined as equivalent to data\-types without size.
 But in layer \extended{}, the size parameter must be declared by using property \name{owl:onDatatype} and data\-type facets \name{xsd:maxLength} or \name{xsd:length} \cite{w3c:owl2-syntax,w3c:xmlschema11-2}, for instance:
@@
 # only for Simple, Standard layers
-expr:STRING22 owl:sameAs expr:STRING .
-
+  expr:STRING22 owl:sameAs xsd:string .
 # only for Extended layer
-expr:STRING22 owl:equivalentClass
-    [   a rdfs:Datatype ;
-        owl:onDatatype  xsd:string ;
-        owl:withRestrictions ( [ xsd:length "22"^^xsd:integer ] ) ] .
+  expr:STRING22 owl:equivalentClass [   a rdfs:Datatype ;
+    owl:onDatatype  xsd:string; owl:withRestrictions ( [ xsd:length 22 ] ) ].
 \end{lstlisting}
@@
 
-\subsection{Enumeration Datatype Declarations}
-\label{subsec:ifcOWL-enum-types}
+\subsubsection{Enumeration Datatypes} represent fixed sets of possible values.
 
-An enumeration data\-type represents a set of possible fixed values.
+\begin{principle}%[Enumeration data\-types]
+Each enumeration data\-type should be defined in \ifcowl{} as an \name{owl:Class} which is a subclass of \name{expr:\-Enum\-Class}.
+Enumeration values must be defined as named individuals of the enumeration class but in different ways in different layers as shown below:
 
-\begin{principle}[Enumeration data\-types]
-Each enumeration data\-type in \textbf{ifc\-OWL} is defined as an \name{owl:Class} which is a subclass of \name{expr:\-Enum\-Class}.
-All possible values of enumeration data\-types are defined as named individuals.
-Enumeration data\-type memberships are defined by means of \name{rdf:type} in the \simple{} layer, and by means of \name{owl:oneOf} in \standard{} and \extended{} layers as shown below:
 % \begin{lstlisting}[caption={Enumeration data\-types}, label=lst:ifcOWL-enum-types]
@@
   ifc:SITE, ifc:FACTORY, ifc:NOTDEFINED rdf:type ifc:IfcAddressTypeEnum .
-
 # only for Standard, Extended layers
@@
 
-The reason of two different approaches is that enumeration with more than one individual or literal are not supported in OWL 2 EL or QL.
-In both cases enumeration values are defined as individuals of one class so in terms of reasnoning logic they are equivalent.
-The enumeration values are defined as individuals because OWL 2 RL supports \name{Object\-One\-Of} to define enumeration of individuals, but not \name{Data\-One\-Of} \cite{w3c:owl2-profiles}.
-The fact that some values such as \name{ifc:NOT\-DE\-FI\-NED} belong to different enumeration data\-types should not cause any problems since all enumeration values in ifc\-RDF datasets are defined together with a concrete enumeration data\-type (see \ref{subsec:ifcRDF-enum-values}).
-It is also very easy to get all possible values of an enumeration data\-type by using simple SPARQL queries.
+Two different approaches are used because enumerations with more than one individual or literal are not supported in OWL 2 EL or QL, like enumerations of literals in OWL 2 RL.
+The fact that some values -- for instance, \name{ifc:NOT\-DE\-FI\-NED} -- belong to different enumeration data\-types should not cause any problems since any enumeration value is always specified with an exact type.
 
-As mentioned above, simple data\-types \name{BOOLEAN} and \name{LOGICAL} are considered by us as enumeration data\-types.
-But these types and their values must be defined in the EXPRESS namespace, e.g.
-\name{expr:LOGICAL}, \name{expr:TRUE}, \name{expr:FALSE} or \name{expr:UNKNOWN}.
+As mentioned above, simple data\-types \name{BOOLEAN} and \name{LOGICAL} are considered as enumeration data\-types.
+But these types and their values must be defined in the EXPRESS-related namespace, e.g.
+\name{expr:LOGICAL}, \name{expr:TRUE}, \name{expr:UNKNOWN}.
 
@@
 
-\subsection{Select Datatype Declarations}
-\label{subsec:ifcOWL-select-types}
+\subsubsection{Select Datatypes} represent unions of other data\-types.
+Particular members of a select data\-type can be entity, defined or select data\-types.
+For instance, type \name{Ifc\-Trim\-ming\-Se\-lect} is an union of entity data\-type \name{Ifc\-Car\-te\-sian\-Point} and defined data\-type \name{Ifc\-Pa\-ra\-me\-ter\-Va\-lue}.
 
-A select data\-type represents an union of other data\-types.
-Union members of a select data\-type usually are either entity data\-types, or defined data\-types all together.
-But there are a few exceptions, for instance, type \name{Ifc\-Trim\-ming\-Se\-lect} is an union of entity data\-type \name{Ifc\-Car\-te\-sian\-Point} and defined data\-type \name{Ifc\-Pa\-ra\-me\-ter\-Va\-lue}.
-
-\begin{principle}[Select data\-types]
+\begin{principle}%[Select data\-types]
 A select data\-type must be converted to an \name{owl:Class} which is a subclass of \name{expr:\-Select\-Class}.
-Union-member data\-types are defined by means of \name{rdfs:\-subclassOf} in the \simple{} layer, and by means of \name{owl:unionOf} in \standard{} and \extended{} layers as shown in the listing below.
+Its particular data\-types are defined as subclasses of the select class, but in different ways depended on the layer:
 \end{principle}
@@
     rdfs:subclassOf ifc:IfcTrimmingSelect .
-
 # only for Standard, Extended layers
@@
 
-Similarly to the above case, two different approaches are used because property \name{owl:unionOf} supported only in OWL 2 RL, but not in OWL 2 EL and OWL 2 QL \cite{w3c:owl2-profiles}.
+Two different approaches are used because property \name{owl:unionOf} is supported only in OWL 2 RL, but not in OWL 2 EL and QL \cite{w3c:owl2-profiles}.
 
@@
 
-\subsection{Defined Datatype Declarations}
-\label{subsec:ifcOWL-defined-types}
+\subsubsection{Defined (Declared, Named) Datatypes} are usually specified as equivalent or subclass of simple data\-types or other defined data\-types.
+For instance, type \name{Ifc\-Length\-Mea\-sure} is equal to \name{REAL}, while type \name{Ifc\-Po\-si\-ti\-ve\-Length\-Mea\-su\-re} is a subset of \name{Ifc\-Length\-Mea\-sure} with some constraint.
 
-A defined (declared, named) data\-type is usually specified as equivalent to a simple data\-type or another defined data\-type, with or without a additional constraint.
-For instance, type \name{Ifc\-Length\-Mea\-sure} is defined as equal to \name{REAL}, and type \name{Ifc\-Po\-si\-ti\-ve\-Length\-Mea\-su\-re} is a set of positive \name{Ifc\-Length\-Mea\-sure} values.
-
-\begin{principle}[Defined data\-types]
-If the defined data\-type is based on a simple data\-type then in ifcOWL it is declared as a \name{owl:Class} which is a subclass of \name{expr:\-Defined\-Class} has a single property \name{rdf:value} with that simple data\-type.
-If the defined data\-type is based on another defined data\-type then it is declared as a subclass of the \name{owl:Class} of that defined data\-type, for instance:
+\begin{principle}%[Defined data\-types]
+A defined data\-type based on a simple data\-type should be declared as an \name{owl:Class} which is a subclass of \name{expr:\-Defined\-Class} and has a single property \name{rdf:value} with that simple data\-type.
+A defined data\-type based on another defined data\-type should be declared as a subclass of the base defined data\-type, for instance:
 
 %\begin{lstlisting}[caption={Defined data\-types}, label=lst:ifcOWL-defined-types]
-\begin{lstlisting}[aboveskip=3pt]
+\begin{lstlisting}
 ifc:IfcLengthMeasure a owl:Class ; rdfs:subClassOf expr:DefinedClass ;
@@
                       owl:allValuesFrom xsd:double ] .
-
-ifc:IfcPositiveLengthMeasure a owl:Class ;
-    rdfs:subClassOf ifc:IfcLengthMeasure .
+ifc:IfcPositiveLengthMeasure rdfs:subClassOf ifc:IfcLengthMeasure .
 \end{lstlisting}
@@
 
-Defined data\-types are not declared as a \name{rdfs:Datatype}, but \name{owl:Class}es which wrap \name{rdfs:Datatype} values because in contract to simple data types they can be used for declaration of select data\-types (see \ref{subsec:ifcOWL-select-types}).
+Defined data\-types are defined as \name{owl:Class}es instead of \name{rdfs:Datatype}s because they can be merged with other \name{owl:Class}es for creating select data\-types.
+In addition, different defined data\-types wrapping the same \name{rdfs:Datatype} are not the same.
+For example, \name{IfcLengthMeasure} is different to \name{IfcMassMeasure} although they both are \name{REAL} and belong to select class \name{IfcMeasureValue}.
 
-Defined data\-type \name{Ifc\-Time\-Stamp} is declared as equivalent to \name{INTEGER} but represents so-called Unix time value that can be easily converted to \name{xsd:date\-Time}.
-So it wraps a \name{xsd:date\-Time} value instead of \name{xsd:integer}.
+In principle type \name{Ifc\-Time\-Stamp} must wrap an integer value which represents so-called Unix time.
+But we decided this value should have type \name{xsd:date\-Time}, and all values of \name{Ifc\-Time\-Stamp} must be converted to a standard datetime values.
 
@@
 
-\subsection{Entity Datatype Declarations}
-\label{subsec:ifcOWL-entity-types}
+\subsubsection{Entity Datatypes} are identical to classes in object-oriented programming languages.
+Each type may have one super\-type.
+An attribute is defined with a data\-type, cardinalities and can be marked as optional.
 
-Entity data\-types are identical to classes in object-oriented programming languages.
-Each type may have not more than one super\-type that it inherits attributes from.
-An attribute is defined with a data\-type, a minimum and a maximum cardinalities, can be marked as optional.
+\begin{principle}%[Entity data\-types]
+An entity data\-type should be converted to an \name{owl:\-Class} which is subclass of its superclass (if any) or \name{expr:\-Entity\-Class}.
+Entity attribute cardinalities can be only from 0 to 1, even if this attribute is a set of values.
+In this case the attribute type is an external aggregated data\-type, for example:
 
-\begin{principle}[Entity data\-types]An entity data\-type is converted to an \name{owl:\-Class} which is subclass of \name{expr:\-Entity\-Class} or its superclass.
-Entity attributes are defined by means of property \name{owl:on\-Pro\-per\-ty}.
-If an attribute is optional then its minimum qualified cardinality is 0, otherwise -- 1.
-Maximum qualified of all attributes are always 1.
-In case if an attribute is a list or a set of some data\-type then its type is considered as an external aggregated data\-type, for example:
 %\begin{lstlisting}[caption={Entity data\-types}, label=lst:ifcOWL-defined-types]
@@
 ifc:IfcProperty a owl:Class ; rdfs:subClassOf expr:EntityClass .
-
 ifc:IfcComplexProperty a owl:Class ; rdfs:subClassOf ifc:IfcProperty ;
@@
 
-\subsection{Aggregated Type Declarations}
-\label{subsec:ifcOWL-aggregated-types}
+\subsubsection{Aggregated Types} may be unordered (\name{SET} and \name{BAG}) or  ordered (\name{LIST} and \name{ARRAY}) collections.
+\name{BAG} may contain duplicated values, unlike \name{SET}.
+\name{ARRAY} is specified with start and end indexes while other types are specified with min and max cardinalities.
 
-Aggregation data\-types may be unordered (\name{SET} and \name{BAG}) or  ordered (\name{LIST} and \name{ARRAY}).
-\name{BAG} may contain a particular value more than once, unlike \name{SET}.
-\name{ARRAY} is specified with starting and ending indices while other types are specified with min and max cardinalities.
+\begin{principle}Any aggregated type must be subclass of one of \name{expr:SetClass}, \name{expr:ListClass}, \name{expr:ArrayClass} and \name{expr:BagClass} which are subclasses of \name{expr:\-Collec\-tion\-Class}.
+The latter has one a nonfunctional property \name{expr:slot} with type \name{expr:Slot} and four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index}.
+Class \name{expr:Slot} has two functional properties \name{expr:item} (the actual collection item value), \name{expr:index} and also optional functional properties \name{expr:previous}, \name{expr:next}.
+Below is a fragment of definition of type \name{SET [2:?] OF IfcProfileDef}:
+\end{principle}
 
-In order to declare these types, new classes were created in a similar way as Ordered List Ontoloty \cite{olo:ontology}:
-\begin{itemize}
-\item Class \name{expr:\-Collec\-tion\-Class} has four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index} and a nonfunctional property \name{expr:slot} with type \name{expr:Slot}.
-\item Class \name{expr:Slot} has two functional properties \name{expr:item} (the actual collection item value), \name{expr:index} and also optional functional properties \name{expr:previous}, \name{expr:next}.
-\item Classes \name{expr:SetClass}, \name{expr:ListClass}, \name{expr:ArrayClass} and \name{expr:BagClass} are subclasses of \name{expr:\-Collec\-tion\-Class}.
-\end{itemize}
+% In order to declare these types, new classes were created in a similar way as Ordered List Ontoloty \cite{olo:ontology}:
+% \begin{itemize}
+% \item Class \name{expr:\-Collec\-tion\-Class} has four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index} and a nonfunctional property \name{expr:slot} with type \name{expr:Slot}.
+% \item Class \name{expr:Slot} has two functional properties \name{expr:item} (the actual collection item value), \name{expr:index} and also optional functional properties \name{expr:previous}, \name{expr:next}.
+% \item Classes \name{expr:SetClass}, \name{expr:ListClass}, \name{expr:ArrayClass} and \name{expr:BagClass} are subclasses of \name{expr:\-Collec\-tion\-Class}.
+% \end{itemize}
 
-Below is a fragment of definition of type \name{SET [2:?] OF IfcProfileDef}:
+% Below is a fragment of definition of type \name{SET [2:?] OF IfcProfileDef}:
 
@@
     rdfs:subClassOf ifc:SET_2_UNDEFINED , ifc:SET_OF_IfcProfileDef .
-
- ifc:SET_2_UNDEFINED, SET_OF_IfcProfileDef a owl:Class ;
-    rdfs:subClassOf expr:SetClass .
+ ifc:SET_2_UNDEFINED, SET_OF_IfcProfileDef rdfs:subClassOf expr:SetClass .
 
@@
     owl:onProperty expr:slot ; owl:minCardinality "2"^^xsd:integer ] .
-
  ifc:SET_OF_IfcProfileDef rdfs:subClassOf [ a owl:Restriction ;
     owl:onProperty expr:slot ; owl:allValuesFrom ifc:SLOT_OF_IfcProfileDef ].
-
  ifc:SLOT_OF_IfcProfileDef rdfs:subClassOf [ a owl:Restriction ; owl:onProperty expr:item ; owl:allValuesFrom ifc:IfcProfileDef ] .
@@
 
-\section{Generating ifcRDF Datasets From IFC Data}
+\subsection{Generating ifcRDF Datasets From IFC Data}
 \label{sec:ifcRDF}
 
-\subsection{Dataset Metadata}
-\label{subsec:ifcRDF-metadata}
+\subsubsection{Dataset Metadata}
+%\label{subsec:ifcRDF-metadata}
 
@@
     ifc:hasProperties  _:b11 .
-
 _:b11 a ifc:SET_1_UNLIMITED_OF_IfcProperty ; # entity type
@@
     expr:slot _:b12, _:b13, _:b14 .
-
 _:b12 a ifc:SLOT_OF_IfcProperty ;

### 2015-04-29 article-00-headers.tex
Changed Principle style

@@
 
-\spnewtheorem{principle}{Principle}{\bfseries}{\rmfamily}
+%\spnewtheorem{principle}{Principle}{\bfseries}{\rmfamily}
+
+\AtBeginDocument{
+    \newtheoremstyle{principleStyle}
+        {1.5\topsep} % Space above, default \topsep
+        {\topsep} % Space below, default \topsep
+        {\itshape} % Body font
+        {} % Indent amount, example: \parindent
+        {\bfseries\itshape} % Theorem head font
+        {.} % Punctuation after theorem head
+        {.5em} % Space after theorem head
+        {} % Theorem head spec (can be left empty, meaning `normal')
+
+    \theoremstyle{principleStyle} \newtheorem{principle}{Principle}
+}
 
 %\spnewtheorem{datasetsRule}{Datasets Rule}{\bfseries}{\rmfamily}
+
+%
+% THEOREMS
+%
+\usepackage{amsthm}
 
@@
 % ifcOWL layer names
-\newcommand{\simple}{\textbf{Simple}}
-\newcommand{\standard}{\textbf{Standard}}
-\newcommand{\extended}{\textbf{Extended}}
+\newcommand{\simple}{\textnormal{\textbf{Simple}}}
+\newcommand{\standard}{\textnormal{\textbf{Standard}}}
+\newcommand{\extended}{\textnormal{\textbf{Extended}}}
 
-\newcommand{\ifcowl}{\textbf{ifcOWL}}
+\newcommand{\ifcowl}{\textnormal{\textbf{ifcOWL}}}
 \newcommand{\ifcsimple}{\ifcowl{}-\simple{}}

### 2015-04-30 article-00-headers.tex
Updates from ShareLaTeX

--- 2015-04-29 article-00-headers.tex
@@
 
-\spnewtheorem{principle}{Principle}{\bfseries}{\rmfamily}
+%\spnewtheorem{principle}{Principle}{\bfseries}{\rmfamily}
+
+\AtBeginDocument{
+    \newtheoremstyle{principleStyle}
+        {1.5\topsep} % Space above, default \topsep
+        {0.5\topsep} % Space below, default \topsep
+        {\itshape} % Body font
+        {} % Indent amount, example: \parindent
+        {\bfseries\itshape} % Theorem head font
+        {.} % Punctuation after theorem head
+        {.5em} % Space after theorem head
+        {} % Theorem head spec (can be left empty, meaning `normal')
+
+    \theoremstyle{principleStyle} \newtheorem{principle}{Principle}
+}
 
 %\spnewtheorem{datasetsRule}{Datasets Rule}{\bfseries}{\rmfamily}
+
+%
+% THEOREMS
+%
+\usepackage{amsthm} % enables \newtheorem{}{}
 
@@
 % ifcOWL layer names
-\newcommand{\simple}{\textbf{Simple}}
-\newcommand{\standard}{\textbf{Standard}}
-\newcommand{\extended}{\textbf{Extended}}
+\newcommand{\bname}[1]{\textnormal{\textbf{#1}}}
+\newcommand{\simple}{\bname{Simple}}
+\newcommand{\standard}{\bname{Standard}}
+\newcommand{\extended}{\bname{Extended}}
 
-\newcommand{\ifcowl}{\textbf{ifcOWL}}
+\newcommand{\ifcowl}{\bname{ifcOWL}}
+\newcommand{\ifcrdf}{\bname{ifcRDF}}
 \newcommand{\ifcsimple}{\ifcowl{}-\simple{}}
@@
 \newcommand{\ifcextended}{\ifcowl{}-\extended{}}
-\newcommand{\ifcrdf}{\textbf{ifcRDF}}
+
+\newcommand{\ifcweb}{\bname{ifcWeb}}
 

### 2015-04-30 article-01-introduction.tex
Updates from ShareLaTeX

@@
 To make BIM models accessible on the Web \cite{torma2014wobd}, we have developed the 
-\name{WebIFC converter} to 
+\bname{WebIFC converter} to 
 derive OWL ontologies from the IFC schema (\ifcowl{}), and RDF datasets from IFC files 
@@
 
-\vspace{2mm}
-% \noindent
-\begin{small}
-\begin{tabular}{|l|l|l|}
+\vspace{0.5\topsep}
+\noindent
+\begin{scriptsize}
+\begin{center}
+    \def\arraystretch{1.2}
+    \setlength{\tabcolsep}{6pt}
+    \begin{tabularx}{0.95\textwidth}{|l|l|X|}
         \hline
-\textbf{Layer} & \textbf{Profiles} & \textbf{Included schema information} \\
+            \textbf{Layer} & \textbf{Profile} & \textbf{Included schema information} \\
         \hline
+            ifcOWL-Simple & OWL 2 EL or & Schema meta data plus: \\
+            & OWL 2 QL or & \ \ - All type definitions and hierarchies \\
+            & OWL 2 RL & \ \ - All entity type properties (names, range types) \\
         \hline 
-\name{ifcOWL-Simple} & OWL 2 EL & Schema meta data \\
-& OWL 2 QL &  All type definitions and hierarchies \\
-& OWL 2 RL & All entity type properties (names, range types) \\
+            ifcOWL-Standard & OWL 2 RL & All above plus: \\
+            & & \ \ - All entity type keys \\
+            & & \ \ - All entity type inverse properties \\
         \hline
-\name{ifcOWL-Standard} & OWL 2 RL & Schema meta data \\
-& & All entity type keys \\
-& & All entity type inverse properties \\
+            ifcOWL-Extended & OWL 2 RL & All above plus: \\
+            & & \ \ - All cardinality constraints \\
         \hline
-\name{ifcOWL-Extended} & OWL 2 RL & All cardinality constraints of properties \\
-& & All cardinality constraints of list types \\
- \hline
-\end{tabular}
-\end{small}
+    \end{tabularx}
+\end{center}
+\end{scriptsize}
 
@@
 \cite{beetz2005ontology,beetz2009ifcowl,pauwels2011interoperability} our approach differs since it
-is based on OWL 2 profiles, produces a flexible multilayer ontology, and is specifically targeted to
+is based on OWL 2 profiles, produces a flexible multilayer ontology, applies \emph{very different methods of data\-type definitions} and is specifically targeted to
 real-life Linked Data applications. It uses stable URIs for entities, in contrast to the previous approaches

### 2015-04-30 article-02-main.tex
Updates from ShareLaTeX

@@
 \label{sec:ifcOWL}
-This section describes the derivation of the ifcOWL ontologies from the IFC specification.
-Since there are different options for the conversion, the choices are presented as conversion principles with justiciations.
+This section describes the derivation of the ifcOWL ontologies from the IFC specification written as an EXPRESS schema.
+An EXPRESS schema defines series of data\-types, functions and rules by using EXPRESS data definition language specified in ISO10303-11 \cite{noauthor:ifc-guide}.
 
-\subsection{Common Conversion Principles}
+\subsection{Common Principles}
 
-\begin{principle}%[Simple OWL profiles]
-\ifcsimple{} should be compatible with as simple OWL 2 profiles as possible (to support a maximum variety of reasoners and other tools). 
-\end{principle}
+Layer \ifcsimple{} should be compatible with with any OWL 2 DL profile (EL, QL or RL) in order to support a maximum variety of reasoners and other tools.
+The other layers \standard{} and \extended{} must be compatible with OWL 2 RL which -- unlike OWL 2 EL or QL -- supports such useful things like enumerations of individuals, union classes and anonymous individuals \cite{motik2013owl}.
 
-The first layer \ifcsimple{} can be specified to be compatible with any OWL 2 DL profile: OWL 2 EL, OWL 2 QL or OWL 2 RL.
-The other layers \ifcstandard{} and \ifcextended{} are compatible with OWL 2 RL.
-OWL 2 RL needed because it -- unlike OWL 2 EL or QL -- supports such necessary things like enumerations of individuals, union classes, disjoint classes, and also anonymous individuals.
-\cite{w3c:owl2-profiles}
+All \ifcowl{} layers must include all type definitions and hierarchies so that any \ifcrdf{} datasets converted from IFC data files could reuse that types without defining new ones.
 
-\begin{principle}%[No domain and range restrictions]
+Conversely, all function and rule definitions, type constraints and derived attributes specified in IFC schema after keywords \name{FUNCTION}, \name{RULE}, \name{WHERE} and \name{RULE} accordingly should not be translated into any \ifcowl{} ontology because that is either impossible, or can involve very complex constraints into \ifcowl{}.
+
 Property names should be kept simple and therefore domain and range constraints (specified by \name{rdfs:domain} and \name{rdfs:range}) should not be included in \ifcowl{} ontologies.
-\end{principle}
+Firstly, there is no need to verify types or to infer types based on properties of \ifcowl{}.
+Secondly, in the IFC schema the property names are \emph{local} and the same property name can appear in multiple
+different classes.
+But in RDF property names are \emph{global} and the basic facilities given by domains and ranges do not provide any direct way to indicate property restrictions that are local to a class \cite{world2014rdf}.
+Finally, direct support for such declarations is provided by OWL constructs, e.g.
+\name{owl:\-onProperty}.
 
-In the IFC schema the property names are \emph{local} to entity classes while in RDF they are \emph{global}. 
-In IFC the same property name can appear in multiple different classes. If the domain and range restrictions
-were included in  \ifcowl{}, the property names would need to be qualified with entity class names. This 
-would reduce the readability and usability of the \ifcowl{} significantly. In \ifcowl{} the domain and 
-range restrictions would mainly be used to infer types, since there is no need to check type constraints 
-like in programming languages \cite{w3c:owl-guide}. Moreover, the basic facilities given by them do not 
-provide any direct way to indicate property restrictions that are local to a class \cite{w3c:rdf-schema}. 
-Although it is possible to combine use \name{rdfs:domain} and \name{rdfs:range} with sub-property hierarchies, 
-direct support for such declarations is provided by OWL constructs, e.g. \name{owl:\-onProperty}.
-
-\begin{principle}%[Namespaces]
-All constructs related to EXPRESS, STEP or IFC specifications must be belong to namespaces shown below with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (instead of \name{IFCXXX} there should be the name of IFC version, e.g.
+All constructs related to EXPRESS, STEP or IFC specifications must belong to separate namespaces with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (instead of \name{IFCXXX} there should be the name of IFC version, e.g.
 \name{IFC4\_ADD1}):
-%\name{IFC2x3} or \name{IFC4}):
-%At this moment the namespaces are defined as below (instead of \name{IFCXXX} there should be IFC versions, e.g. 
-%\name{IFC2x3} or \name{IFC4}):
 
@@
 \end{lstlisting}
-\end{principle}
-
-\begin{principle}All definitions of functions, rules, type constraints and derived attributes specified in EXPRESS after keywords \name{FUNCTION}, \name{RULE}, \name{WHERE} and \name{RULE} accordingly should be ignored.
-\end{principle}
-
-\TODO{Explain why functions, rules, etc. should be ignored}
 
@@
 
-\TODO{\textit{Write about EXPRESS in the Introduction section}:\\
-The IFC specification is written as an EXPRESS schema.
-An EXPRESS schema defines series of data\-types, functions and rules by using EXPRESS data definition language specified in ISO10303-11 \cite{wiki:express,noauthor:ifc-guide}.}
-
 An EXPRESS schema includes six groups of data\-types: built-in simple data\-types, entity data\-types, enumeration data\-types, defined data\-types, select data\-types and aggregation data\-types.
@@
 \subsubsection{Simple Datatypes} built in EXPRESS are: \name{STRING}, \name{BI\-NA\-RY}, \name{IN\-TE\-GER}, \name{REAL}, \name{NUM\-BER}, \name{BOO\-LEAN} and \name{LO\-GI\-CAL}.
-Types \name{STRING} and \name{BINARY} data\-types may be specified with a size parameter: \name{STRING (max\-Length\-In\-Bytes)} or \name{BINARY(size\-In\-Bits)}.
-
-\begin{principle}%[Simple data\-types]
-A simple data\-type (except \emph{BOOLEAN} and \emph{LOGICAL}) should be declared in \ifcowl{} as \name{owl:sameAs} of its most similar primitive XSD data\-type among all data\-types supported by any OWL 2 profile \cite{w3c:owl2-profiles} and preferred in Linked Data Platform \cite{w3c:ldp-best-practices}:
+A simple data\-type (except \name{BOOLEAN} and \name{LOGICAL}) should be declared in \ifcowl{} as \name{owl:sameAs} of its most similar XSD data\-type among data\-types supported by all OWL 2 profiles \cite{motik2013owl} and preferred in Linked Data Platform \cite{ldp-best-practices}:
 
@@
 \end{lstlisting}
-\end{principle}
 
 Type \name{LOGICAL} -- with its three possible values \name{TRUE}, \name{FALSE} and \name{UNKNOWN} -- is considered as an enumeration data\-type.
-The same to \name{BOOLEAN} because OWL 2 EL and QL do not support type \name{xsd:boolean} \cite{w3c:owl2-profiles}.
+The same to \name{BOOLEAN} because OWL 2 EL and QL do not support type \name{xsd:boolean} \cite{motik2013owl}.
 
-Statement `most similar XSD data\-type' means the type from XSD schema that conform to the value spaces of the EXPRESS simple data\-type better than any other type.
-For instance, type \name{xsd:double} is more similar to \name{REAL} than \name{xsd:decimal} as it is an IEEE 64-bit floating-point data\-type and supports scientific notation \cite{w3c:xsd,datapic:xsd}.
+Type \name{xsd:double} is more similar to \name{REAL} than \name{xsd:decimal} as it is an IEEE 64-bit floating-point data\-type and supports scientific notation.
 When it is needed to refer to an EXPRRESS simple data\-type, we recommend to use its XSD equivalent type directly.
 
-\begin{principle}%[Simple data\-types with sizes]
-Size parameters of types \name{STRING} and \name{BINARY} (if any) should be ignored in layers \simple{} and \standard{} and should be declared obviously -- by means of OWL 2 data\-type facets -- in layer \extended{}, for instance:
-
-% they are defined as equivalent to data\-types without size.
-But in layer \extended{}, the size parameter must be declared by using property \name{owl:onDatatype} and data\-type facets \name{xsd:maxLength} or \name{xsd:length} \cite{w3c:owl2-syntax,w3c:xmlschema11-2}, for instance:
+Types \name{STRING} and \name{BINARY} data\-types may be specified with a size parameter: \name{STRING(max\-Length\-In\-Bytes)} or \name{BINARY(size\-In\-Bits)}.
+In layers \simple{} and \standard{} the size should be ignored, while in layer \extended{} it should be declared:
 
@@
 \end{lstlisting}
-\end{principle}
 
@@
 
-\subsubsection{Enumeration Datatypes} represent fixed sets of possible values.
-
-\begin{principle}%[Enumeration data\-types]
-Each enumeration data\-type should be defined in \ifcowl{} as an \name{owl:Class} which is a subclass of \name{expr:\-Enum\-Class}.
-Enumeration values must be defined as named individuals of the enumeration class but in different ways in different layers as shown below:
+\subsubsection{Enumeration Datatypes} represent fixed sets of possible string values.
+Each enumeration data\-type should be defined in \ifcowl{} as an \name{owl:Class} which is a subclass of \name{expr:\-Enum\-Class}\footnote{Hereinafter an object that has property \name{rdfs:subClassOf} is always an \name{owl:Class}}.
+Enumeration values must be defined as named individuals of the enumeration class but in different ways according to the layer:
 
@@
 \end{lstlisting}
-\end{principle}
 
-Two different approaches are used because enumerations with more than one individual or literal are not supported in OWL 2 EL or QL, like enumerations of literals in OWL 2 RL.
+The enumeration values cannot be defined as literals because enumerations of literals is not supported by OWL 2 RL.
+Two different approaches are used because enumerations with more than one individual or literal are not supported in OWL 2 EL or QL.
 The fact that some values -- for instance, \name{ifc:NOT\-DE\-FI\-NED} -- belong to different enumeration data\-types should not cause any problems since any enumeration value is always specified with an exact type.
 
-As mentioned above, simple data\-types \name{BOOLEAN} and \name{LOGICAL} are considered as enumeration data\-types.
-But these types and their values must be defined in the EXPRESS-related namespace, e.g.
+As mentioned, simple data\-types \name{BOOLEAN} and \name{LOGICAL} are considered as enumerations.
+But they and their values must be defined in the EXPRESS-related namespace, e.g.
 \name{expr:LOGICAL}, \name{expr:TRUE}, \name{expr:UNKNOWN}.
@@
 For instance, type \name{Ifc\-Trim\-ming\-Se\-lect} is an union of entity data\-type \name{Ifc\-Car\-te\-sian\-Point} and defined data\-type \name{Ifc\-Pa\-ra\-me\-ter\-Va\-lue}.
-
-\begin{principle}%[Select data\-types]
-A select data\-type must be converted to an \name{owl:Class} which is a subclass of \name{expr:\-Select\-Class}.
-Its particular data\-types are defined as subclasses of the select class, but in different ways depended on the layer:
-\end{principle}
+A select data\-type must be converted to a subclass of \name{expr:\-Select\-Class}.
+Its particular data\-types are defined as subclasses of the select class, but in different ways according to the layer:
 
@@
 
-Two different approaches are used because property \name{owl:unionOf} is supported only in OWL 2 RL, but not in OWL 2 EL and QL \cite{w3c:owl2-profiles}.
+Two different approaches are used because property \name{owl:unionOf} is not supported in OWL 2 EL and OWL 2 QL, but only in OWL 2 RL. \cite{motik2013owl}.
 
@@
 
-\begin{principle}%[Defined data\-types]
-A defined data\-type based on a simple data\-type should be declared as an \name{owl:Class} which is a subclass of \name{expr:\-Defined\-Class} and has a single property \name{rdf:value} with that simple data\-type.
+A defined data\-type based on a simple data\-type should be declared as a subclass of \name{expr:\-Defined\-Class} and has a single property \name{rdf:value} with that simple data\-type.
 A defined data\-type based on another defined data\-type should be declared as a subclass of the base defined data\-type, for instance:
@@
 \end{lstlisting}
-\end{principle}
 
-Defined data\-types are defined as \name{owl:Class}es instead of \name{rdfs:Datatype}s because they can be merged with other \name{owl:Class}es for creating select data\-types.
-In addition, different defined data\-types wrapping the same \name{rdfs:Datatype} are not the same.
-For example, \name{IfcLengthMeasure} is different to \name{IfcMassMeasure} although they both are \name{REAL} and belong to select class \name{IfcMeasureValue}.
+Defined data\-types are declared as \name{owl:Class}es using boxing technique because they must be able to merge with other \name{owl:Class}es for creating select data\-types.
+In addition, values of defined data\-types must be qualified with type name.
+For example, value \name{"5.0"^^xsd:double} \TODO{}
+
+for \name{IfcLengthMeasure} is different to \name{IfcMassMeasure} although they both are \name{REAL} and belong to select class \name{IfcMeasureValue}.
 
@@
 An attribute is defined with a data\-type, cardinalities and can be marked as optional.
-
-\begin{principle}%[Entity data\-types]
 An entity data\-type should be converted to an \name{owl:\-Class} which is subclass of its superclass (if any) or \name{expr:\-Entity\-Class}.
@@
 
-\begin{principle}Any aggregated type must be subclass of one of \name{expr:SetClass}, \name{expr:ListClass}, \name{expr:ArrayClass} and \name{expr:BagClass} which are subclasses of \name{expr:\-Collec\-tion\-Class}.
+Any aggregated type in \ifcowl{} must be subclass of one of \name{expr:SetClass}, \name{expr:ListClass}, \name{expr:ArrayClass} and \name{expr:BagClass} which are subclasses of \name{expr:\-Collec\-tion\-Class}.
 The latter has one a nonfunctional property \name{expr:slot} with type \name{expr:Slot} and four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index}.
@@
 Below is a fragment of definition of type \name{SET [2:?] OF IfcProfileDef}:
-\end{principle}
 
@@
 
-\subsubsection{Dataset Metadata}
+\subsubsection{Header Section} ....
 %\label{subsec:ifcRDF-metadata}
@@
 \end{lstlisting}
+
+\subsubsection{Data Section} ....
 

### 2015-04-30 article-01-introduction.tex
Updates from ShareLaTeX

@@
 
-BIM models could provide a natural framework to organize and manage information about buildings.
-However, a big obstacle is the difficulty to access and utilize BIM models which have traditionally been shared shared through point-to-point file exchange.
+BIM models could provide a natural framework to organize and manage building related information.
+However, a big obstacle is the difficulty to access and utilize BIM models which have traditionally been shared through point-to-point file exchange.
 They cannot be used in an online, granular, and
-interlinked manner and provides a poor basis for open utilization of building information and the 
+interlinked manner. This provides a poor basis for open utilization of building information and the 
 emergence of new, innovative applications.
@@
 there are possibilities and reasons to convert the IFC schema to OWL in several different ways. Firstly, there is a \emph{mismatch} between the modeling constructs of OWL and 
-EXPRESS data definition language used to specify the IFC schema. Secondly, a simple version of \ifcowl{} \emph{compatible with
-restricted OWL profiles} (OWL 2 EL or QL) \cite{motik2009owl} is useful to support ontology
+EXPRESS data definition language \cite{schenck1994information} used to specify the IFC schema. Secondly, a simple version of \ifcowl{} \emph{compatible with
+restricted OWL profiles} (OWL 2 EL or QL) \cite{motik2012owl} is useful to support ontology
 reasoning or inexact reasoning. Thirdly, for most practical purposes only a subset of ontology
 information is required, as no validation or checking of \ifcrdf{} data is needed; it can
-be considered \emph{practically valid} w.r.t. the schema, since BIM tools use certified converters, and icfRDF can sensibly only be used in a
+be considered \emph{practically valid} w.r.t. the schema, since BIM tools use certified converters. Moreover, icfRDF can sensibly only be used in a
 \emph{read-only} fashion, since BIM models will be maintained in the native formats of BIM
-tools. Fourthly, a full conversion of the IFC schema will make \ifcowl{}
-\emph{unnecessarily large}, and difficult to understand or manipulate.
+tools. In addition, a full conversion of the IFC schema produces a huge ontology, difficult to understand or manipulate.
 
@@
 
+CityGML \cite{kolbe2005citygml} is a surface model for buildings, which can be useful to visitors for 
+indoor navigation. IFC contains complementary information about building design, structure, 
+and equipment for a wider range of stakeholders and use cases. Moreover, IFC data can be one source 
+when generating CityGML models. % GityGML has also been converted into OWL.
+
+In the rest of this paper we describe the conversions to \ifcowl{} and \ifcrdf{}, present the deployment of ontologies, example datasets and the converter, and finish with summary and future work.
+
 % \begin{figure}

### 2015-04-30 article-02-main.tex
Updates from ShareLaTeX

@@
 
-\section{Principles of Generating ifcOWL Ontologies And ifcRDF Datasets}
+\section{Generating ifcOWL Ontologies And ifcRDF Datasets}
 \label{sec:ifcOWL}
-This section describes the derivation of the ifcOWL ontologies from the IFC specification written as an EXPRESS schema.
-An EXPRESS schema defines series of data\-types, functions and rules by using EXPRESS data definition language specified in ISO10303-11 \cite{noauthor:ifc-guide}.
+This section describes the derivation of the ifcOWL ontologies from the IFC schema.
+The schema defines series of datatypes, functions and rules by using EXPRESS data definition language (ISO10303-11 \cite{schenck1994information}).
 
@@
 
-Layer \ifcsimple{} should be compatible with with any OWL 2 DL profile (EL, QL or RL) in order to support a maximum variety of reasoners and other tools.
-The other layers \standard{} and \extended{} must be compatible with OWL 2 RL which -- unlike OWL 2 EL or QL -- supports such useful things like enumerations of individuals, union classes and anonymous individuals \cite{motik2013owl}.
+Layer \ifcsimple{} has been designed to be compatible with with any OWL 2 DL profile (EL, QL or RL) in order to support a maximum variety of reasoners and other Semantic Web tools.
+The other layers \standard{} and \extended{} are compatible with OWL 2 RL which -- unlike OWL 2 EL or QL -- supports representational constructs of EXPRESS, such as enumerations of individuals, union classes and anonymous individuals \cite{motik2012owl}.
+All \ifcowl{} layers include all type definitions and hierarchies so that any \ifcrdf{} datasets converted from IFC data files could reuse types without defining new ones.
 
-All \ifcowl{} layers must include all type definitions and hierarchies so that any \ifcrdf{} datasets converted from IFC data files could reuse that types without defining new ones.
+The function and rule definitions, type constraints and derived attributes specified in IFC schema after keywords \name{FUNCTION}, \name{RULE}, and \name{WHERE} are not translated into any \ifcowl{} ontology either because of their procedural nature, or the complexity of the resulting constraints.
 
-Conversely, all function and rule definitions, type constraints and derived attributes specified in IFC schema after keywords \name{FUNCTION}, \name{RULE}, \name{WHERE} and \name{RULE} accordingly should not be translated into any \ifcowl{} ontology because that is either impossible, or can involve very complex constraints into \ifcowl{}.
+One goal in the conversion is to keep the property names simple and understandable.
+Therefore domain and range constraints (specified by \name{rdfs:domain} and \name{rdfs:range}) are not included in \ifcowl{}.
+Firstly, there is no need to verify types or to infer types based on properties of \ifcowl{}.
+Secondly, in the IFC schema the property names are \emph{local} and the same property name can appear in multiple different classes.
+In RDF property names are \emph{global} and the basic facilities given by domains and ranges do not provide any direct way to indicate property restrictions that are local to a class \cite{world2014rdf}.
+%Finally, direct support for such declarations is provided by OWL constructs, e.g. \name{owl:\-onProperty}.
 
-Property names should be kept simple and therefore domain and range constraints (specified by \name{rdfs:domain} and \name{rdfs:range}) should not be included in \ifcowl{} ontologies.
-Firstly, there is no need to verify types or to infer types based on properties of \ifcowl{}.
-Secondly, in the IFC schema the property names are \emph{local} and the same property name can appear in multiple
-different classes.
-But in RDF property names are \emph{global} and the basic facilities given by domains and ranges do not provide any direct way to indicate property restrictions that are local to a class \cite{world2014rdf}.
-Finally, direct support for such declarations is provided by OWL constructs, e.g.
-\name{owl:\-onProperty}.
-
-All constructs related to EXPRESS, STEP or IFC specifications must belong to separate namespaces with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (instead of \name{IFCXXX} there should be the name of IFC version, e.g.
+All constructs related to EXPRESS, STEP or IFC specifications belong to separate namespaces shown with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (instead of \name{IFCXXX} there should be the name of IFC version, e.g.
 \name{IFC4\_ADD1}):
@@
 \subsubsection{Simple Datatypes} built in EXPRESS are: \name{STRING}, \name{BI\-NA\-RY}, \name{IN\-TE\-GER}, \name{REAL}, \name{NUM\-BER}, \name{BOO\-LEAN} and \name{LO\-GI\-CAL}.
-A simple data\-type (except \name{BOOLEAN} and \name{LOGICAL}) should be declared in \ifcowl{} as \name{owl:sameAs} of its most similar XSD data\-type among data\-types supported by all OWL 2 profiles \cite{motik2013owl} and preferred in Linked Data Platform \cite{ldp-best-practices}:
+A simple data\-type (except \name{BOOLEAN} and \name{LOGICAL}) should be declared in \ifcowl{} as \name{owl:sameAs} of its most similar XSD data\-type among data\-types supported by all OWL 2 profiles \cite{motik2012owl} and preferred in Linked Data Platform \cite{ldp-best-practices}:
 
@@
 Type \name{LOGICAL} -- with its three possible values \name{TRUE}, \name{FALSE} and \name{UNKNOWN} -- is considered as an enumeration data\-type.
-The same to \name{BOOLEAN} because OWL 2 EL and QL do not support type \name{xsd:boolean} \cite{motik2013owl}.
+The same to \name{BOOLEAN} because OWL 2 EL and QL do not support type \name{xsd:boolean} \cite{motik2012owl}.
 
-Type \name{xsd:double} is more similar to \name{REAL} than \name{xsd:decimal} as it is an IEEE 64-bit floating-point data\-type and supports scientific notation.
-When it is needed to refer to an EXPRRESS simple data\-type, we recommend to use its XSD equivalent type directly.
+Type \name{xsd:double} is closer to \name{REAL} than \name{xsd:decimal} as it is an IEEE 64-bit floating-point data\-type and supports scientific notation.
+If a simple datatype of EXPRESS is needed, its XSD equivalent type should be used directly.
+% When it is needed to refer to an EXPRESS simple data\-type, we recommend to use its XSD equivalent type directly.
 
@@
 Each enumeration data\-type should be defined in \ifcowl{} as an \name{owl:Class} which is a subclass of \name{expr:\-Enum\-Class}\footnote{Hereinafter an object that has property \name{rdfs:subClassOf} is always an \name{owl:Class}}.
-Enumeration values must be defined as named individuals of the enumeration class but in different ways according to the layer:
+Enumeration values are defined as named individuals of the enumeration class but in different ways according to the layer:
 
@@
 As mentioned, simple data\-types \name{BOOLEAN} and \name{LOGICAL} are considered as enumerations.
-But they and their values must be defined in the EXPRESS-related namespace, e.g.
+But they and their values should be defined in the EXPRESS-related namespace, e.g.
 \name{expr:LOGICAL}, \name{expr:TRUE}, \name{expr:UNKNOWN}.
@@
 For instance, type \name{Ifc\-Trim\-ming\-Se\-lect} is an union of entity data\-type \name{Ifc\-Car\-te\-sian\-Point} and defined data\-type \name{Ifc\-Pa\-ra\-me\-ter\-Va\-lue}.
-A select data\-type must be converted to a subclass of \name{expr:\-Select\-Class}.
+A select data\-type is converted to a subclass of \name{expr:\-Select\-Class}.
 Its particular data\-types are defined as subclasses of the select class, but in different ways according to the layer:
@@
 
-Two different approaches are used because property \name{owl:unionOf} is not supported in OWL 2 EL and OWL 2 QL, but only in OWL 2 RL. \cite{motik2013owl}.
+Two different approaches are used because property \name{owl:unionOf} is not supported in OWL 2 EL and OWL 2 QL, but only in OWL 2 RL. \cite{motik2012owl}.
 
@@
 
-Defined data\-types are declared as \name{owl:Class}es using boxing technique because they must be able to merge with other \name{owl:Class}es for creating select data\-types.
-In addition, values of defined data\-types must be qualified with type name.
-For example, value \name{"5.0"^^xsd:double} \TODO{}
+Defined data\-types are declared as \name{owl:Class}es using boxing technique because they should be able to merge with other \name{owl:Class}es for creating select data\-types.
+In addition, values of defined data\-types are qualified with type name.
+For example, value \name{"5.0"^^xsd:double} for \name{IfcLengthMeasure} is different to \name{IfcMassMeasure} although they both are \name{REAL} and belong to select class \name{IfcMeasureValue}.
 
-for \name{IfcLengthMeasure} is different to \name{IfcMassMeasure} although they both are \name{REAL} and belong to select class \name{IfcMeasureValue}.
-
-In principle type \name{Ifc\-Time\-Stamp} must wrap an integer value which represents so-called Unix time.
-But we decided this value should have type \name{xsd:date\-Time}, and all values of \name{Ifc\-Time\-Stamp} must be converted to a standard datetime values.
+In principle type \name{Ifc\-Time\-Stamp} wraps an integer value which represents so-called Unix time.
+However, we have chosen the type \name{xsd:date\-Time} for these values, and all values of \name{Ifc\-Time\-Stamp} are converted to a standard datetime values.
 
@@
 
-\subsubsection{Entity Datatypes} are identical to classes in object-oriented programming languages.
+\subsubsection{Entity Datatypes} correspond to classes in object-oriented languages.
 Each type may have one super\-type.
@@
 Entity attribute cardinalities can be only from 0 to 1, even if this attribute is a set of values.
-In this case the attribute type is an external aggregated data\-type, for example:
+In this case the attribute type is an external aggregated data\-type, as in:
 
@@
 
-\subsubsection{Aggregated Types} may be unordered (\name{SET} and \name{BAG}) or  ordered (\name{LIST} and \name{ARRAY}) collections.
-\name{BAG} may contain duplicated values, unlike \name{SET}.
+\subsubsection{Aggregated Types} may be unordered (\name{SET} and \name{BAG}) or  ordered (\name{LIST} and \name{ARRAY}) collections. %\name{BAG} may contain duplicated values, unlike \name{SET}. 
 \name{ARRAY} is specified with start and end indexes while other types are specified with min and max cardinalities.
 
-Any aggregated type in \ifcowl{} must be subclass of one of \name{expr:SetClass}, \name{expr:ListClass}, \name{expr:ArrayClass} and \name{expr:BagClass} which are subclasses of \name{expr:\-Collec\-tion\-Class}.
+Any aggregated type in \ifcowl{} is a subclass of \name{expr:SetClass},
+\name{expr:ListClass}, \name{expr:ArrayClass} or \name{expr:BagClass} which are subclasses of \name{expr:\-Collec\-tion\-Class}.
 The latter has one a nonfunctional property \name{expr:slot} with type \name{expr:Slot} and four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index}.
@@
 
-\subsection{Generating ifcRDF Datasets From IFC Data}
+\section{Generating ifcRDF Datasets From IFC Data}
 \label{sec:ifcRDF}
 
-\subsubsection{Header Section} ....
-%\label{subsec:ifcRDF-metadata}
-
-IFC data are exchanged by using ``STEP physical files'' (SPF).
-SPF format is defined in ISO 10303-21 \cite{wiki:step-file}.
-A SPF header section includes a list of STEP-specific metadata entities with types \name{FileDescription} \name{FileName} and \name{FileSchema} described in \cite{buildingSMART:ifc-header}.
+\subsubsection{Header Section} IFC data is exchanged as \emph{STEP physical files} (SPF).
+SPF format is defined in ISO 10303-21 \cite{ISO10303-21}.
+A SPF header section includes a list of STEP-specific metadata entities with types \name{FileDescription} \name{FileName} and \name{FileSchema}.
+%described in \cite{buildingSMART:ifc-header}.
 They are exported as objects of STEP entities, some attributes are duplicated as Dublin Core (DC) metadata annotations, for example:
@@
 
-\subsubsection{Data Section} ....
-
-In a SPF data section each entity is declared with a line number, an entity data\-type and a list of values.
-All entities of \name{Ifc\-Root}-derived types are converted into URI resources with format \name{GUID\_\textless{}guid\textgreater{}} where \name{guid} must be an normal GUID, not a compressed IFC GUID which may include a sign \name{\$}.
-Regarding to other entities, user can choose whether to export them as anonymous individuals or name them in a format like \name{LINE\_\textless{}lineNumber\textgreater{}}.
+\subsubsection{Data Section} In a SPF data section each entity is identified with a line number.
+%, an entity data\-type and a list of values. 
+All entities of \name{Ifc\-Root}-derived types are converted into URI resources with format \name{GUID\_\textless{}guid\textgreater{}} where \name{guid} is a GUID as a hexadecimal string (not including the character \name{\$}).
+Regarding to anonymous IFC entities, it can be chosen whether to export them as blank nodes or create them a URI in a format like \name{LINE\_\textless{}lineNumber\textgreater{}} (in a user selected namespace).
 Note that line numbers are used only for linking entities inside a SPF, and they are unstable.
-Naming is required when the dataset is used together with OWL 2 EL or OWL 2 QL because these profiles do not allow anonymous individuals.
-The namespace of named entities is also defined by user.
-Below is an example of IFC data exported to an RDF dataset which is compatible with OWL 2 RL:
+Naming is required when the dataset is used together with OWL 2 EL/QL as these profiles do not allow anonymous individuals.
+Below is an example of IFC data exported to an RDF dataset compatible with OWL 2 RL:
 
@@
 
-\section{Conclusion}
+\section{Deployment}
+
+The \textbf{WebIFC converter}, \ifcowl{} ontologies, and example \ifcrdf{} datasets (together with the original IFC files) are available for use from the addresses:
+
+\vspace{2mm}
+\begin{tabular}{|l|l|}
+\hline
+Ontologies: & \texttt{http://drumbeat.cs.hut.fi/owl/} \\
+\hline
+Example datasets & \\
+- IFC files: & \texttt{http://drumbeat.cs.hut.fi/ifc/} \\
+- RDF datasets: & \texttt{http://drumbeat.cs.hut.fi/rdf/} \\
+\hline
+Converter & \\
+- service: & \texttt{http://drumbeat.cs.hut.fi/home/} \\
+- open source: & \texttt{https://github.com/Web-of-Building-Data/Ifc2Rdf} \\
+\hline
+\end{tabular}
+
+\section{Summary and future work}
 \label{sec:conclusion}
+
+We have presented \ifcowl{}, a multilayer ontology of building designs based on the widely adopted IFC standard. The ontology is 
+derived from the IFC schema with the \texfbf{WebIFC} converter that can also produce RDF datasets from IFC files.
+
+The \ifcowl{} ontologies and \ifcrdf{} datasets are currently used in a variety of Linked Data oriented use cases \cite{torma2013semantic} developed in the DRUMBEAT project.
+We are exploring the possibilities to use the WebIFC converter in the generation of metadata for \ifcrdf{} datasets and, in some cases, also linksets between datasets.

### 2015-04-30 article-01-introduction.tex
Updates from ShareLaTeX
Edited by Seppo

@@
 \centering
-\includegraphics[angle=0,width=0.9\textwidth]{images/ifcOWL-multilayers.png}
+\includegraphics[angle=0,width=1.0\textwidth]{images/ifcOWL-multilayers.png}
 %\includegraphics{images/ifcOWL-multilayers.jpg}
@@
 \cite{beetz2005ontology,beetz2009ifcowl,pauwels2011interoperability} our approach differs since it
-is based on OWL 2 profiles, produces a flexible multilayer ontology, applies \emph{very different methods of data\-type definitions} and is specifically targeted to
+is based on OWL 2 profiles, produces a flexible multilayer ontology, applies other methods of data\-type definitions and is specifically targeted to
 real-life Linked Data applications. It uses stable URIs for entities, in contrast to the previous approaches

### 2015-04-30 article-02-main.tex
Edited by Seppo

@@
 
-\section{Generating ifcOWL Ontologies And ifcRDF Datasets}
+\section{Generating ifcOWL Ontologies from the IFC Schema}
 \label{sec:ifcOWL}
@@
 
-\subsection{Common Principles}
+\subsection{Basic Principles}
 
-Layer \ifcsimple{} has been designed to be compatible with with any OWL 2 DL profile (EL, QL or RL) in order to support a maximum variety of reasoners and other Semantic Web tools.
-The other layers \standard{} and \extended{} are compatible with OWL 2 RL which -- unlike OWL 2 EL or QL -- supports representational constructs of EXPRESS, such as enumerations of individuals, union classes and anonymous individuals \cite{motik2012owl}.
+Layer \ifcsimple{} has been designed to be compatible with any OWL 2 DL profile (EL, QL or RL) in order to support a maximum variety of reasoners and other Semantic Web tools.
+The other layers \standard{} and \extended{} are compatible with OWL 2 RL which -- unlike OWL 2 EL or QL -- supports many important representational constructs of EXPRESS, such as enumerations of individuals, union classes and anonymous individuals \cite{motik2012owl}.
 All \ifcowl{} layers include all type definitions and hierarchies so that any \ifcrdf{} datasets converted from IFC data files could reuse types without defining new ones.
 
-The function and rule definitions, type constraints and derived attributes specified in IFC schema after keywords \name{FUNCTION}, \name{RULE}, and \name{WHERE} are not translated into any \ifcowl{} ontology either because of their procedural nature, or the complexity of the resulting constraints.
+The function and rule definitions, type constraints and derived attributes specified in the IFC schema after keywords \name{FUNCTION}, \name{RULE}, and \name{WHERE} are not translated into any \ifcowl{} layer either because of their procedural nature, or the complexity of the resulting constraints.
 
@@
 
-All constructs related to EXPRESS, STEP or IFC specifications belong to separate namespaces shown with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (instead of \name{IFCXXX} there should be the name of IFC version, e.g.
+All constructs related to EXPRESS, STEP or IFC specifications belong to separate namespaces shown below with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (instead of \name{IFCXXX} there should be the name of IFC version, e.g.
 \name{IFC4\_ADD1}):
@@
 Types \name{STRING} and \name{BINARY} data\-types may be specified with a size parameter: \name{STRING(max\-Length\-In\-Bytes)} or \name{BINARY(size\-In\-Bits)}.
-In layers \simple{} and \standard{} the size should be ignored, while in layer \extended{} it should be declared:
+In layers \simple{} and \standard{} the size should be ignored, while in layer \extended{} it can be declared:
 
@@
 
-The enumeration values cannot be defined as literals because enumerations of literals is not supported by OWL 2 RL.
+The enumeration values cannot be defined as literals because enumerations of literals are not supported by OWL 2 RL.
 Two different approaches are used because enumerations with more than one individual or literal are not supported in OWL 2 EL or QL.
@@
 \subsubsection{Select Datatypes} represent unions of other data\-types.
-Particular members of a select data\-type can be entity, defined or select data\-types.
+A particular member of a select data\-type can be an entity, or a defined or a select data\-type.
 For instance, type \name{Ifc\-Trim\-ming\-Se\-lect} is an union of entity data\-type \name{Ifc\-Car\-te\-sian\-Point} and defined data\-type \name{Ifc\-Pa\-ra\-me\-ter\-Va\-lue}.
@@
 
-Defined data\-types are declared as \name{owl:Class}es using boxing technique because they should be able to merge with other \name{owl:Class}es for creating select data\-types.
-In addition, values of defined data\-types are qualified with type name.
-For example, value \name{"5.0"^^xsd:double} for \name{IfcLengthMeasure} is different to \name{IfcMassMeasure} although they both are \name{REAL} and belong to select class \name{IfcMeasureValue}.
+Defined data\-types are declared as \name{owl:Class}es using boxing technique because if they were \name{rdfs:Datatype}s they could not be be united with other \name{owl:Class}es for creating select data\-types.
+In addition, values of defined data\-types must be qualified with type names.
+For example, a \name{xsd:double} value with type \name{IfcLengthMeasure} is not the same as with type \name{IfcMassMeasure} although these types belong to select class \name{IfcMeasureValue} and defined as equal to \name{REAL}.
 
-In principle type \name{Ifc\-Time\-Stamp} wraps an integer value which represents so-called Unix time.
-However, we have chosen the type \name{xsd:date\-Time} for these values, and all values of \name{Ifc\-Time\-Stamp} are converted to a standard datetime values.
+In IFC type \name{Ifc\-Time\-Stamp} is defined as \name{INTEGER} which represents so-called Unix time value.
+However, it was decided that these values must be converted to standard datetime values during reading IFC files.
+Thus, \name{ifc:Ifc\-Time\-Stamp} wraps a \name{xsd:date\-Time} value instead of \name{xsd:integer}.
 
@@
 An entity data\-type should be converted to an \name{owl:\-Class} which is subclass of its superclass (if any) or \name{expr:\-Entity\-Class}.
-Entity attribute cardinalities can be only from 0 to 1, even if this attribute is a set of values.
-In this case the attribute type is an external aggregated data\-type, as in:
+It was decided to make all entity properties be functional (the maximum cardinality is 1).
+
+Entity attribute cardinalities can be only from 0 to 1, even if this attribute is a set of values. In this case the attribute type is an external aggregated data\-type, as in:
 
@@
 \name{expr:ListClass}, \name{expr:ArrayClass} or \name{expr:BagClass} which are subclasses of \name{expr:\-Collec\-tion\-Class}.
-The latter has one a nonfunctional property \name{expr:slot} with type \name{expr:Slot} and four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index}.
+The latter has one nonfunctional property \name{expr:slot} with type \name{expr:Slot} and four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index}.
 Class \name{expr:Slot} has two functional properties \name{expr:item} (the actual collection item value), \name{expr:index} and also optional functional properties \name{expr:previous}, \name{expr:next}.
@@
 %, an entity data\-type and a list of values. 
-All entities of \name{Ifc\-Root}-derived types are converted into URI resources with format \name{GUID\_\textless{}guid\textgreater{}} where \name{guid} is a GUID as a hexadecimal string (not including the character \name{\$}).
-Regarding to anonymous IFC entities, it can be chosen whether to export them as blank nodes or create them a URI in a format like \name{LINE\_\textless{}lineNumber\textgreater{}} (in a user selected namespace).
+All entities of \name{Ifc\-Root}-derived types have a GUID in IFC.
+They are converted into URI resources with format \name{GUID\_\textless{}guid\textgreater{}} where \name{guid} is in the hexadecimal string representation (not including the character \name{\$}).
+Regarding to anonymous IFC entities, it can be chosen whether to export them as blank nodes or create them a URI in a format \name{LINE\_\textless{}lineNumber\textgreater{}} (in a user defined namespace).
 Note that line numbers are used only for linking entities inside a SPF, and they are unstable.

### 2015-04-30 article-01-introduction.tex
Created new figure

@@
 To make BIM models accessible on the Web \cite{torma2014wobd}, we have developed the 
-\bname{WebIFC converter} to 
+\bname{IFC2LD converter} to 
 derive OWL ontologies from the IFC schema (\ifcowl{}), and RDF datasets from IFC files 
@@
 \centering
-\includegraphics[angle=0,width=1.0\textwidth]{images/ifcOWL-multilayers.png}
-%\includegraphics{images/ifcOWL-multilayers.jpg}
-\caption{Multiple layers of \ifcowl{} ontology}
+% \includegraphics[angle=0,width=1.0\textwidth]{images/ifcOWL-multilayers.png}
+\includegraphics{images/ifcOWL-multilayers-6.png}
+%\caption{Multiple layers of \ifcowl{} ontology}
+\caption{Producing IFC ontologies and datasets for use in Linked Data Platform}
 \label{fig:ifcOWL-layers}
@@
 In the rest of this paper we describe the conversions to \ifcowl{} and \ifcrdf{}, present the deployment of ontologies, example datasets and the converter, and finish with summary and future work.
-
-% \begin{figure}
-% \centering
-% %\def\svgwidth{\textwidth}
-% \def\svgscale{1}
-% \input{images/image.pdf_tex}
-% %{\includesvg{images/test-2}}
-% %\includegraphics[width=0.75\textwidth]{images/ifcOWL-multilayer-2.png}
-% \caption{A multilayer ifcOWL ontology}
-% \label{fig:ifcOWL-multilayered}
-% \end{figure}

### 2015-04-30 article-02-main.tex
Created new figure

@@
 
-The \textbf{WebIFC converter}, \ifcowl{} ontologies, and example \ifcrdf{} datasets (together with the original IFC files) are available for use from the addresses:
+The \textbf{IFC2LD converter}, \ifcowl{} ontologies, and example \ifcrdf{} datasets (together with the original IFC files) are available for use from the addresses:
 
@@
 We have presented \ifcowl{}, a multilayer ontology of building designs based on the widely adopted IFC standard. The ontology is 
-derived from the IFC schema with the \texfbf{WebIFC} converter that can also produce RDF datasets from IFC files.
+derived from the IFC schema with the \texfbf{IFC2LD} converter that can also produce RDF datasets from IFC files.
 
 The \ifcowl{} ontologies and \ifcrdf{} datasets are currently used in a variety of Linked Data oriented use cases \cite{torma2013semantic} developed in the DRUMBEAT project.
-We are exploring the possibilities to use the WebIFC converter in the generation of metadata for \ifcrdf{} datasets and, in some cases, also linksets between datasets.
+We are exploring the possibilities to use the IFC2LD converter in the generation of metadata for \ifcrdf{} datasets and, in some cases, also linksets between datasets.

### 2015-04-30 article-02-main.tex
First version submitted

@@
 
-All constructs related to EXPRESS, STEP or IFC specifications belong to separate namespaces shown below with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (instead of \name{IFCXXX} there should be the name of IFC version, e.g.
+All constructs related to EXPRESS, STEP or IFC specifications belong to separate namespaces shown below with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (\name{IFCXXX} should be replaced by the name of an IFC version, e.g.
 \name{IFC4\_ADD1}):
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Entity Datatype Declarations
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsubsection{Entity Datatypes} correspond to classes in object-oriented languages.
-Each type may have one super\-type.
-An attribute is defined with a data\-type, cardinalities and can be marked as optional.
-An entity data\-type should be converted to an \name{owl:\-Class} which is subclass of its superclass (if any) or \name{expr:\-Entity\-Class}.
-It was decided to make all entity properties be functional (the maximum cardinality is 1).
-
-Entity attribute cardinalities can be only from 0 to 1, even if this attribute is a set of values. In this case the attribute type is an external aggregated data\-type, as in:
-
-%\begin{lstlisting}[caption={Entity data\-types}, label=lst:ifcOWL-defined-types]
-\begin{lstlisting}
-ifc:IfcProperty a owl:Class ; rdfs:subClassOf expr:EntityClass .
-ifc:IfcComplexProperty a owl:Class ; rdfs:subClassOf ifc:IfcProperty ;
-    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:coordinates;
-                      owl:allValuesFrom ifc:LIST_1_3_IfcLengthMeasure ] ;
-    # only for Extended layer
-    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:coordinates;
-                      owl:cardinality "1"^^xsd:integer ] .
-
-# only for Standard & Extended layers
-[] a owl:AllDisjointClasses ;
-    owl:members ( ifc:IfcComplexProperty ifc:IfcSimpleProperty ) .
-\end{lstlisting}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Aggregated Datatype Declarations
@@
 Any aggregated type in \ifcowl{} is a subclass of \name{expr:SetClass},
-\name{expr:ListClass}, \name{expr:ArrayClass} or \name{expr:BagClass} which are subclasses of \name{expr:\-Collec\-tion\-Class}.
-The latter has one nonfunctional property \name{expr:slot} with type \name{expr:Slot} and four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index}.
+\name{expr:ListClass}, \name{expr:ArrayClass} or \name{expr:BagClass}.
+Their common superclass \name{expr:\-Collec\-tion\-Class} has one nonfunctional property \name{expr:slot} with type \name{expr:Slot} and four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index}.
 Class \name{expr:Slot} has two functional properties \name{expr:item} (the actual collection item value), \name{expr:index} and also optional functional properties \name{expr:previous}, \name{expr:next}.
@@
 
-% In order to declare these types, new classes were created in a similar way as Ordered List Ontoloty \cite{olo:ontology}:
-% \begin{itemize}
-% \item Class \name{expr:\-Collec\-tion\-Class} has four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index} and a nonfunctional property \name{expr:slot} with type \name{expr:Slot}.
-% \item Class \name{expr:Slot} has two functional properties \name{expr:item} (the actual collection item value), \name{expr:index} and also optional functional properties \name{expr:previous}, \name{expr:next}.
-% \item Classes \name{expr:SetClass}, \name{expr:ListClass}, \name{expr:ArrayClass} and \name{expr:BagClass} are subclasses of \name{expr:\-Collec\-tion\-Class}.
-% \end{itemize}
-
-% Below is a fragment of definition of type \name{SET [2:?] OF IfcProfileDef}:
-
 \begin{lstlisting}
 # all layers:
- ifc:SET_2_UNDEFINED_OF_IfcProfileDef a owl:Class ;
-    rdfs:subClassOf ifc:SET_2_UNDEFINED , ifc:SET_OF_IfcProfileDef .
- ifc:SET_2_UNDEFINED, SET_OF_IfcProfileDef rdfs:subClassOf expr:SetClass .
+ ifc:SET_2_UNLIMITED_OF_IfcProfileDef a owl:Class ;
+    rdfs:subClassOf ifc:SET_2_UNLIMITED , ifc:SET_OF_IfcProfileDef .
+ ifc:SET_2_UNLIMITED, SET_OF_IfcProfileDef rdfs:subClassOf expr:SetClass .
 
 # only for layer Extended:
- ifc:SET_2_UNDEFINED rdfs:subClassOf rdfs:subClassOf [ a owl:Restriction ;
+ ifc:SET_2_UNLIMITED rdfs:subClassOf rdfs:subClassOf [ a owl:Restriction ;
     owl:onProperty expr:slot ; owl:minCardinality "2"^^xsd:integer ] .
@@
  ifc:SLOT_OF_IfcProfileDef rdfs:subClassOf [ a owl:Restriction ; owl:onProperty expr:item ; owl:allValuesFrom ifc:IfcProfileDef ] .
+\end{lstlisting}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Entity Datatype Declarations
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsubsection{Entity Datatypes} correspond to classes in object-oriented languages.
+Each type may have one super\-type.
+An attribute is defined with a data\-type, can be a list or a set with \emph{any} cardinalities and can be marked as optional.
+An entity data\-type should be converted to a subclass of its super\-class (\name{expr:\-Entity\-Class} is the most generic).
+It was decided to make maximum cardinality of any entity property be \name{1}.
+In case if the attribute type is a list or a set then a new external aggregated data\-type is created.
+Below is an example for a type with attributes \name{Profiles : SET [2:?] OF IfcProfileDef} and \name{Label : OPTIONAL IfcLabel}:
+
+%\begin{lstlisting}[caption={Entity data\-types}, label=lst:ifcOWL-defined-types]
+\begin{lstlisting}
+ifc:IfcCompositeProfileDef a owl:Class ; rdfs:subClassOf ifc:IfcProfileDef ;
+  # only for Standard, Extended layers
+    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:profiles;
+                      owl:allValuesFrom ifc:SET_2_UNLIMITED_IfcProfileDef ] ;
+    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:label;
+                      owl:allValuesFrom ifc:Label ] ;
+  # only for Extended layer
+    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:profiles;
+                      owl:cardinality "1"^^xsd:integer ] ;
+    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:label;
+                      owl:minCardinality "0"^^xsd:integer ] ;
+    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:label;
+                      owl:maxCardinality "1"^^xsd:integer ] .
 \end{lstlisting}
@@
 
-\subsubsection{Header Section} IFC data is exchanged as \emph{STEP physical files} (SPF).
-SPF format is defined in ISO 10303-21 \cite{ISO10303-21}.
-A SPF header section includes a list of STEP-specific metadata entities with types \name{FileDescription} \name{FileName} and \name{FileSchema}.
+\subsubsection{Header Section.} IFC data is exchanged as \emph{STEP physical files} (SPF) format of which is defined in ISO 10303-21 \cite{ISO10303-21}.
+A SPF header section includes a list of STEP-specific metadata entities with types \name{FileDescription}, \name{FileName} and \name{FileSchema}.
 %described in \cite{buildingSMART:ifc-header}.
@@
 
-\subsubsection{Data Section} In a SPF data section each entity is identified with a line number.
+\subsubsection{Data Section.} In a SPF data section each entity is identified with a line number.
 %, an entity data\-type and a list of values. 
 All entities of \name{Ifc\-Root}-derived types have a GUID in IFC.
-They are converted into URI resources with format \name{GUID\_\textless{}guid\textgreater{}} where \name{guid} is in the hexadecimal string representation (not including the character \name{\$}).
-Regarding to anonymous IFC entities, it can be chosen whether to export them as blank nodes or create them a URI in a format \name{LINE\_\textless{}lineNumber\textgreater{}} (in a user defined namespace).
+They are converted into URI resources with format \name{GUID\_\emph{\textless{}guid\textgreater{}}} where \emph{guid} is in the hexadecimal string representation (not including the character \name{\$}).
+Regarding to anonymous IFC entities, it can be chosen whether to export them as blank nodes or create them a URI in a format \name{LINE\_\emph{\textless{}lineNumber\textgreater{}}} (in a user defined namespace).
 Note that line numbers are used only for linking entities inside a SPF, and they are unstable.
@@
 \begin{lstlisting}
-:GUID_1916794F-5605-4499-9AB0-F155DE8D3B6C
+:GUID_1916794F-5605-4499-9AB0-F155DE8D3B6C     # entity as an URI
     a            ifc:IfcPropertySet ;          # entity type
-    ifc:globalId    [ a ifc:IfcGloballyUniqueId ; # defined type
+    ifc:globalId [ a ifc:IfcGloballyUniqueId ; # property with defined type
                    rdf:value "0P5dbFLWL4cPgmyLNUZJji"^^xsd:string ] ;
-    ifc:hasProperties  _:b11 .
-_:b11 a ifc:SET_1_UNLIMITED_OF_IfcProperty ; # entity type
+    ifc:hasProperties  _:b11 .                 # property w/ aggregated type
+_:b11 a ifc:SET_1_UNLIMITED_OF_IfcProperty ;
     expr:size "3"^^xsd:integer ;
@@
 _:b12 a ifc:SLOT_OF_IfcProperty ;
-    expr:item [ a ifc:IfcPropertySingleValue ; # entity type
+    expr:item [ a ifc:IfcPropertySingleValue ;  # entity as a blank node
                 ifc:name "initial_GUID"^^xsd:string ;
@@
 
-\vspace{2mm}
-\begin{tabular}{|l|l|}
+% \vspace{0.5\topsep}
+\noindent
+\begin{scriptsize}
+\begin{center}
+    \def\arraystretch{1.2}
+    \setlength{\tabcolsep}{6pt}
+    \begin{tabularx}{0.90\textwidth}{|l|X|}
         \hline
@@
         \hline
-Example datasets & \\
-- IFC files: & \texttt{http://drumbeat.cs.hut.fi/ifc/} \\
+        Example files & \\
+        - IFC data: & \texttt{http://drumbeat.cs.hut.fi/ifc/} \\
         - RDF datasets: & \texttt{http://drumbeat.cs.hut.fi/rdf/} \\
         \hline
-Converter & \\
-- service: & \texttt{http://drumbeat.cs.hut.fi/home/} \\
-- open source: & \texttt{https://github.com/Web-of-Building-Data/Ifc2Rdf} \\
+        IFC2LD Converter & \\
+        - online service: & \texttt{http://drumbeat.cs.hut.fi/home/} \\
+        - download binary: & \texttt{http://drumbeat.cs.hut.fi/ifc2ld/} \\
+        - source code: & \texttt{https://github.com/Web-of-Building-Data/Ifc2Rdf} \\
         \hline
-\end{tabular}
+    \end{tabularx}
+\end{center}
+\end{scriptsize}
 
-\section{Summary and future work}
-\label{sec:conclusion}
+\section{Summary and future works}
+\label{sec:summary}
 
 We have presented \ifcowl{}, a multilayer ontology of building designs based on the widely adopted IFC standard. The ontology is 
-derived from the IFC schema with the \texfbf{IFC2LD} converter that can also produce RDF datasets from IFC files.
+derived from the IFC schema with the \ifcld{} converter that can also produce RDF datasets from IFC files.
 
 The \ifcowl{} ontologies and \ifcrdf{} datasets are currently used in a variety of Linked Data oriented use cases \cite{torma2013semantic} developed in the DRUMBEAT project.
-We are exploring the possibilities to use the IFC2LD converter in the generation of metadata for \ifcrdf{} datasets and, in some cases, also linksets between datasets.
+We are exploring the possibilities to use the \ifcld{} converter in the generation of metadata for \ifcrdf{} datasets and, in some cases, also linksets between datasets.

### 2015-05-01 article-00-headers.tex
Updates from ShareLaTeX
Created new figure
First version submitted
Changed bibliography style to 'splncs'
Version 1.2

--- 2015-04-29 article-00-headers.tex
@@
 \usepackage{graphicx}   % allows \includegraphics
+
 \usepackage{cite}       % enables autonumbering in cites
@@
 
-\spnewtheorem{principle}{Principle}{\bfseries}{\rmfamily}
+%\spnewtheorem{principle}{Principle}{\bfseries}{\rmfamily}
+
+% \AtBeginDocument{
+%     \newtheoremstyle{principleStyle}
+%         {1.5\topsep} % Space above, default \topsep
+%         {0.5\topsep} % Space below, default \topsep
+%         {\itshape} % Body font
+%         {} % Indent amount, example: \parindent
+%         {\bfseries\itshape} % Theorem head font
+%         {.} % Punctuation after theorem head
+%         {.5em} % Space after theorem head
+%         {} % Theorem head spec (can be left empty, meaning `normal')
+
+%     \theoremstyle{principleStyle} \newtheorem{principle}{Principle}
+% }
+
+\usepackage{titlesec}   % enables \titlespacing
+
+\titlespacing*{\section}
+    {0pt}{\baselineskip}{0.5\baselineskip}    % {<left>}{<before-sep>}{<after-sep>}
+
+\titlespacing*{\subsection}
+    {0pt}{0.8\baselineskip}{0.5\baselineskip}
+
+\titlespacing*{\subsubsection}
+    {0pt}{0.8\baselineskip}{0.5em}
 
 %\spnewtheorem{datasetsRule}{Datasets Rule}{\bfseries}{\rmfamily}
+
+%
+% THEOREMS
+%
+%\usepackage{amsthm} % enables \newtheorem{}{}
 
@@
 
-\graphicspath{{images/}}
+% \graphicspath{{images/}}
+
+\usepackage{caption}    % enables \captionsetup for tables and figures
+
+\AtBeginDocument{
+    \captionsetup{font=small}
+    \captionsetup{labelfont=bf}
+    % \captionsetup[table]{aboveskip=0pt}
+    \captionsetup[table]{belowskip=0pt}
+
+    % \setlength{\textfloatsep}{10pt}  % distance between floats on the top or the bottom and the text;
+    % \setlength{\floatsep}{0pt}      % distance between two floats;
+    \setlength{\intextsep}{\baselineskip}     % distance between floats inserted inside the page text (using h) and the text proper.
+
+%     The default values in the article document class with the 10pt option are:
+% \textfloatsep: 20.0pt plus 2.0pt minus 4.0pt;
+% \floatsep: 12.0pt plus 2.0pt minus 2.0pt;
+% \intextsep: 12.0pt plus 2.0pt minus 2.0pt.
+}
+
+%
+% REFERENCES
+%
+
+% \usepackage[sorting=none]{biblatex} % http://www.ctan.org/tex-archive/help/Catalogue/entries/biblatex.html
 
@@
 % ifcOWL layer names
-\newcommand{\simple}{\textbf{Simple}}
-\newcommand{\standard}{\textbf{Standard}}
-\newcommand{\extended}{\textbf{Extended}}
+\newcommand{\bname}[1]{\textnormal{\textbf{#1}}}
 
-\newcommand{\ifcowl}{\textbf{ifcOWL}}
+\newcommand{\simple}{\bname{Simple}}
+\newcommand{\standard}{\bname{Standard}}
+\newcommand{\extended}{\bname{Extended}}
+
+\newcommand{\ifcowl}{\bname{ifc\-OWL}}
+\newcommand{\ifcrdf}{\bname{ifc\-RDF}}
 \newcommand{\ifcsimple}{\ifcowl{}-\simple{}}
@@
 \newcommand{\ifcextended}{\ifcowl{}-\extended{}}
-\newcommand{\ifcrdf}{\textbf{ifcRDF}}
+
+\newcommand{\ifcconverter}{\bname{IFC2LD}}
+\newcommand{\ifcld}{\ifcconverter}
 
@@
 %
-\titlerunning{Building ontology}  % abbreviated title (for running head)
+% \titlerunning{Building ontology}  % abbreviated title (for running head)
 %                                     also used for the TOC unless
@@
 
-The volume of data produced in the construction industry is rapidly growing but, as of now, it is not available for open and flexible access and utilization.
-We present an ontology derived from the conceptualization of IFC, the existing standard for representing building designs, and a tool to convert building designs from IFC files to RDF datasets that conform to the ontology.
-Replacing the traditional point-to-point exchange of IFC files with publication of RDF datasets using Linked Data principles and tools enables information integration across different IFC-based and other smart city datasets, and the granular access to individual components.
-It paves the way to a distributed and loosely coupled architecture for building information management, and provides opportunities for innovative applications.
+The volume of data produced in the construction industry is rapidly growing but, as of now, it is not available for open, flexible access and utilization.
+We propose a new approach to the automatic derivation of a building ontology from the conceptualization of Industry Foundation Classes (IFC), the existing standard for representing building designs.
+The resulted multilayer ontology and the tool for converting building designs from IFC files to RDF datasets that conform to the ontology are introduced.
+Replacing the traditional point-to-point exchange of IFC files with publication of RDF datasets using Linked Data tools enables information integration across different IFC-based and other smart city datasets, and the granular access to individual components.
+It paves the way to a distributed, loosely coupled architecture for building information management, and provides opportunities for innovative applications.
 
@@
 
-\input{article-03-bibliography.tex}
+% According to LNCS manual, the Acknowledgements section must be declared by using \subsubsection{}
+
+\subsubsection{Acknowledgements.} This work has been carried out at Aalto University in research projects DRUM (RYM/PRE 2010-2014) and DRUMBEAT (2014-2017), funded by Tekes, Aalto University, and the participating companies.
+
+%\input{article-03-bibliography.tex}
+
+% \bibliographystyle{splncs03}
+\bibliographystyle{splncs}
+\bibliography{article.bib}
+
+% \printbibliography{}
 

### 2015-05-01 article-01-introduction.tex
Version 1.2

@@
 
-Currently, huge amounts of information about building designs is produced  
-using Building Information Modeling (BIM) tools \cite{eastman2011bim}. A widely adopted standard, 
-IFC (Industry Foundation Classes) \cite{ISO16739,liebich2010unveiling} provides a common representation for BIM models of 
+Currently, huge amounts of building designs information is produced by 
+using Building Information Modeling (BIM) tools \cite{eastman2011bim}. 
+IFC \cite{ISO16739,liebich2010unveiling} -- a widely adopted standard -- provides a common representation for BIM models of 
 different design disciplines. All major BIM tools support IFC and can export BIM models to IFC files for exchange with other parties.
@@
 (\ifcrdf{}). The \ifcowl{} specifies concepts, relations, and constraints for \ifcrdf{} datasets that replicate the content of IFC files. 
-All terms defined in \ifcowl{}  must thus be based on the corresponding constructs in 
+Thus, all terms defined in \ifcowl{}  must be based on the corresponding constructs in 
 the IFC schema.
@@
 without information loss.
-Ideally, there would also be just one \ifcowl{} ontology for each version of the IFC schema. However,
-there are possibilities and reasons to convert the IFC schema to OWL in several different ways. Firstly, there is a \emph{mismatch} between the modeling constructs of OWL and 
-EXPRESS data definition language \cite{schenck1994information} used to specify the IFC schema. Secondly, a simple version of \ifcowl{} \emph{compatible with
+Ideally, there would be just one \ifcowl{} ontology for each version of the IFC schema. However,
+there are possibilities and reasons to convert the IFC schema to OWL in several different ways. Firstly, there is a \emph{mismatch} between the modeling constructs of OWL and the 
+EXPRESS data definition language \cite{schenck1994information} used to specify the IFC schema. Secondly, a simple version of \ifcowl{} that is \emph{compatible with
 restricted OWL profiles} (OWL 2 EL or QL) \cite{motik2012owl} is useful to support ontology
 reasoning or inexact reasoning. Thirdly, for most practical purposes only a subset of ontology
-information is required, as no validation or checking of \ifcrdf{} data is needed; it can
-be considered \emph{practically valid} w.r.t. the schema, since BIM tools use certified converters. Moreover, icfRDF can sensibly only be used in a
+information is required, as no validation or checking of \ifcrdf{} data is needed; it is considered \emph{practically valid} w.r.t.
+the schema, since BIM tools use certified converters.
+Moreover, \ifcrdf{} can sensibly only be used in a
 \emph{read-only} fashion, since BIM models will be maintained in the native formats of BIM
@@
 in a way that \emph{\ifcrdf{} is compatible with any of the layers} (Fig. \ref{fig:ifcOWL-layers}). 
-New representational constructs are introduced at more comprehensive levels which therefore 
-provide increasing detailed type information about \ifcrdf{} data:
+New representational constructs are introduced at more comprehensive levels which, therefore, 
+provide increasing detailed type information about \ifcrdf{} data (Table \ref{tab:ifcOWL-layers}).
 
-\vspace{0.5\topsep}
-\noindent
-\begin{scriptsize}
-\begin{center}
-    \def\arraystretch{1.2}
-    \setlength{\tabcolsep}{6pt}
+\begin{table}
+    \caption{Contents of ifcOWL layers}
+    \label{tab:ifcOWL-layers}
+    \centering
+    \scriptsize
+    \def\arraystretch{1.2}          % vertical space
+    \setlength{\tabcolsep}{6pt}     % horizontal space
     \begin{tabularx}{0.95\textwidth}{|l|l|X|}
         \hline
-            \textbf{Layer} & \textbf{Profile} & \textbf{Included schema information} \\
+            \textbf{Layer} & \textbf{Compatible profile} & \textbf{Contents} \\
         \hline
-            ifcOWL-Simple & OWL 2 EL or & Schema meta data plus: \\
-            & OWL 2 QL or & \ \ - All type definitions and hierarchies \\
-            & OWL 2 RL & \ \ - All entity type properties (names, range types) \\
+            ifcOWL-Simple & OWL 2 EL or & Ontology meta data plus: \\
+            & OWL 2 QL or & \ \ - All basic constructs \\
+            & OWL 2 RL & \ \ - All type definitions and hierarchies \\
         \hline 
             ifcOWL-Standard & OWL 2 RL & All above plus: \\
-            & & \ \ - All entity type keys \\
-            & & \ \ - All entity type inverse properties \\
+            & & \ \ - All entity type properties and keys \\
         \hline
@@
     \end{tabularx}
-\end{center}
-\end{scriptsize}
+\end{table}
 
@@
 
-From the previous IFC-to-OWL/RDF conversions
-\cite{beetz2005ontology,beetz2009ifcowl,pauwels2011interoperability} our approach differs since it
+Our approach differs from the previous IFC-to-OWL/RDF conversions
+\cite{beetz2005ontology,beetz2009ifcowl,pauwels2011interoperability}. It
 is based on OWL 2 profiles, produces a flexible multilayer ontology, applies other methods of data\-type definitions and is specifically targeted to
 real-life Linked Data applications. It uses stable URIs for entities, in contrast to the previous approaches
-where the URIs were instable across models versions.
+where the URIs were unstable across models versions.
 
-CityGML \cite{kolbe2005citygml} is a surface model for buildings, which can be useful to visitors for 
+CityGML \cite{kolbe2005citygml}, a surface model for buildings, can be useful to visitors for 
 indoor navigation. IFC contains complementary information about building design, structure, 
@@
 
-In the rest of this paper we describe the conversions to \ifcowl{} and \ifcrdf{}, present the deployment of ontologies, example datasets and the converter, and finish with summary and future work.
+\TODO{Update the description of sections}
+
+In the rest of this paper, we describe the conversions to \ifcowl{} and \ifcrdf{}, present the deployment of ontologies, example datasets and the converter, and finish with summary and future work.

### 2015-05-01 article-02-main.tex
Version 1.2

@@
 \label{sec:ifcOWL}
-This section describes the derivation of the ifcOWL ontologies from the IFC schema.
+This section describes the derivation of the ifcOWL ontologies from an IFC schema.
 The schema defines series of datatypes, functions and rules by using EXPRESS data definition language (ISO10303-11 \cite{schenck1994information}).
@@
 
-Layer \ifcsimple{} has been designed to be compatible with any OWL 2 DL profile (EL, QL or RL) in order to support a maximum variety of reasoners and other Semantic Web tools.
-The other layers \standard{} and \extended{} are compatible with OWL 2 RL which -- unlike OWL 2 EL or QL -- supports many important representational constructs of EXPRESS, such as enumerations of individuals, union classes and anonymous individuals \cite{motik2012owl}.
+The \ifcsimple{} layer has been designed to be compatible with any OWL 2 DL profile (EL, QL or RL) in order to support a maximum variety of reasoners and other Semantic Web tools.
+The other layers, i.e.
+\standard{} and \extended{}, are compatible with OWL 2 RL which -- unlike OWL 2 EL or QL -- supports many important representational constructs of EXPRESS, such as enumerations, union classes and anonymous resources \cite{motik2012owl}.
 All \ifcowl{} layers include all type definitions and hierarchies so that any \ifcrdf{} datasets converted from IFC data files could reuse types without defining new ones.
@@
 One goal in the conversion is to keep the property names simple and understandable.
-Therefore domain and range constraints (specified by \name{rdfs:domain} and \name{rdfs:range}) are not included in \ifcowl{}.
+Therefore, domain and range constraints (specified by \name{rdfs:domain} and \name{rdfs:range}) are not included in \ifcowl{}.
 Firstly, there is no need to verify types or to infer types based on properties of \ifcowl{}.
-Secondly, in the IFC schema the property names are \emph{local} and the same property name can appear in multiple different classes.
-In RDF property names are \emph{global} and the basic facilities given by domains and ranges do not provide any direct way to indicate property restrictions that are local to a class \cite{world2014rdf}.
+Secondly, while in the IFC schema the same property name can appear in multiple different classes, in RDF, property names are \emph{global} and the basic facilities given by domains and ranges do not provide any direct way to indicate that properties are \emph{local} to a class \cite{world2014rdf}.
 %Finally, direct support for such declarations is provided by OWL constructs, e.g. \name{owl:\-onProperty}.
 
-All constructs related to EXPRESS, STEP or IFC specifications belong to separate namespaces shown below with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (\name{IFCXXX} should be replaced by the name of an IFC version, e.g.
+All constructs related to EXPRESS, STEP or IFC specifications belong to separate namespaces shown below with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (\name{IFCxxx} should be replaced by the name of an IFC version, e.g.
 \name{IFC4\_ADD1}):
@@
 @prefix step: <http://drumbeat.cs.hut.fi/owl/STEP#>
-@prefix ifc: <http://drumbeat.cs.hut.fi/owl/IFCXXX#>
+@prefix ifc:  <http://drumbeat.cs.hut.fi/owl/IFCxxx#>
 \end{lstlisting}
@@
 
-An EXPRESS schema includes six groups of data\-types: built-in simple data\-types, entity data\-types, enumeration data\-types, defined data\-types, select data\-types and aggregation data\-types.
+An EXPRESS schema includes six groups of data\-types, namely: \emph{simple}, \emph{entity}, \emph{enumeration}, \emph{defined}, \emph{select}, and \emph{aggregation} data\-types.
 
@@
 
-\subsubsection{Simple Datatypes} built in EXPRESS are: \name{STRING}, \name{BI\-NA\-RY}, \name{IN\-TE\-GER}, \name{REAL}, \name{NUM\-BER}, \name{BOO\-LEAN} and \name{LO\-GI\-CAL}.
-A simple data\-type (except \name{BOOLEAN} and \name{LOGICAL}) should be declared in \ifcowl{} as \name{owl:sameAs} of its most similar XSD data\-type among data\-types supported by all OWL 2 profiles \cite{motik2012owl} and preferred in Linked Data Platform \cite{ldp-best-practices}:
+\subsubsection{Simple Datatypes} built in EXPRESS are: \name{STRING}, \name{BI\-NA\-RY}, \name{IN\-TE\-GER}, \name{REAL}, \name{NUM\-BER}, \name{BOO\-LEAN}, and \name{LO\-GI\-CAL}.
+Any simple data\-type, except \name{BOOLEAN} and \name{LOGICAL}, should be declared in \ifcowl{} as \name{owl:sameAs} of its most similar XSD data\-type among the data\-types that are supported by all OWL 2 profiles \cite{motik2012owl} and preferred in Linked Data Platform \cite{ldp-best-practices}:
 
@@
 Type \name{LOGICAL} -- with its three possible values \name{TRUE}, \name{FALSE} and \name{UNKNOWN} -- is considered as an enumeration data\-type.
-The same to \name{BOOLEAN} because OWL 2 EL and QL do not support type \name{xsd:boolean} \cite{motik2012owl}.
+The same to \name{BOOLEAN} because OWL 2 EL and QL do not support data\-type \name{xsd:boolean} \cite{motik2012owl}.
 
-Type \name{xsd:double} is closer to \name{REAL} than \name{xsd:decimal} as it is an IEEE 64-bit floating-point data\-type and supports scientific notation.
-If a simple datatype of EXPRESS is needed, its XSD equivalent type should be used directly.
+XSD type \name{xsd:double} is closer to \name{REAL} than \name{xsd:decimal} as it is an IEEE 64-bit floating-point data\-type and supports scientific notation.
+Whenever a simple data\-type of EXPRESS is needed, its XSD equivalent type should be used directly.
 % When it is needed to refer to an EXPRESS simple data\-type, we recommend to use its XSD equivalent type directly.
 
-Types \name{STRING} and \name{BINARY} data\-types may be specified with a size parameter: \name{STRING(max\-Length\-In\-Bytes)} or \name{BINARY(size\-In\-Bits)}.
-In layers \simple{} and \standard{} the size should be ignored, while in layer \extended{} it can be declared:
+Types \name{STRING} and \name{BINARY} data\-types may be specified with a size parameter: \name{STRING(\emph{max\-Length\-In\-Bytes})} or \name{BINARY(\emph{size\-In\-Bits})}.
+It should be ignored in the \simple{} and \standard{} layers, and must be declared in the \extended{} layer:
 
@@
 \begin{lstlisting}
-# only for Simple, Standard layers
+# only for the Simple, Standard layers
   expr:STRING22 owl:sameAs xsd:string .
-# only for Extended layer
+# only for the Extended layer
   expr:STRING22 owl:equivalentClass [ a rdfs:Datatype ;
-    owl:onDatatype  xsd:string; owl:withRestrictions ( [ xsd:length 22 ] ) ].
+    owl:onDatatype xsd:string; owl:withRestrictions ( [ xsd:length 22 ] ) ] .
 \end{lstlisting}
@@
 \subsubsection{Enumeration Datatypes} represent fixed sets of possible string values.
-Each enumeration data\-type should be defined in \ifcowl{} as an \name{owl:Class} which is a subclass of \name{expr:\-Enum\-Class}\footnote{Hereinafter an object that has property \name{rdfs:subClassOf} is always an \name{owl:Class}}.
-Enumeration values are defined as named individuals of the enumeration class but in different ways according to the layer:
+Each enumeration data\-type should be defined in \ifcowl{} as a \emph{subclass}\footnote{Any class that is a subclass of another one hereinafter is an \name{owl:Class}.} of \name{expr:\-Enum\-Class}.
+Enumeration values are defined as \emph{named individuals} of the enumeration class, but in different ways, according to the layer, since enumerations with more than one individual or literal are not supported in OWL 2 EL or QL:
 
@@
 \begin{lstlisting}
-# only for Simple layer
+# only for the Simple layer
   ifc:IfcAssemblyPlaceEnum a owl:Class ; rdfs:subclassOf expr:EnumClass .
-  ifc:SITE, ifc:FACTORY, ifc:NOTDEFINED rdf:type ifc:IfcAddressTypeEnum .
-# only for Standard, Extended layers
+  ifc:SITE, ifc:FACTORY, ifc:NOTDEFINED a ifc:IfcAddressTypeEnum .
+# only for the Standard, Extended layers
   ifc:IfcAssemblyPlaceEnum a owl:Class ; rdfs:subclassOf expr:EnumClass ;
@@
 
-The enumeration values cannot be defined as literals because enumerations of literals are not supported by OWL 2 RL.
-Two different approaches are used because enumerations with more than one individual or literal are not supported in OWL 2 EL or QL.
-The fact that some values -- for instance, \name{ifc:NOT\-DE\-FI\-NED} -- belong to different enumeration data\-types should not cause any problems since any enumeration value is always specified with an exact type.
+The enumeration values are not defined as literals because enumerations of literals are not supported by OWL 2 RL.
+The fact that some values -- for instance, \name{ifc:NOT\-DE\-FI\-NED} -- belong to different enumeration data\-types should not cause any problems as an enumeration value can be assigned only to an entity attribute with known enumeration data\-type.
 
-As mentioned, simple data\-types \name{BOOLEAN} and \name{LOGICAL} are considered as enumerations.
-But they and their values should be defined in the EXPRESS-related namespace, e.g.
-\name{expr:LOGICAL}, \name{expr:TRUE}, \name{expr:UNKNOWN}.
+As mentioned above, simple data\-types \name{BOOLEAN} and \name{LOGICAL} are considered as enumerations.
+However, they and their values should be defined in the EXPRESS-related namespace, e.g.
+\name{expr:LOGICAL}, \name{expr:TRUE}, \name{expr:UNKNOWN}, etc.
 
@@
 \subsubsection{Select Datatypes} represent unions of other data\-types.
-A particular member of a select data\-type can be an entity, or a defined or a select data\-type.
+A particular member of a select data\-type can be either an entity, or a defined, or a select data\-type.
 For instance, type \name{Ifc\-Trim\-ming\-Se\-lect} is an union of entity data\-type \name{Ifc\-Car\-te\-sian\-Point} and defined data\-type \name{Ifc\-Pa\-ra\-me\-ter\-Va\-lue}.
-A select data\-type is converted to a subclass of \name{expr:\-Select\-Class}.
-Its particular data\-types are defined as subclasses of the select class, but in different ways according to the layer:
+A select data\-type is converted into a subclass of \name{expr:\-Select\-Class}.
+Its particular data\-types are defined as subclasses of the select class, but in different ways, depended on the layer, since property \name{owl:unionOf} is not supported in OWL 2 EL and OWL 2 QL \cite{motik2012owl}:
 
@@
 \begin{lstlisting}
-# only for Simple layer
+# only for the Simple layer
   ifc:IfcTrimmingSelect a owl:Class ; rdfs:subclassOf expr:SelectClass .
@@
         rdfs:subclassOf ifc:IfcTrimmingSelect .
-# only for Standard, Extended layers
+# only for the Standard, Extended layers
   ifc:IfcTrimmingSelect a owl:Class ; rdfs:subclassOf expr:SelectClass ;
@@
 \end{lstlisting}
-
-Two different approaches are used because property \name{owl:unionOf} is not supported in OWL 2 EL and OWL 2 QL, but only in OWL 2 RL. \cite{motik2012owl}.
 
@@
 
-\subsubsection{Defined (Declared, Named) Datatypes} are usually specified as equivalent or subclass of simple data\-types or other defined data\-types.
-For instance, type \name{Ifc\-Length\-Mea\-sure} is equal to \name{REAL}, while type \name{Ifc\-Po\-si\-ti\-ve\-Length\-Mea\-su\-re} is a subset of \name{Ifc\-Length\-Mea\-sure} with some constraint.
+\subsubsection{Defined (Declared, Named) Datatypes} are usually specified as equivalent sets or subsets of simple data\-types or other defined data\-types.
+For instance, \name{Ifc\-Length\-Mea\-sure} is equal to \name{REAL}, while \name{Ifc\-Po\-si\-ti\-ve\-Length\-Mea\-su\-re} is a subset of \name{Ifc\-Length\-Mea\-sure}, specified by some constraint rule.
 
-A defined data\-type based on a simple data\-type should be declared as a subclass of \name{expr:\-Defined\-Class} and has a single property \name{rdf:value} with that simple data\-type.
+A defined data\-type must not be converted into a \name{rdfs:Datatype} even if it is equivalent to a simple data\-type.
+Let us suppose, there is an attribute with a select type \name{IfcMeasureValue} and it has a \name{xsd:double} value, then it is hard to say whether this is a \name{IfcLengthMeasure} value, a \name{IfcMassMeasure} value, or something else.
+Furthermore, it is not allowed to use a \name{rdfs:Datatype} as a member class of a select class like \name{ifc:ifcMeasureValue}.
+
+Thus, a defined data\-type based on a simple data\-type should be converted into an \name{expr:\-Defined\-Class}' subclass which has only property \name{rdf:value} with range is the base simple data\-type.
 A defined data\-type based on another defined data\-type should be declared as a subclass of the base defined data\-type, for instance:
@@
 ifc:IfcLengthMeasure a owl:Class ; rdfs:subClassOf expr:DefinedClass ;
+    # only for the Standard, Extended layers
     rdfs:subClassOf [ a owl:Restriction; owl:onProperty rdf:value;
@@
 
-Defined data\-types are declared as \name{owl:Class}es using boxing technique because if they were \name{rdfs:Datatype}s they could not be be united with other \name{owl:Class}es for creating select data\-types.
-In addition, values of defined data\-types must be qualified with type names.
-For example, a \name{xsd:double} value with type \name{IfcLengthMeasure} is not the same as with type \name{IfcMassMeasure} although these types belong to select class \name{IfcMeasureValue} and defined as equal to \name{REAL}.
-
-In IFC type \name{Ifc\-Time\-Stamp} is defined as \name{INTEGER} which represents so-called Unix time value.
+In IFC, type \name{Ifc\-Time\-Stamp} is defined as \name{INTEGER} which represents so-called Unix time value.
 However, it was decided that these values must be converted to standard datetime values during reading IFC files.
-Thus, \name{ifc:Ifc\-Time\-Stamp} wraps a \name{xsd:date\-Time} value instead of \name{xsd:integer}.
+Thus, \name{ifc:Ifc\-Time\-Stamp} wraps a \name{xsd:date\-Time} property instead of \name{xsd:integer}.
 
@@
 \name{expr:ListClass}, \name{expr:ArrayClass} or \name{expr:BagClass}.
-Their common superclass \name{expr:\-Collec\-tion\-Class} has one nonfunctional property \name{expr:slot} with type \name{expr:Slot} and four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index}.
-Class \name{expr:Slot} has two functional properties \name{expr:item} (the actual collection item value), \name{expr:index} and also optional functional properties \name{expr:previous}, \name{expr:next}.
-Below is a fragment of definition of type \name{SET [2:?] OF IfcProfileDef}:
+Their common superclass \name{expr:\-Collec\-tion\-Class} has one \emph{nonfunctional} property \name{expr:slot} with type \name{expr:Slot} and four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, and \name{expr:\-end\-Index}.
+Class \name{expr:Slot} has functional properties \name{expr:item}, \name{expr:index}, and optional functional properties \name{expr:previous}, \name{expr:next}.
+A fragment of the definition of type \name{SET [2:?] OF IfcProfileDef} is shown as follows:
 
@@
 
-# only for layer Extended:
+# only for the Extended layer:
  ifc:SET_2_UNLIMITED rdfs:subClassOf rdfs:subClassOf [ a owl:Restriction ;
@@
 Each type may have one super\-type.
-An attribute is defined with a data\-type, can be a list or a set with \emph{any} cardinalities and can be marked as optional.
-An entity data\-type should be converted to a subclass of its super\-class (\name{expr:\-Entity\-Class} is the most generic).
-It was decided to make maximum cardinality of any entity property be \name{1}.
-In case if the attribute type is a list or a set then a new external aggregated data\-type is created.
-Below is an example for a type with attributes \name{Profiles : SET [2:?] OF IfcProfileDef} and \name{Label : OPTIONAL IfcLabel}:
+Thus, an entity data\-type should be converted into a subclass of its super\-class or of \name{expr:\-Entity\-Class}.
+For instance, a type with attributes \name{Profiles : SET [2:?] OF IfcProfileDef} and \name{Label : OPTIONAL IfcLabel} is defined as follow:
 
@@
 
+The minimum property cardinality of \name{ifc:label} is 0 because it is optional.
+The maximum cardinality of any property is 1.
+If an attribute type is a set (or a list), then a new specific aggregation must be defined, for example, \name{SET\_\-2\_\-UNLIMITED\_\-IfcProfileDef} is created to specify property \name{ifc:profiles}.
+This approach allows to represent lists, sets and arrays in the same manner and provides possibilities to constraint the attribute cardinalities.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Generating ifcRDF Datasets From IFC Data}
+\section{Generating ifcRDF Datasets from IFC Data}
 \label{sec:ifcRDF}

### 2015-05-01 article-01-introduction.tex
Version 1.5

@@
 \section{Introduction}
-\label{sec:Introduction}
+\label{sec:introduction}
 
@@
 
-CityGML \cite{kolbe2005citygml}, a surface model for buildings, can be useful to visitors for 
-indoor navigation. IFC contains complementary information about building design, structure, 
-and equipment for a wider range of stakeholders and use cases. Moreover, IFC data can be one source 
-when generating CityGML models. % GityGML has also been converted into OWL.
+% CityGML \cite{kolbe2005citygml}, a surface model for buildings, can be useful to visitors for 
+% indoor navigation. IFC contains complementary information about building design, structure, 
+% and equipment for a wider range of stakeholders and use cases. Moreover, IFC data can be one source 
+% when generating CityGML models. % GityGML has also been converted into OWL.
 
-\TODO{Update the description of sections}
+% \TODO{Update the description of sections}
 
-In the rest of this paper, we describe the conversions to \ifcowl{} and \ifcrdf{}, present the deployment of ontologies, example datasets and the converter, and finish with summary and future work.
+% In the rest of this paper, we describe the conversions to \ifcowl{} and \ifcrdf{}, present the deployment of ontologies, example datasets and the converter, and finish with summary and future work.
+
+In the Sect.
+\ref{sec:ifcOWL}, we describe the principles of IFC-EXPRESS-to-OWL conversion.
+Sect.
+\ref{sec:ifcRDF} provides some sample fragments of IFC-SPF-to-RDF conversion; where SPF (STEP Physical File) defined in ISO 10303-21 \cite{ISO10303-21} is the main exchange format of IFC data.
+Sect.
+\ref{sec:deployment} includes short descriptions about public access to the \textbf{IFC2LD} converter and the resulting ifcOWL ontologies.

### 2015-05-01 article-02-main.tex
Version 1.5

@@
 Thus, an entity data\-type should be converted into a subclass of its super\-class or of \name{expr:\-Entity\-Class}.
-For instance, a type with attributes \name{Profiles : SET [2:?] OF IfcProfileDef} and \name{Label : OPTIONAL IfcLabel} is defined as follow:
+
+The minimum cardinality of an optional entity property is \name{0}.
+The maximum cardinality is always \name{1}.
+If an attribute type is set, list, or array, then a new aggregation data\-type is defined, for example, \name{ifc:\-SET\_\-2\_\-UNLIMITED\_\-IfcProfileDef} was created for property \name{Profiles : SET [2:?] OF IfcProfileDef}.
+This approach allows to represent all multiple attributes in the same manner and provides possibilities to constrain the cardinalities, even when the minimum cardinality is not \name{0} or \name{1}.
+
+% Below is an example of definition of a type with attributes \name{Profiles : SET [2:?] OF IfcProfileDef} and \name{Label : OPTIONAL IfcLabel}:
 
@@
 ifc:IfcCompositeProfileDef a owl:Class ; rdfs:subClassOf ifc:IfcProfileDef ;
-  # only for Standard, Extended layers
+  # only for the Standard, Extended layers
     rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:profiles;
@@
                       owl:allValuesFrom ifc:Label ] ;
-  # only for Extended layer
+  # only for the Extended layer
     rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:profiles;
@@
 
-The minimum property cardinality of \name{ifc:label} is 0 because it is optional.
-The maximum cardinality of any property is 1.
-If an attribute type is a set (or a list), then a new specific aggregation must be defined, for example, \name{SET\_\-2\_\-UNLIMITED\_\-IfcProfileDef} is created to specify property \name{ifc:profiles}.
-This approach allows to represent lists, sets and arrays in the same manner and provides possibilities to constraint the attribute cardinalities.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\subsubsection{Header Section.} IFC data is exchanged as \emph{STEP physical files} (SPF) format of which is defined in ISO 10303-21 \cite{ISO10303-21}.
-A SPF header section includes a list of STEP-specific metadata entities with types \name{FileDescription}, \name{FileName} and \name{FileSchema}.
-%described in \cite{buildingSMART:ifc-header}.
-They are exported as objects of STEP entities, some attributes are duplicated as Dublin Core (DC) metadata annotations, for example:
+As mentioned in Sect.
+\ref{sec:introduction}, IFC data is exchanged in STEP Physical Files (SPF).
+The IFC-SPF-to-RDF conversion is aimed to produce \ifcrdf{} datasets including \emph{all} information stored in SPFs and compatible with any \ifcowl{} layer.
+
+\subsubsection{Header Section} includes STEP-specific metadata as normal data-section entities.
+Their types -- \name{FileDescription}, \name{FileName}, and \name{FileSchema} -- are defined in STEP.
+They are exported as normal entities, but some attributes of them are duplicated as Dublin Core (DC) metadata annotations, for example:
 
@@
 
-\subsubsection{Data Section.} In a SPF data section each entity is identified with a line number.
+\subsubsection{Data Section} includes entities, each of which is identified with a line number.
 %, an entity data\-type and a list of values. 
@@
 
-\section{Deployment}
+\section{Implementation and Deployment}
+\label{sec:deployment}
 
-The \textbf{IFC2LD converter}, \ifcowl{} ontologies, and example \ifcrdf{} datasets (together with the original IFC files) are available for use from the addresses:
+The algorithm of generating \ifcowl{} ontologies and \ifcrdf{} datasets, including the principles described in Sect.
+\ref{sec:ifcOWL}-\ref{sec:ifcRDF}, are implemented in our Java program, namely \textbf{IFC2LD converter}.
+One can use its command-line interface, web interface or application programming interface for converting any IFC schema or data file into \ifcowl{} or \ifcrdf{}.
+By changing the configuration file, one can also force the converter to adapt various OWL profiles starting from OWL Lite.
 
-% \vspace{0.5\topsep}
-\noindent
-\begin{scriptsize}
-\begin{center}
-    \def\arraystretch{1.2}
-    \setlength{\tabcolsep}{6pt}
+The \textbf{IFC2LD converter}, \ifcowl{} ontologies including different layers for different IFC schema versions, and some sample \ifcrdf{} datasets (together with the original IFC files) are available on the Internet (Table \ref{tab:ifc2ld-web}).
+
+\begin{table}
+    \caption{Contents of ifcOWL layers}
+    \label{tab:ifc2ld-web}
+    \centering
+    \scriptsize
+    \def\arraystretch{1.2}          % vertical space
+    \setlength{\tabcolsep}{6pt}     % horizontal space
     \begin{tabularx}{0.90\textwidth}{|l|X|}
@@
         - download binary: & \texttt{http://drumbeat.cs.hut.fi/ifc2ld/} \\
-        - source code: & \texttt{https://github.com/Web-of-Building-Data/Ifc2Rdf} \\
+        % - source code: & \texttt{https://github.com/Web-of-Building-Data/Ifc2Rdf} \\
         \hline
     \end{tabularx}
-\end{center}
-\end{scriptsize}
+\end{table}
 
@@
 
-We have presented \ifcowl{}, a multilayer ontology of building designs based on the widely adopted IFC standard. The ontology is 
+In this paper, we proposed a new approach and principles of creating \ifcowl{}, a multilayer ontology of building designs based on the widely adopted IFC standard. The ontology is 
 derived from the IFC schema with the \ifcld{} converter that can also produce RDF datasets from IFC files.
 
-The \ifcowl{} ontologies and \ifcrdf{} datasets are currently used in a variety of Linked Data oriented use cases \cite{torma2013semantic} developed in the DRUMBEAT project.
-We are exploring the possibilities to use the \ifcld{} converter in the generation of metadata for \ifcrdf{} datasets and, in some cases, also linksets between datasets.
+The \ifcowl{} ontologies and \ifcrdf{} datasets are currently used in a variety of Linked-Data-oriented use cases \cite{torma2013semantic} which are developed in the DRUMBEAT project.
+We are exploring the possibilities to use the \ifcld{} converter in the terms of generation of metadata for linksets between \ifcrdf{} datasets.

### 2015-05-01 article-01-introduction.tex
Version 1.6 (edited by Seppo)

@@
 
-Currently, huge amounts of building designs information is produced by 
-using Building Information Modeling (BIM) tools \cite{eastman2011bim}. 
-IFC \cite{ISO16739,liebich2010unveiling} -- a widely adopted standard -- provides a common representation for BIM models of 
+Currently, huge amounts of building designs information is produced
+using Building Information Modeling (BIM) tools \cite{eastman2011bim}. Industry Foundation Classes 
+(IFC) \cite{ISO16739,liebich2010unveiling} -- a widely adopted standard -- provides a common representation for BIM models of 
 different design disciplines. All major BIM tools support IFC and can export BIM models to IFC files for exchange with other parties.
@@
 Consequently, we convert the IFC schema into a \emph{flexible, multilayer \ifcowl{} ontology}
-in a way that \emph{\ifcrdf{} is compatible with any of the layers} (Fig. \ref{fig:ifcOWL-layers}). 
-New representational constructs are introduced at more comprehensive levels which, therefore, 
-provide increasing detailed type information about \ifcrdf{} data (Table \ref{tab:ifcOWL-layers}).
+in a way that \emph{\ifcrdf{} is compatible with any of the layers} (Fig. \ref{fig:ifcOWL-layers}), and in which 
+%New representational constructs are included at 
+the more comprehensive levels % which, therefore, 
+provide increasing detailed type information about the same \ifcrdf{} data (Table \ref{tab:ifcOWL-layers}).
 
-\begin{table}
+\begin{table}[b]
     \caption{Contents of ifcOWL layers}
@@
 
+It should be noted that surface models targeted to building users can be shared in CityGML \cite{kolbe2005citygml}.
+IFC, however, provides complementary, original information about building design for a wider range of stakeholders and use cases.
+%, and can also be one source when generating CityGML models.
+
 % CityGML \cite{kolbe2005citygml}, a surface model for buildings, can be useful to visitors for 

### 2015-05-01 article-02-main.tex
Version 1.6 (edited by Seppo)

@@
 Therefore, domain and range constraints (specified by \name{rdfs:domain} and \name{rdfs:range}) are not included in \ifcowl{}.
-Firstly, there is no need to verify types or to infer types based on properties of \ifcowl{}.
-Secondly, while in the IFC schema the same property name can appear in multiple different classes, in RDF, property names are \emph{global} and the basic facilities given by domains and ranges do not provide any direct way to indicate that properties are \emph{local} to a class \cite{world2014rdf}.
+There is no need to verify types or to infer types based on properties of \ifcowl{}.
+Moreover, while in the IFC schema the same property name can appear in multiple different classes, in RDF, property names are \emph{global} and the basic facilities given by domains and ranges do not provide any direct way to indicate that properties are \emph{local} to a class \cite{world2014rdf}.
 %Finally, direct support for such declarations is provided by OWL constructs, e.g. \name{owl:\-onProperty}.

### 2015-05-01 article-02-main.tex
Version 2.0 (final)

@@
 All entities of \name{Ifc\-Root}-derived types have a GUID in IFC.
-They are converted into URI resources with format \name{GUID\_\emph{\textless{}guid\textgreater{}}} where \emph{guid} is in the hexadecimal string representation (not including the character \name{\$}).
+They are converted into URI resources with format \name{GUID\_\emph{\textless{}guid\textgreater{}}}; where \emph{guid} is in the hexadecimal string representation (not including the character \name{\$}).
 Regarding to anonymous IFC entities, it can be chosen whether to export them as blank nodes or create them a URI in a format \name{LINE\_\emph{\textless{}lineNumber\textgreater{}}} (in a user defined namespace).
@@
 
-\begin{table}
-    \caption{Contents of ifcOWL layers}
+\begin{table}[b]
+    \caption{Internet links to the related research results}
     \label{tab:ifc2ld-web}
